```{r}
# bookdown::render_book("index.Rmd", "bookdown::gitbook")
```

# Figure 1: Integration of Ex Vivo Drug Screening and Transcriptomics

This chapter describes a comprehensive analysis integrating ex vivo cytarabine (Ara-C) response profiles with transcriptomic data from AML patient cohorts, including BeatAML and RJAML. The goal is to identify predictive biomarkers of Ara-C sensitivity and characterize gene expression/pathway enrichment differences between clinical response groups.

## Ara-C Response Profiling in BeatAML

We first assessed the distribution of Ara-C sensitivity across patients using the BeatAML dataset. Patients were classified into sensitivity groups based on the z-score transformed area under the curve (AUC) values for cytarabine.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

```{r plot-arac-sensitivity-distribution, eval=TRUE, message = FALSE, warning = FALSE}
library(ggplot2)

sample1 <- read.csv("../02.data/BeatAML_AraC_value_sampleInfo.csv", header = TRUE)
sample1$z_auc <- scale(sample1$auc)
sample1$group <- ifelse(sample1$z_auc < 0, "Sensitive", "Resistant")

# Visualize the distribution using ggplot2
p <- ggplot(sample1, aes(x = z_auc)) +
  geom_histogram(binwidth = 0.3, fill = "#3182bd", color = "black") +
  labs(
    title = "Distribution of Ara-C Sensitivity (Z-scored AUC)",
    x = "Z-Scaled AUC", y = "Number of Patients"
  ) +
  theme_minimal(base_size = 13)

p

ggsave(p, filename = "../03.out/figure1/beataml_auc_zscore.pdf", width = 5.4, height = 5)
```

## Identifying Ara-C-Responsive Genes by Partial Correlation

To identify genes whose expression levels are associated with cytarabine sensitivity, we calculated partial correlations between gene expression and Ara-C AUC, adjusting for bone marrow blast percentage.

```{r partial-correlation-analysis, eval=FALSE, message = FALSE, warning = FALSE}
# Load libraries
library(pbapply)
library(dplyr)
library(tibble)
library(readr)
library(stringr)
library(stats)

# ========== 1. Load and preprocess metadata ==========
# Load AUC info
sample_info <- read_csv("../02.data/BeatAML_AraC_value_sampleInfo.csv")

arac <- sample_info %>%
  select(Sample, auc) %>%
  rename(AraC_AUC = auc) %>%
  column_to_rownames("Sample")

# Load BM blast data
bm_blast <- read.table("../02.data/beataml_bm_blasts.txt", header = TRUE) %>%
  arrange(Sample) %>%
  column_to_rownames("Sample")

# ========== 2. Load and filter expression matrix ==========
exp_raw <- read.table("../02.data/BeatAML_all_735_TPM_matrix.txt", header = TRUE, row.names = 1)

# Keep only meaningful gene types
valid_types <- c(
  "protein_coding", "antisense", "lincRNA", "3prime_overlapping_ncrna",
  "macro_lncRNA", "non_coding", "processed_transcript", "TEC", "scaRNA",
  "sense_intronic", "sense_overlapping", "transcribed_processed_pseudogene",
  "transcribed_unitary_pseudogene", "transcribed_unprocessed_pseudogene",
  "unprocessed_pseudogene"
)

exp_filtered <- exp_raw %>%
  filter(Type %in% valid_types)

# Get common samples
common_samples <- Reduce(intersect, list(colnames(exp_filtered)[-c(1:2)], rownames(arac), rownames(bm_blast)))

# Subset and log2-transform expression
exp_log2 <- exp_filtered[, common_samples, drop = FALSE] %>%
  as.matrix() %>%
  t() %>%
  { log2(. + 1) }

# Keep genes with ≥25% non-NA and non-zero variance
exp_log2 <- exp_log2[, apply(exp_log2, 2, function(x) {
  sum(!is.na(x)) >= (0.25 * length(x)) && sd(x, na.rm = TRUE) > 0
})]

# Ensure sample order consistency
exp_log2 <- exp_log2[common_samples, ]
arac <- arac[common_samples, , drop = FALSE]
bm_blast <- bm_blast[common_samples, , drop = FALSE]

# ========== 3. Define partial correlation function ==========
partial_correlation <- function(gene_exp, auc, blast) {
  df <- na.omit(data.frame(Gene = gene_exp, AraC_AUC = auc, BM_Blast = blast))
  if (nrow(df) <= 2) return(c(NA, NA))
  
  r12 <- cor(df$Gene, df$AraC_AUC)
  r13 <- cor(df$Gene, df$BM_Blast)
  r23 <- cor(df$AraC_AUC, df$BM_Blast)
  
  denom <- sqrt((1 - r13^2) * (1 - r23^2))
  if (is.na(denom) || denom == 0) return(c(NA, NA))
  
  partial_r <- (r12 - r13 * r23) / denom
  stat <- partial_r * sqrt((nrow(df) - 3) / (1 - partial_r^2))
  pval <- 2 * pnorm(-abs(stat))
  
  return(c(partial_r, pval))
}

# ========== 4. Apply to all genes ==========
results <- pblapply(seq_len(ncol(exp_log2)), function(i) {
  partial_correlation(exp_log2[, i], arac$AraC_AUC, bm_blast$BM_Blast)
})

pcor_matrix <- do.call(rbind, results)

# ========== 5. Assemble results ==========
results_df <- tibble(
  Gene = colnames(exp_log2),
  PearsonRho = pcor_matrix[, 1],
  P.Value = pmin(pcor_matrix[, 2], 1)
) %>%
  mutate(
    FDR = p.adjust(P.Value, method = "BH")
  )

# Compute Z-score only for non-NA PearsonRho
results_df$Z_score <- NA
valid_idx <- which(!is.na(results_df$PearsonRho))
results_df$Z_score[valid_idx] <- scale(results_df$PearsonRho[valid_idx])[, 1]

# Compute AraCRS 
results_df <- results_df %>%
  mutate(
    AraCRS = ifelse(is.na(Z_score), NA, -log2(FDR + 1e-40) * abs(Z_score))
  )

# ========== 6. Add gene annotation ==========
gene_anno <- exp_filtered[, c("Symbol", "Type")] %>%
  rownames_to_column("Gene")

results_df <- left_join(results_df, gene_anno, by = "Gene")

# View top ranked genes
results_df <- arrange(results_df, desc(AraCRS))

# Save to file
# write_csv(results_df, "../02.data/BeatAML_AraC_AUC_Gene_Corr.csv")

```

## Visualization of Gene-AraC Correlation Results

```{r plot-partial-correlation-results, eval=TRUE, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(ggrepel)
library(readr)
library(ggrastr)

# Load the data
df <- read_csv("../02.data/BeatAML_AraC_AUC_Gene_Corr.csv")

# Define genes to label
genes_to_label <- c(
  "GAS6-AS2", "SLC29A1", "CDA", "PCBP3", "TUFT1", "DCK", "TM4SF19-AS1",
  "RP11-434H6.6", "RAB43", "LAMC1", "RP11-1029J19.5", "DCTD", "UBA6-AS1",
  "CHEK2", "RP11-162D16.2", "LDLRAD2", "FAM166B", "S100A1"
)

# Rank by Z_score
df_ranked <- df %>%
  arrange(desc(Z_score)) %>%
  mutate(
    rank = row_number(),
    abs_z = abs(Z_score),
    is_labeled = Symbol %in% genes_to_label,
    label = ifelse(is_labeled, Symbol, NA)
  )

# Get Z-score cutoffs for top and bottom 1000 genes
top_1000_cutoff <- df_ranked$Z_score[1000]
bottom_1000_cutoff <- df_ranked$Z_score[nrow(df_ranked) - 999]

# Create plot
p <- ggplot(df_ranked, aes(x = rank, y = Z_score)) +
  # Main colored points
  geom_point_rast(aes(size = abs_z, color = Z_score), alpha = 0.8) +

  # Black outline circles (same size) on labeled genes
  geom_point(
    data = df_ranked %>% filter(is_labeled),
    aes(size = abs_z),
    shape = 21, stroke = 0.8, fill = NA, color = "black"
  ) +

  # Text labels
  geom_text_repel(
    data = df_ranked %>% filter(is_labeled),
    aes(label = Symbol),
    size = 3,
    max.overlaps = Inf
  ) +

  # Horizontal dashed lines for top/bottom 1000 gene Z-score cutoffs
  geom_hline(yintercept = c(top_1000_cutoff, bottom_1000_cutoff),
             linetype = "dashed", color = "grey40") +

  # Scales and themes
  scale_color_distiller(palette = "PRGn", direction = -1) +
  scale_size_continuous(range = c(1, 4)) +
  labs(
    x = "Gene Rank (by Z-score)",
    y = "Z-score",
    color = "Z-score",
    size = "|Z-score|"
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

p

# ggsave(p, filename = "../03.out/figure1/01.gene_arac_dotplot_beataml.pdf", width = 6.5, height = 5.4)
```

## Processing the CRISPRa data from the 2018 Cell paper

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(readxl)
library(tidyverse)
library(glue)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(ggrepel)
library(ggrastr)
library(patchwork)

# Define constants
abundance_cols <- c("d0_ctrl1", "d0_ctrl2", "d14_arac1", "d14_arac2")

# General-purpose CRISPRa processing function
process_crispr_sheet <- function(file, sheet, id_col, keytype = NULL, map_to_symbol = TRUE) {
  message(glue("Processing sheet {sheet}..."))

  df <- read_xlsx(file, sheet = sheet)

  # Normalize to CPM
  df <- df %>%
    mutate(across(all_of(abundance_cols), ~ .x / sum(.x, na.rm = TRUE) * 1e6))

  # Map to gene symbols if required
  if (map_to_symbol) {
    df$Symbol <- mapIds(
      org.Hs.eg.db,
      keys = df[[id_col]],
      column = "SYMBOL",
      keytype = keytype,
      multiVals = "first"
    )
    df <- df %>% filter(!is.na(Symbol))
  } else {
    df <- df %>% mutate(Symbol = .data[[id_col]])
  }

  # Aggregate by symbol and calculate CRISPR score (log2 fold-change)
  df_agg <- df %>%
    group_by(Symbol) %>%
    summarise(across(all_of(abundance_cols), ~ mean(.x, na.rm = TRUE)), .groups = "drop") %>%
    mutate(CRISPR_score = log2(
      pmax(rowMeans(across(starts_with("d14")), na.rm = TRUE), 1e-6) /
      pmax(rowMeans(across(starts_with("d0_")), na.rm = TRUE), 1e-6)
    ))

  return(df_agg)
}

# Wrapper function to create dot plots
plot_crispr_dot <- function(df, title) {
  ranked <- df %>%
    arrange(desc(CRISPR_score)) %>%
    mutate(
      rank = row_number(),
      abs_score = abs(CRISPR_score)
    )

  label_genes <- ranked %>%
    slice_head(n = 50) %>%
    bind_rows(slice_tail(ranked, n = 50))

  ggplot(ranked, aes(x = rank, y = CRISPR_score)) +
    geom_point_rast(aes(size = abs_score, color = CRISPR_score), alpha = 0.8) +
    geom_text_repel(
      data = label_genes,
      aes(label = Symbol),
      size = 2,
      max.overlaps = Inf
    ) +
    scale_color_distiller(palette = "BrBG", direction = 1) +
    scale_size_continuous(range = c(1, 4)) +
    labs(
      title = title,
      x = "Gene Rank (by CRISPR Score)",
      y = "CRISPR Score",
      color = "CRISPR Score",
      size = "|CRISPR Score|"
    ) +
    theme_classic(base_size = 14)
}

# Input file path
file_path <- "../02.data/Cell_2018_CRISPRa.xlsx"

# Process protein-coding genes (Sheet 1, RefSeq ID)
crispr_protein_coding <- process_crispr_sheet(
  file = file_path,
  sheet = 1,
  id_col = "Gene",
  keytype = "REFSEQ",
  map_to_symbol = TRUE
)

# Process non-coding genes (Sheet 2, use gene_id as symbol)
crispr_noncoding <- process_crispr_sheet(
  file = file_path,
  sheet = 2,
  id_col = "Gene",
  map_to_symbol = FALSE
)

# Generate plots
p1 <- plot_crispr_dot(crispr_protein_coding, "Ara-C Responsive Protein-Coding Genes")
p2 <- plot_crispr_dot(crispr_noncoding, "Ara-C Responsive Non-Coding Genes")

plots <- p1 + p2

plots

ggsave(plots, filename = "../03.out/figure1/CRISPRa_dotplots.pdf", width = 13, height = 5)

# Optional: save results
# write_csv(crispr_protein_coding, "../02.data/CRISPRa_protein_coding.csv")
# write_csv(crispr_noncoding, "../02.data/CRISPRa_noncoding.csv")
```

## Define the AraCRS (Ara-C Response Score) to predict functional Ara-C modulators 
We defined the top identified genes by CRISPR activation screen as the Ara-C Responsive Genes (Responsive) and the others as the NonResponsive group. Then we compare their AraCRS.

```{r, eval=TRUE, message = FALSE, warning = FALSE}
# Load libraries
library(pROC)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)

# Step 1: Select top 50 sensitive and resistant genes based on CRISPR scores
top_n <- 50

# Sort and select protein-coding genes
sensitive_gene <- crispr_protein_coding %>%
  arrange(CRISPR_score) %>%
  slice_head(n = top_n) %>%
  mutate(type = "Sensitive")

resistant_gene <- crispr_protein_coding %>%
  arrange(desc(CRISPR_score)) %>%
  slice_head(n = top_n) %>%
  mutate(type = "Resistant")

# Sort and select non-coding genes
sensitive_nonc <- crispr_noncoding %>%
  arrange(CRISPR_score) %>%
  slice_head(n = top_n) %>%
  mutate(type = "Sensitive")

resistant_nonc <- crispr_noncoding %>%
  arrange(desc(CRISPR_score)) %>%
  slice_head(n = top_n) %>%
  mutate(type = "Resistant")

# Combine sensitive and resistant genes
arac_responsive_gene <- bind_rows(sensitive_gene, resistant_gene, sensitive_nonc, resistant_nonc) %>%
  dplyr::select(Symbol, type)  # Keep only necessary columns
# write.csv(arac_responsive_gene, "../02.data/arac_responsive_gene_top200.csv")

# Step 2: Load Rank_score matrix and label responsive vs non-responsive genes
gene_corr_mat <- read_csv("../02.data/BeatAML_AraC_AUC_Gene_Corr.csv")

mat_labeled <- gene_corr_mat %>%
  left_join(arac_responsive_gene, by = "Symbol") %>%
  mutate(type = replace_na(type, "NonResponsive"),
         type = if_else(type != "NonResponsive", "Responsive", "NonResponsive"))

# Step 3: Compute ROC

roc_obj <- roc(response = mat_labeled$type,
               predictor = mat_labeled$AraCRS,
               levels = c("NonResponsive", "Responsive"),
               direction = "<") # Higher AraCRS values indicates responsiveness

# Step 4: Compute AUC and 95% CI
auc_val <- auc(roc_obj)
ci_auc <- ci.auc(roc_obj)

# Step 5: Manual one-sample Z-test (test if AUC > 0.5)
n1 <- sum(mat_labeled$type == "Responsive")
n2 <- sum(mat_labeled$type == "NonResponsive")

q1 <- auc_val / (2 - auc_val)
q2 <- (2 * auc_val^2) / (1 + auc_val)
se_auc <- sqrt((auc_val * (1 - auc_val) +
                (n1 - 1) * (q1 - auc_val^2) +
                (n2 - 1) * (q2 - auc_val^2)) / (n1 * n2))

z_score <- (auc_val - 0.5) / se_auc
p_value_auc <- 2 * pnorm(-abs(z_score))  # two-sided

# Format text to display
auc_ci_text <- paste0("AUC = ", round(auc_val, 3),
                      " [", round(ci_auc[1], 3), "-", round(ci_auc[3], 3), "]")
pval_text <- paste("P-value =", formatC(p_value_auc, format = "e", digits = 2))

# Get optimal cutoff (Youden's index)
opt_cut <- coords(roc_obj, "best", ret = "threshold", transpose = FALSE)

# Step 6: Plot ROC curve
roc_plot <- ggroc(roc_obj, linewidth = 1.2, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50", linewidth = 0.8) +
  
  # Annotate cutoff, AUC with CI, and p-value
  annotate("text", x = 0.6, y = 0.3,
           label = paste("Optimal cutoff:", round(opt_cut, 3)),
           size = 5, color = "darkred") +
  
  annotate("text", x = 0.6, y = 0.2,
           label = auc_ci_text,
           size = 5, color = "black") +

  annotate("text", x = 0.6, y = 0.1,
           label = pval_text,
           size = 5, color = "black") +

  labs(
    title = "ROC Curve for AraC Response Prediction",
    x = "1 - Specificity", y = "Sensitivity"
  ) +
  theme_classic(base_size = 14)

roc_plot

gene_corr_mat$group <- ifelse(gene_corr_mat$AraCRS > as.numeric(opt_cut), "Selected", "Non")
table(gene_corr_mat$group)

```

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)

# Set factor levels for consistent ordering
mat_labeled <- mat_labeled %>%
  mutate(type = factor(type, levels = c("NonResponsive", "Responsive")))

# Compute group-wise mean and SEM for AraCRS
summary_df <- mat_labeled %>%
  group_by(type) %>%
  summarise(
    mean = mean(AraCRS, na.rm = TRUE),
    sem = sd(AraCRS, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Perform Wilcoxon rank-sum test
wilcox_p <- wilcox.test(AraCRS ~ type, data = mat_labeled)$p.value
p_label <- paste0("Wilcoxon test, p = ", formatC(wilcox_p, format = "e", digits = 2))

# Generate bar plot with mean ± SEM
p <- ggplot(summary_df, aes(x = type, y = mean, fill = type)) +
  geom_bar(stat = "identity", width = 0.6, color = "black", alpha = 0.85) +
  geom_errorbar(
    aes(ymin = mean - sem, ymax = mean + sem),
    width = 0.3, size = 0.8
  ) +
  annotate(
    "text",
    x = 1.5,
    y = max(summary_df$mean + summary_df$sem) * 1.05,
    label = p_label,
    size = 4,
    fontface = "italic"
  ) +
  scale_fill_manual(values = c("NonResponsive" = "#E69F00", "Responsive" = "#56B4E9")) +
  labs(
    x = "Response Type",
    y = "AraC Response Score (AraCRS)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )

p

ggsave(p, filename = "../03.out/figure1/04.CRISPRa_boxplot.pdf", width = 3, height = 4)
```

### Define Ara-C responsive genes

```{r}
# Define resistance and sensitivity genes
resistance_genes <- df_ranked %>% 
  filter(Z_score >= top_1000_cutoff) %>%
  mutate(category = "AraC_Resistance")

sensitivity_genes <- df_ranked %>%
  filter(Z_score <= bottom_1000_cutoff) %>%
  mutate(category = "AraC_Sensitivity")

# Combine and optionally save or inspect
araC_gene_sets <- bind_rows(resistance_genes, sensitivity_genes)

# write_csv(araC_gene_sets, "../02.data/AraC_TopBottom1000_Genes.csv")
```

## Identification of AraC-Associated Genes Using CCLE Datasets

To extend our analysis across cancer cell lines, we integrated RNA-seq data and Ara-C sensitivity profiles from the Cancer Cell Line Encyclopedia (CCLE). We assessed gene-drug correlations and visualized representative AraC-correlated genes across multiple lineages.

```{r plot-ccle-correlation-landscape, eval=TRUE, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(ggrepel)
library(readr)
library(ggrastr)

# Load datasets
ccle_auc <- read.csv("../02.data/GDSC1_RNA-seq_AraC_AUC.csv")

# Histogram of Ara-C sensitivity
p_hist <- ggplot(ccle_auc, aes(x = Z_SCORE)) +
  geom_histogram(binwidth = 0.3, fill = "#3182bd", color = "black") +
  labs(
    x = "Z-Scaled AUC (Ara-C Sensitivity)",
    y = "Number of Cell Lines",
    title = "Distribution of Ara-C Sensitivity Across CCLE"
  ) +
  theme_minimal()

p_hist
ggsave(p_hist, filename = "../03.out/figure1/gene_arac_hist_ccle.pdf", width = 5.4, height = 5)

# Prepare correlation data for plotting
# Load the data
ccle_df <- read.csv("../02.data/CCLE_CorrOut.csv")

# Define genes to label
genes_to_label <- c(
  "TUFT1", "ZBP1", "SETBP1", "CDA", "MUL1", "CXCL17", "RMDN2-AS1",
  "DCK", "SLC29A1", "CHEK2", "DCTD", "GADD45G", "ATP6V0A1", "CLDN3",
  "RP11-794G24.1", "NPM3", "NAP1L1", "ITPRIP", "RPL14", "LINC01222"
)

# Rank and annotate
ccle_ranked <- ccle_df %>%
  arrange(desc(Z_score)) %>%
  mutate(
    rank = row_number(),
    abs_z = abs(Z_score),
    is_labeled = Symbol %in% genes_to_label,
    label = ifelse(is_labeled, Symbol, NA)
  )

# Get Z-score cutoffs
top1000_z_ccle <- ccle_ranked$Z_score[1000]
tail1000_z_ccle <- ccle_ranked$Z_score[nrow(ccle_ranked) - 999]

# Plot CCLE Z-score ranking
p_ccle <- ggplot(ccle_ranked, aes(x = rank, y = Z_score)) +
  geom_point_rast(aes(size = abs_z, color = Z_score), alpha = 0.8) +
  geom_point(
    data = filter(ccle_ranked, is_labeled),
    aes(size = abs_z), shape = 21, stroke = 0.8, fill = NA, color = "black"
  ) +
  geom_text_repel(
    data = filter(ccle_ranked, is_labeled),
    aes(label = Symbol),
    size = 3, max.overlaps = Inf
  ) +
  geom_hline(yintercept = c(top1000_z_ccle, tail1000_z_ccle),
             linetype = "dashed", color = "grey40") +
  scale_color_distiller(palette = "PRGn", direction = -1) +
  scale_size_continuous(range = c(1, 4)) +
  labs(
    title = "Ara-C-Responsive Genes in CCLE Cell Lines (In vitro)",
    x = "Gene Rank (by Z-score)",
    y = "Z-score",
    color = "Z-score",
    size = "|Z-score|"
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

p_ccle

ggsave(p_ccle, filename = "../03.out/figure1/gene_arac_dotplot_ccle.pdf", width = 6.5, height = 5.4)
```

## Top genes overlap

```{r}
library(ggvenn)
library(tidyverse)

patient_df <- read_csv("../02.data/BeatAML_AraC_AUC_Gene_Corr.csv")
ccle_df <- read.csv("../02.data/CCLE_CorrOut.csv")
ccle_df <- ccle_df[!duplicated(ccle_df$Symbol), ]
# Rank both
top_bottom_genes <- function(df) {
  df <- df %>% arrange(desc(Z_score))
  top1000 <- df$Symbol[1:1000]
  bottom1000 <- df$Symbol[(nrow(df) - 999):nrow(df)]
  unique(na.omit(c(top1000, bottom1000)))
}

genes_patient <- top_bottom_genes(patient_df)
genes_ccle <- top_bottom_genes(ccle_df)

gene_sets <- list(
  Patient = genes_patient,
  CCLE = genes_ccle
)

# Create Venn plot with ggvenn
p_venn <- ggvenn(
  gene_sets,
  fill_color = c("#E69F00", "#56B4E9"),
  stroke_size = 0.5,
  set_name_size = 5,
  text_size = 4
)
p_venn
# Define gene universe = genes that exist in both datasets
gene_universe <- intersect(patient_df$Symbol, ccle_df$Symbol)

# Logical vectors for membership
in_patient <- gene_universe %in% genes_patient
in_ccle <- gene_universe %in% genes_ccle

# Contingency table and p-value
contingency <- table(Patient = in_patient, CCLE = in_ccle)
fisher_p <- fisher.test(contingency)$p.value
p_label <- paste0("Fisher's Exact Test\np = ", formatC(fisher_p, format = "e", digits = 2))
p_label

ggsave(p_venn, filename = "../03.out/figure1/arac_genes_overlap.pdf", width = 5, height = 5)

```

## Ex vivo screens in the Functional Precision Medicine Tumor Board AML

```{r fpmtb_normalize_tpm, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)

# -------------------------------
# Load data
# -------------------------------
count_df <- read_csv("../02.data/File_8_RNA_seq_Raw_Reads_163S_4Healthy.csv") %>%
  rename(id2 = 1)

gene_length_df <- read_csv("../02.data/gencode.v22_geneLength.csv")

# -------------------------------
# Merge gene length with expression counts
# -------------------------------
merged_df <- inner_join(gene_length_df, count_df, by = "id2")

# Extract count matrix and corresponding gene lengths
count_mat <- merged_df %>%
  column_to_rownames("id2") %>%
  select(-(1:4))

length_vec <- merged_df %>%
  select(id2, length) %>%
  column_to_rownames("id2") %>%
  pull(length)

# -------------------------------
# Calculate TPM
# -------------------------------
# Convert to matrix for vectorized math
count_mat_numeric <- as.matrix(count_mat)

# TPM calculation: Normalize for gene length and library size
TPM <- t(t(count_mat_numeric / length_vec) * 1e6 / colSums(count_mat_numeric / length_vec))

# Convert TPM matrix back to tibble
TPM_df <- as_tibble(TPM, rownames = "id2") %>%
  left_join(select(gene_length_df, id2, Symbol, Type), by = "id2") %>%
  relocate(Symbol, Type)

# -------------------------------
# Load drug profiling data
# -------------------------------
drug_df <- read_csv("../02.data/File_3_Drug_response_sDSS_164S_17Healthy.csv")

# -------------------------------
# Find shared samples and filter data
# -------------------------------
shared_samples <- intersect(colnames(TPM_df), colnames(drug_df))

tpm_filtered <- TPM_df %>%
  select(id2, Symbol, Type, all_of(shared_samples))

count_filtered <- count_mat_numeric[, shared_samples]

drug_filtered <- drug_df %>%
  select(Drug_name, all_of(shared_samples))

# -------------------------------
# Extract 7+3 chemotherapy drugs
# -------------------------------
chemo_drugs <- c("Cytarabine", "Daunorubicin", "Idarubicin")

chemo_df <- drug_filtered %>%
  filter(Drug_name %in% chemo_drugs) %>%
  column_to_rownames("Drug_name")

# -------------------------------
# Write outputs
# -------------------------------
# write.csv(count_filtered, "../02.data/fpmtb_drug_gene_count.csv")
# write.csv(tpm_filtered, "../02.data/fpmtb_drug_gene_tpm.csv")
# write.csv(chemo_df, "../02.data/fpmtb_chemo_drug_sensitivity.csv")

```

### Unsupervised clustering for "Ara-C + Idarubicin" (2 major groups)

```{r}
# -------------------------------
# Load required libraries
# -------------------------------
library(pheatmap)
library(ggpubr)
library(dplyr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)
library(ggbeeswarm)

# -------------------------------
# Load and subset drug sensitivity matrix
# -------------------------------
core_mat <- read.csv("../02.data/fpmtb_chemo_drug_sensitivity.csv", row.names = 1)
selected_drugs <- c("Cytarabine", "Idarubicin")
core_mat_sub <- core_mat[selected_drugs, ]

# Convert to matrix and scale
core_mat_scaled <- as.matrix(core_mat_sub)

# Define color scale
heat_colors <- colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100)

# -------------------------------
# First heatmap clustering to assign labels
# -------------------------------
heat_result <- pheatmap(
  core_mat_scaled,
  color = heat_colors,
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  scale = "row",
  cutree_cols = 2,
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  border_color = "black",
  clustering_method = "ward.D"
)

# -------------------------------
# Extract clustering and assign labels
# -------------------------------
# Assign cluster from pheatmap
sample_cluster <- data.frame(
  Sample = colnames(core_mat_scaled),
  Cluster = cutree(heat_result$tree_col, k = 2)
)

# Calculate mean sDSS per sample (Cytarabine + Idarubicin)
sample_cluster$Mean_sDSS <- colMeans(core_mat_scaled)

# Determine which cluster is more sensitive (higher sDSS)
cluster_avg <- sample_cluster %>%
  group_by(Cluster) %>%
  summarize(Avg_sDSS = mean(Mean_sDSS)) %>%
  arrange(desc(Avg_sDSS)) %>%
  mutate(Label = c("Sensitive", "Resistant"))

# Merge back to assign labels
sample_cluster <- sample_cluster %>%
  left_join(cluster_avg, by = "Cluster") %>%
  dplyr::select(Sample, Cluster, Label) %>%
  column_to_rownames("Sample")

# Prepare for heatmap annotation
annotation_col <- data.frame(Label = factor(sample_cluster$Label))
rownames(annotation_col) <- rownames(sample_cluster)

ann_colors <- list(
  Label = c("Sensitive" = "#e9b383", "Resistant" = "#7c9d97")
)

# -------------------------------
# Final annotated heatmap
# -------------------------------
pdf("../03.out/figure1/fpmtb_exVivo_arac_ida_heatmap.pdf", width = 10, height = 4)
pheatmap(
  core_mat_scaled,
  color = heat_colors,
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  cutree_cols = 2,
  scale = "row",
  border_color = "black",
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  annotation_col = annotation_col,
  annotation_colors = ann_colors,
  clustering_method = "ward.D"
)
dev.off()

# -------------------------------
# Boxplot of drug sensitivity by cluster
# -------------------------------
# Prepare long-format data
drug_df <- as.data.frame(t(core_mat_sub)) %>%
  rownames_to_column("Sample") %>%
  left_join(sample_cluster %>% rownames_to_column("Sample"), by = "Sample") %>%
  pivot_longer(
    cols = all_of(selected_drugs),
    names_to = "Drug",
    values_to = "sDSS"
  )

# Reorder drug factor levels
drug_df$Drug <- factor(drug_df$Drug, levels = selected_drugs)

# Plot
boxplot_fig <- ggplot(drug_df, aes(x = Label, y = sDSS, color = Label)) +
  geom_boxplot_jitter(width = 0.5, alpha = 0.8, outlier.shape = NA) +
  geom_quasirandom(width = 0.15, alpha = 0.75) +
  facet_wrap(~Drug, ncol = 2) +
  theme_pubr() +
  labs(x = NULL, y = "Drug sensitivity (sDSS)") +
  stat_compare_means(
    label = "p.signif",
    comparisons = list(c("Resistant", "Sensitive")),
    method = "wilcox.test"
  ) +
  scale_color_manual(values = c("Sensitive" = "#e9b383", "Resistant" = "#7c9d97")) +
  theme(
    strip.text.x = element_text(size = 13),
    legend.position = "none"
  )

boxplot_fig

ggsave(boxplot_fig, filename =  "../03.out/figure1/fpmtb_exVivo_arac_ida_boxplot.pdf", width = 5, height = 2)
# -------------------------------
## Exact the data for Defferential express analysis
# -------------------------------

# library(DESeq2)
# library(readxl)
# sample_cluster <- sample_cluster[order(sample_cluster$Label, decreasing = TRUE), ]
# 
# table(sample_cluster$Label)
# 
# count <- read.csv("../02.data/fpmtb_drug_gene_count.csv", header = T, row.names = 1)
# tpm <- read.csv("../02.data/fpmtb_drug_gene_tpm.csv", header = T, row.names = 1)
# colnames(tpm)[1] <- "ID"
# 
# tpm <- subset(tpm, Type %in% c("protein_coding", "miRNA", "lincRNA", "antisense",
#                                "processed_transcript",
#                                "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene",
#                                "transcribed_unprocessed_pseudogene","macro_lncRNA", "TEC",
#                                "3prime_overlapping_ncRNA","sense_intronic", "sense_overlapping",
#                                "unprocessed_pseudogene"))
# tpm <- tpm[, c("ID", "Symbol", "Type", as.character(row.names(sample_cluster)))]
# 
# # filter low-express genes
# tpm <- tpm[rowMeans(tpm[, 4:ncol(tpm)]) >= 0.05, ]
# 
# count <- count[tpm$ID, as.character(row.names(sample_cluster))]
# 
# ## run DESEQ2
# source("../01.script/RunDESeq2.R")
# 
# RunDESeq2(count_mat = count, n.cont = 67, n.treat = 65,
#           prefix = "../02.data/fpmtb_AraC_Ida_DESeq2_out_R_S", sort.p = FALSE,
#           merge.normalized = TRUE,
#           normalized_mat = tpm)

```

### Unsupervised clustering for "Ara-C + Daunorubicin" (4 clusters)

```{r}
# -------------------------------
# Load required libraries
# -------------------------------
library(pheatmap)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)
library(ggbeeswarm)

# -------------------------------
# Load and subset drug sensitivity matrix
# -------------------------------
core_mat <- read.csv("../02.data/fpmtb_chemo_drug_sensitivity.csv", row.names = 1)
selected_drugs <- c("Cytarabine", "Idarubicin")
core_mat_sub <- core_mat[selected_drugs, ]
core_mat_scaled <- scale(t(core_mat_sub)) %>% t()  # scale row-wise

# -------------------------------
# Clustering: 4 subclusters (2 sensitive, 2 resistant)
# -------------------------------
heat_result <- pheatmap(
  core_mat_scaled,
  color = colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100),
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  scale = "none",
  cutree_cols = 4,
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  border_color = "black",
  clustering_method = "ward.D"
)

# Extract cluster assignments and compute mean sensitivity
sample_cluster <- data.frame(
  Sample = colnames(core_mat_scaled),
  Cluster = cutree(heat_result$tree_col, k = 4)
)
sample_cluster$Mean_sDSS <- colMeans(core_mat_scaled)

# Rank clusters by sensitivity (higher sDSS = more sensitive)
cluster_labels <- sample_cluster %>%
  group_by(Cluster) %>%
  summarise(Avg_sDSS = mean(Mean_sDSS)) %>%
  arrange(desc(Avg_sDSS)) %>%
  mutate(Label = c("Sensitive-1", "Sensitive-2", "Resistant-1", "Resistant-2"))

# Merge cluster label assignments
sample_cluster <- sample_cluster %>%
  left_join(cluster_labels, by = "Cluster") %>%
  select(Sample, Cluster, Label) %>%
  column_to_rownames("Sample")

# -------------------------------
# Prepare annotation for heatmap
# -------------------------------
annotation_col <- data.frame(
  Label = factor(sample_cluster$Label,
                 levels = c("Sensitive-1", "Sensitive-2", "Resistant-1", "Resistant-2"))
)
rownames(annotation_col) <- rownames(sample_cluster)

ann_colors <- list(
  Label = c(
    "Sensitive-1" = "#e9b383", "Sensitive-2" = "#e9b383",
    "Resistant-1" = "#7c9d97", "Resistant-2" = "#7c9d97"
  )
)

# -------------------------------
# Final annotated heatmap
# -------------------------------
 pdf("../03.out/figure1/02.exVivo_arac_ida_heatmap_fpmtb.pdf", width = 10, height = 2.5)
pheatmap(
  core_mat_scaled,
  color = colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100),
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  cutree_cols = 4,
  scale = "none",
  border_color = "black",
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  annotation_col = annotation_col,
  annotation_colors = ann_colors,
  clustering_method = "ward.D"
)
 dev.off()

# -------------------------------
# Boxplot using scaled matrix
# -------------------------------

# Prepare data 

drug_df <- core_mat_scaled %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  left_join(sample_cluster %>% rownames_to_column("Sample"), by = "Sample") %>%
  pivot_longer(
    cols = all_of(selected_drugs),
    names_to = "Drug", values_to = "sDSS_scaled"
  )

drug_df$Drug <- factor(drug_df$Drug, levels = selected_drugs)
drug_df$Label <- factor(drug_df$Label, levels = c("Resistant-1", "Resistant-2", "Sensitive-1", "Sensitive-2"))

# -------------------------------
# Plot: Compare Cytarabine vs Idarubicin within each cluster
# -------------------------------
drug_comparisons <- list(c("Cytarabine", "Idarubicin"))

boxplot_fig <- ggplot(drug_df, aes(x = Drug, y = sDSS_scaled, fill = Drug)) +
  geom_boxplot(width = 0.5, alpha = 0.8, outlier.shape = NA) +
  geom_quasirandom(width = 0.2, alpha = 0.75, color = "black") +
  facet_wrap(~Label, nrow = 1) +
  theme_pubr() +
  labs(x = NULL, y = "Scaled drug sensitivity (sDSS)") +
  stat_compare_means(
    label = "p.signif",
    method = "wilcox.test",
    comparisons = drug_comparisons
  ) +
  scale_fill_manual(values = c("Cytarabine" = "#e9b383", "Idarubicin" = "#7c9d97")) +
  theme(
    strip.text.x = element_text(size = 13),
    legend.position = "none"
  )

boxplot_fig
ggsave(boxplot_fig, filename = "../03.out/figure1/03.exVivo_arac_ida_boxplot_fpmtb.pdf", width = 8, height = 3)

```

### Unsupervised clustering for "Ara-C + Daunorubicin" (2 major groups)

```{r}
# -------------------------------
# Load required libraries
# -------------------------------
library(pheatmap)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)
library(ggbeeswarm)

# -------------------------------
# Load and subset drug sensitivity matrix
# -------------------------------
core_mat <- read.csv("../02.data/fpmtb_chemo_drug_sensitivity.csv", row.names = 1)
selected_drugs <- c("Cytarabine", "Daunorubicin")
core_mat_sub <- core_mat[selected_drugs, ]

# Convert to matrix and scale
core_mat_scaled <- as.matrix(core_mat_sub)

# Define color scale
heat_colors <- colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100)

# -------------------------------
# First heatmap clustering to assign labels
# -------------------------------
heat_result <- pheatmap(
  core_mat_scaled,
  color = heat_colors,
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  scale = "row",
  cutree_cols = 2,
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  border_color = "black",
  clustering_method = "ward.D"
)

# -------------------------------
# Extract clustering and assign labels
# -------------------------------
# Assign cluster from pheatmap
sample_cluster <- data.frame(
  Sample = colnames(core_mat_scaled),
  Cluster = cutree(heat_result$tree_col, k = 2)
)

# Calculate mean sDSS per sample (Cytarabine + Idarubicin)
sample_cluster$Mean_sDSS <- colMeans(core_mat_scaled)

# Determine which cluster is more sensitive (higher sDSS)
cluster_avg <- sample_cluster %>%
  group_by(Cluster) %>%
  summarize(Avg_sDSS = mean(Mean_sDSS)) %>%
  arrange(desc(Avg_sDSS)) %>%
  mutate(Label = c("Sensitive", "Resistant"))

# Merge back to assign labels
sample_cluster <- sample_cluster %>%
  left_join(cluster_avg, by = "Cluster") %>%
  dplyr::select(Sample, Cluster, Label) %>%
  column_to_rownames("Sample")

# Prepare for heatmap annotation
annotation_col <- data.frame(Label = factor(sample_cluster$Label))
rownames(annotation_col) <- rownames(sample_cluster)

ann_colors <- list(
  Label = c("Sensitive" = "#e9b383", "Resistant" = "#7c9d97")
)

# -------------------------------
# Final annotated heatmap
# -------------------------------
pdf("../03.out/figure1/fpmtb_exVivo_arac_dau_heatmap.pdf", width = 10, height = 4)
pheatmap(
  core_mat_scaled,
  color = heat_colors,
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  cutree_cols = 2,
  scale = "row",
  border_color = "black",
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  annotation_col = annotation_col,
  annotation_colors = ann_colors,
  clustering_method = "ward.D"
)
dev.off()

# -------------------------------
# Boxplot of drug sensitivity by cluster
# -------------------------------
# Prepare long-format data
drug_df <- as.data.frame(t(core_mat_sub)) %>%
  rownames_to_column("Sample") %>%
  left_join(sample_cluster %>% rownames_to_column("Sample"), by = "Sample") %>%
  pivot_longer(
    cols = all_of(selected_drugs),
    names_to = "Drug",
    values_to = "sDSS"
  )

# Reorder drug factor levels
drug_df$Drug <- factor(drug_df$Drug, levels = selected_drugs)

# Plot
boxplot_fig <- ggplot(drug_df, aes(x = Label, y = sDSS, color = Label)) +
  geom_boxplot_jitter(width = 0.5, alpha = 0.8, outlier.shape = NA) +
  geom_quasirandom(width = 0.15, alpha = 0.75) +
  facet_wrap(~Drug, ncol = 2) +
  theme_pubr() +
  labs(x = NULL, y = "Drug sensitivity (sDSS)") +
  stat_compare_means(
    label = "p.signif",
    comparisons = list(c("Resistant", "Sensitive")),
    method = "wilcox.test"
  ) +
  scale_color_manual(values = c("Sensitive" = "#e9b383", "Resistant" = "#7c9d97")) +
  theme(
    strip.text.x = element_text(size = 13),
    legend.position = "none"
  )

boxplot_fig
ggsave(boxplot_fig, filename =  "../03.out/figure1/fpmtb_exVivo_arac_dau_boxplot.pdf", width = 5, height = 2)
# -------------------------------
## Exact the data for Defferential express analysis
# -------------------------------

# library(DESeq2)
# library(readxl)
# sample_cluster <- sample_cluster[order(sample_cluster$Label, decreasing = TRUE), ]
# 
# table(sample_cluster$Label)
# 
# count <- read.csv("../02.data/fpmtb_drug_gene_count.csv", header = T, row.names = 1)
# tpm <- read.csv("../02.data/fpmtb_drug_gene_tpm.csv", header = T, row.names = 1)
# colnames(tpm)[1] <- "ID"
# 
# tpm <- subset(tpm, Type %in% c("protein_coding", "miRNA", "lincRNA", "antisense",
#                                "processed_transcript",
#                                "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene",
#                                "transcribed_unprocessed_pseudogene","macro_lncRNA", "TEC",
#                                "3prime_overlapping_ncRNA","sense_intronic", "sense_overlapping",
#                                "unprocessed_pseudogene"))
# tpm <- tpm[, c("ID", "Symbol", "Type", as.character(row.names(sample_cluster)))]
# 
# # filter low-express genes
# tpm <- tpm[rowMeans(tpm[, 4:ncol(tpm)]) >= 0.05, ]
# 
# count <- count[tpm$ID, as.character(row.names(sample_cluster))]
# 
# ## run DESEQ2
# source("../01.script/RunDESeq2.R")
# 
# RunDESeq2(count_mat = count, n.cont = 51, n.treat = 81,
#           prefix = "../02.data/fpmtb_AraC_Dau_DESeq2_out_R_S", sort.p = FALSE,
#           merge.normalized = TRUE,
#           normalized_mat = tpm)

```

### Unsupervised clustering for "Ara-C + Daunorubicin" (4 clusters)

```{r}
# -------------------------------
# Load required libraries
# -------------------------------
library(pheatmap)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)
library(ggbeeswarm)

# -------------------------------
# Load and subset drug sensitivity matrix
# -------------------------------
core_mat <- read.csv("../02.data/fpmtb_chemo_drug_sensitivity.csv", row.names = 1)
selected_drugs <- c("Cytarabine", "Daunorubicin")
core_mat_sub <- core_mat[selected_drugs, ]
core_mat_scaled <- scale(t(core_mat_sub)) %>% t()  # scale row-wise

# -------------------------------
# Clustering: 4 subclusters (2 sensitive, 2 resistant)
# -------------------------------
heat_result <- pheatmap(
  core_mat_scaled,
  color = colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100),
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  scale = "none",
  cutree_cols = 4,
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  border_color = "black",
  clustering_method = "ward.D"
)

# Extract cluster assignments and compute mean sensitivity
sample_cluster <- data.frame(
  Sample = colnames(core_mat_scaled),
  Cluster = cutree(heat_result$tree_col, k = 4)
)
sample_cluster$Mean_sDSS <- colMeans(core_mat_scaled)

# Rank clusters by sensitivity (higher sDSS = more sensitive)
cluster_labels <- sample_cluster %>%
  group_by(Cluster) %>%
  summarise(Avg_sDSS = mean(Mean_sDSS)) %>%
  arrange(desc(Avg_sDSS)) %>%
  mutate(Label = c("Sensitive-1", "Sensitive-2", "Resistant-1", "Resistant-2"))

# Merge cluster label assignments
sample_cluster <- sample_cluster %>%
  left_join(cluster_labels, by = "Cluster") %>%
  dplyr::select(Sample, Cluster, Label) %>%
  column_to_rownames("Sample")

# -------------------------------
# Prepare annotation for heatmap
# -------------------------------
annotation_col <- data.frame(
  Label = factor(sample_cluster$Label,
                 levels = c("Sensitive-1", "Sensitive-2", "Resistant-1", "Resistant-2"))
)
rownames(annotation_col) <- rownames(sample_cluster)

ann_colors <- list(
  Label = c(
    "Sensitive-1" = "#e9b383", "Sensitive-2" = "#e9b383",
    "Resistant-1" = "#7c9d97", "Resistant-2" = "#7c9d97"
  )
)

# -------------------------------
# Final annotated heatmap
# -------------------------------
pdf("../03.out/figure1/02.exVivo_arac_dau_heatmap_fpmtb.pdf", width = 10, height = 2.5)
pheatmap(
  core_mat_scaled,
  color = colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100),
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  cutree_cols = 4,
  scale = "none",
  border_color = "black",
  show_colnames = FALSE,
  show_rownames = TRUE,
  angle_col = 90,
  annotation_col = annotation_col,
  annotation_colors = ann_colors,
  clustering_method = "ward.D"
)
dev.off()

# -------------------------------
# Boxplot using scaled matrix
# -------------------------------

# Prepare data 

drug_df <- core_mat_scaled %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  left_join(sample_cluster %>% rownames_to_column("Sample"), by = "Sample") %>%
  pivot_longer(
    cols = all_of(selected_drugs),
    names_to = "Drug", values_to = "sDSS_scaled"
  )

drug_df$Drug <- factor(drug_df$Drug, levels = selected_drugs)
drug_df$Label <- factor(drug_df$Label, levels = c("Sensitive-1", "Sensitive-2", "Resistant-1", "Resistant-2"))

# -------------------------------
# Plot: Compare Cytarabine vs Idarubicin within each cluster
# -------------------------------
drug_comparisons <- list(c("Cytarabine", "Daunorubicin"))

boxplot_fig <- ggplot(drug_df, aes(x = Drug, y = sDSS_scaled, fill = Drug)) +
  geom_boxplot(width = 0.5, alpha = 0.8, outlier.shape = NA) +
  geom_quasirandom(width = 0.2, alpha = 0.75, color = "black") +
  facet_wrap(~Label, nrow = 1) +
  theme_pubr() +
  labs(x = NULL, y = "Scaled drug sensitivity (sDSS)") +
  stat_compare_means(
    label = "p.signif",
    method = "wilcox.test",
    comparisons = drug_comparisons
  ) +
  scale_fill_manual(values = c("Cytarabine" = "#e9b383", "Daunorubicin" = "#7c9d97")) +
  theme(
    strip.text.x = element_text(size = 13),
    legend.position = "none"
  )

boxplot_fig

ggsave(boxplot_fig, filename = "../03.out/figure1/03.exVivo_arac_dau_boxplot_fpmtb.pdf", width = 8, height = 3)

```

## Pathway Enrichment via GSEA Using AraC-Correlated Gene List

```{r gsea-analysis, eval=TRUE, message = FALSE, warning = FALSE}
library(clusterProfiler)
library(tidyverse)
library(readr)

# Load KEGG gene sets
geneSet <- read.gmt("../02.data/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

# Define GSEA wrapper function with saving capability
run_gsea <- function(file_path, symbol_col, score_col, tag, save_dir = "../02.data/GSEA_outputs") {
  # Read file
  df <- if (grepl("\\.csv$", file_path)) read_csv(file_path) else read.table(file_path, header = TRUE)
  
  # Prepare ranked gene list
  gene_df <- df %>%
    select(ID = all_of(symbol_col), Score = all_of(score_col)) %>%
    distinct(ID, .keep_all = TRUE) %>%
    drop_na() %>%
    mutate(rank = rank(Score, ties.method = "random"))
  
  geneList <- gene_df$Score
  names(geneList) <- gene_df$ID
  geneList <- sort(geneList, decreasing = TRUE)
  
  # Run GSEA
  set.seed(123456)
  gsea_result <- GSEA(
    geneList,
    TERM2GENE = geneSet,
    pvalueCutoff = 1,
    nPermSimple = 100000,
    maxGSSize = 800,
    pAdjustMethod = "BH"
  )
  
  # Save full gseaResult object to file for plotting later
  if (!dir.exists(save_dir)) dir.create(save_dir)
  saveRDS(gsea_result, file = file.path(save_dir, paste0("GSEA_result_", tag, ".rds")))
  
  # Return formatted results for merging
  gsea_df <- gsea_result@result %>%
    as_tibble() %>%
    mutate(source = tag)
  
  return(gsea_df)
}

# Run and save for each dataset
# gsea_ccle <- run_gsea("../02.data/CCLE_CorrOut.csv", "Symbol", "Z_score", "CCLE_inVitro")
gsea_patient <- run_gsea("../02.data/BeatAML_AraC_AUC_Gene_Corr.csv", "Symbol", "Z_score", "Patient_exVivo")
gsea_ida <- run_gsea("../02.data/fpmtb_AraC_Ida_DESeq2_out_R_S_with_normalized_mat.txt", "Symbol", "log2FoldChange", "AraC_Ida")
gsea_dau <- run_gsea("../02.data/fpmtb_AraC_Dau_DESeq2_out_R_S_with_normalized_mat.txt", "Symbol", "log2FoldChange", "AraC_Dau")

# Combine all results
gsea_all <- bind_rows(gsea_patient, gsea_ida, gsea_dau) # gsea_ccle, 
gsea_all$Description <- gsub("KEGG_", "", gsea_all$Description)
# write_csv(gsea_all, "../02.data/gsea_all_sources.csv")

```

```{r}
library(ggplot2)
library(dplyr)
library(forcats)
library(readr)

# Load data
gsea_all <- read_csv("../02.data/gsea_all_sources.csv")

# Set factor levels for consistency
gsea_all <- gsea_all %>%
  mutate(
    source = factor(source, levels = c("Patient_exVivo", "AraC_Ida", "AraC_Dau", "CCLE_inVitro")),
    sig_label = ifelse(pvalue < 0.05, "*", "")
  )

# Define selected KEGG pathways
selected_pathways <- c(
  "JAK_STAT_SIGNALING_PATHWAY", "MAPK_SIGNALING_PATHWAY", "CHEMOKINE_SIGNALING_PATHWAY",
  "CYTOKINE_CYTOKINE_RECEPTOR_INTERACTION", "TGF_BETA_SIGNALING_PATHWAY",
  "FOCAL_ADHESION", "ADHERENS_JUNCTION", "REGULATION_OF_ACTIN_CYTOSKELETON",
  "NOD_LIKE_RECEPTOR_SIGNALING_PATHWAY", "HEMATOPOIETIC_CELL_LINEAGE",
  "TOLL_LIKE_RECEPTOR_SIGNALING_PATHWAY", "ANTIGEN_PROCESSING_AND_PRESENTATION",
  "FC_GAMMA_R_MEDIATED_PHAGOCYTOSIS", "VEGF_SIGNALING_PATHWAY",
  "ACUTE_MYELOID_LEUKEMIA", "APOPTOSIS", "RIBOSOME", "GLYCEROLIPID_METABOLISM",
  "ASCORBATE_AND_ALDARATE_METABOLISM", "PYRIMIDINE_METABOLISM", "PURINE_METABOLISM",
  "RNA_DEGRADATION", "DNA_REPLICATION", "PYRUVATE_METABOLISM",
  "FATTY_ACID_METABOLISM", "OXIDATIVE_PHOSPHORYLATION"
)

# Subset and order pathways
gsea_selected <- gsea_all %>%
  filter(Description %in% selected_pathways) %>%
  mutate(Description = factor(Description, levels = rev(selected_pathways)))

# Plot
p <- ggplot(gsea_selected, aes(x = Description, y = source)) +
  geom_point(aes(color = NES, size = -log10(p.adjust)), alpha = 0.9) +
  geom_text(
    data = filter(gsea_selected, pvalue < 0.05),
    aes(label = sig_label),
    size = 4, fontface = "bold", color = "black"
  ) +
  scale_color_gradient2(
    low = "#4575b4", mid = "white", high = "#d73027", midpoint = 0,
    name = "NES"
  ) +
  scale_size_continuous(range = c(2, 6), name = expression(-log[10](FDR))) +
  labs(x = "", y = "") +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    legend.key.height = unit(0.6, "cm")
  )

p

ggsave(p, filename = "../03.out/figure1/06.gsea_dot.pdf", width = 10, height = 5.8)

```

## Resistant vs. Sensitive (Valcano plot)

### Ara-C + Ida

```{r}
library(clusterProfiler)
library(enrichplot)
library(dplyr)

degene_mat <- read.table("../02.data/fpmtb_AraC_Ida_DESeq2_out_R_S_with_normalized_mat.txt", header = TRUE)

# Prepare ranked gene list for GSEA
DE_list <- degene_mat %>%
  dplyr::select(ID = 9, Corr = 3) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na() %>%
  mutate(rank = rank(Corr, ties.method = "random"))

geneList <- DE_list$Corr
names(geneList) <- DE_list$ID
geneList <- sort(geneList, decreasing = TRUE)

# Load AraC gene sets
araC_gene_sets <- read_csv("../02.data/AraC_TopBottom1000_Genes.csv")
AraC_GeneSets_df <- araC_gene_sets %>%
  dplyr::select(term = category, gene = Symbol) %>%
  distinct()  

library(fgsea)

# Convert TERM2GENE to a list
gene_sets_list <- split(AraC_GeneSets_df$gene, AraC_GeneSets_df$term)

# Run standard fgsea (instead of GSEA from clusterProfiler)
fgsea.res <- fgsea(
  pathways = gene_sets_list,
  stats = geneList,
  minSize = 1,
  maxSize = 1200,
  nperm = 100000
)

fgsea.res %>%
  arrange(padj)

fgsea.res$ES

# Plot enrichment for AraC_Resistance
pdf("../03.out/figure1/AraC_Resistance_gsea.pdf", width = 5, height = 2)
plotEnrichment(gene_sets_list[["AraC_Resistance"]], geneList) +
  ggtitle("Enrichment of AraC_Resistance Genes")
dev.off()

# Plot enrichment for AraC_Sensitivity
pdf("../03.out/figure1/AraC_Sensitivity_gsea.pdf", width = 5, height = 2)
plotEnrichment(gene_sets_list[["AraC_Sensitivity"]], geneList) +
  ggtitle("Enrichment of AraC_Sensitivity Genes")
dev.off()

```

```{r}
# Load necessary libraries
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(ggrastr)

# Load data
diff <- read.table("../02.data/fpmtb_AraC_Ida_DESeq2_out_R_S_with_normalized_mat.txt", header = TRUE)

# Cap log2FoldChange to [-4, 4] and padj to [1e-6, 1]
diff <- diff %>%
  mutate(
    log2FoldChange = case_when(
      log2FoldChange > 4 ~ 4,
      log2FoldChange < -4 ~ -4,
      TRUE ~ log2FoldChange
    ),
    padj = case_when(
      padj < 1e-6 ~ 1e-6,
      TRUE ~ padj
    )
  )

# Classify genes based on log2FoldChange and adjusted p-value
diff <- diff %>%
  mutate(g = case_when(
    log2FoldChange > 0.5 & padj < 0.05 ~ "up",
    log2FoldChange < -0.5 & padj < 0.05 ~ "down",
    TRUE ~ "stable"
  ))

# Extract significantly upregulated and downregulated genes
diff_up <- diff %>% filter(g == "up")
diff_down <- diff %>% filter(g == "down")

# Load AraC gene sets
araC_gene_sets <- read_csv("../02.data/AraC_TopBottom1000_Genes.csv")

# Extract AraC_Resistance and AraC_Sensitivity gene symbols
AraC_Resistance <- araC_gene_sets %>% filter(category == "AraC_Resistance") %>% pull(Symbol)
AraC_Sensitivity <- araC_gene_sets %>% filter(category == "AraC_Sensitivity") %>% pull(Symbol)

# Classify genes in the differential expression dataset

diff <- diff %>%
  mutate(gene_col = case_when(
    Symbol %in% AraC_Resistance ~ "AraC_Resistance",
    Symbol %in% AraC_Sensitivity ~ "AraC_Sensitivity",
    TRUE ~ "all"
  ))

# Prepare data subsets
data_bg <- diff %>% filter(gene_col == "all")
data_geneset <- diff %>% filter(grepl("AraC_", gene_col))
data_AraC_Resistance <- diff %>% filter(gene_col == "AraC_Resistance")
data_AraC_Sensitivity <- diff %>% filter(gene_col == "AraC_Sensitivity")

# Define the base plot
p <- ggplot(diff, aes(x = log2FoldChange, y = -log10(padj))) +
  # Add shaded regions for significant up- and down-regulated genes
  annotate(
    geom = "rect", fill = "#ea838e", alpha = 0.3,
    xmin = min(diff_up$log2FoldChange), xmax = max(diff_up$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  annotate(
    geom = "rect", fill = "#827fd6", alpha = 0.3,
    xmin = min(diff_down$log2FoldChange), xmax = max(diff_down$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  # Plot background points
  geom_point_rast(
    data = data_bg, shape = 21, color = "black", alpha = 0.2,
    size = 1.3, stroke = 0.6
  ) +
  # Plot gene set and key points
  geom_point_rast(
    data = data_geneset,
    aes(fill = gene_col), shape = 21, color = "black",
    size = 2, stroke = 0.8
  ) +
  # Custom fill colors for gene sets and key genes
  scale_fill_manual(
    values = c(
      "AraC_Resistance" = "#faa400",
      "AraC_Sensitivity" = "#bebebe"
    )
  ) +
  # Add reference lines
  geom_vline(xintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(
    yintercept = -log10(0.05), color = "#ff7d82",
    linewidth = 0.6, linetype = "dotted"
  ) +
  # Customize x-axis
  scale_x_continuous(limits = c(-4, 4), breaks = seq(-4, 4, 1)) +
  # Customize y-axis
  scale_y_continuous(limits = c(-2, 7), breaks = seq(0, 6.5, 2), expand = c(0, 0)) +
  # Apply classic theme and customize
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  ) +
  # Customize axis labels
  xlab("Resistant vs Sensitive (log2FC)") +
  ylab("Significance (-log10 adj.p-val.)")

# Add lineranges and annotations to the plot (p1)
p1 <- p +
  # Add lineranges for AraC_Resistance genes
  geom_linerange(
    data = data_AraC_Resistance,
    aes(x = log2FoldChange, ymin = -1, ymax = -0.35),
    color = "#faa400", size = 0.5, linewidth = 0.1
  ) +
  # Add lineranges for AraC_Sensitivity genes
  geom_linerange(
    data = data_AraC_Sensitivity,
    aes(x = log2FoldChange, ymin = -1.75, ymax = -1.1),
    color = "#bebebe", size = 0.5, linewidth = 0.1
  ) +
  # Add text annotations for AraC_Resistance and AraC_Sensitivity
  annotate(
    "text", x = -3, y = -0.7, label = "AraC_Resistance",
    color = "#faa400", size = 4.8, hjust = 0
  ) +
  annotate(
    "text", x = -3, y = -1.4, label = "AraC_Sensitivity",
    color = "#bebebe", size = 4.8, hjust = 0
  ) +
  # Add enrichment statistics for AraC_Resistance
  annotate(
    "text", x = 3.0, y = -0.7, label = "ES: 0.38\npadj < 0.001",
    color = "#faa400", size = 4.5, lineheight = 0.8, hjust = 1
  ) +
  # Add enrichment statistics for AraC_Sensitivity
  annotate(
    "text", x = 3.0, y = -1.4, label = "NES: -0.495\npadj < 0.001",
    color = "#bebebe", size = 4.5, lineheight = 0.8, hjust = 1
  )

# Display the final plot
p1
ggsave(p1, filename = "../03.out/figure1/05.fpmtb_AraC_Ida_valcano.pdf", width = 6.5, height = 6)
```

### Ara-C + Dau

```{r}
library(clusterProfiler)
library(enrichplot)
library(fgsea)
library(dplyr)
library(tidyverse)

degene_mat <- read.table("../02.data/fpmtb_AraC_Dau_DESeq2_out_R_S_with_normalized_mat.txt", header = TRUE)
# Prepare ranked gene list for GSEA
DE_list <- degene_mat %>%
  select(ID = 9, Corr = 3) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na() %>%
  mutate(rank = rank(Corr, ties.method = "random"))

geneList <- DE_list$Corr
names(geneList) <- DE_list$ID
geneList <- sort(geneList, decreasing = TRUE)

# Load AraC gene sets
araC_gene_sets <- read_csv("../02.data/AraC_TopBottom1000_Genes.csv")
AraC_GeneSets_df <- araC_gene_sets %>%
  select(term = category, gene = Symbol) %>%
  distinct()  

# Convert TERM2GENE to a list
gene_sets_list <- split(AraC_GeneSets_df$gene, AraC_GeneSets_df$term)

# Run standard fgsea (instead of GSEA from clusterProfiler)
fgsea.res <- fgsea(
  pathways = gene_sets_list,
  stats = geneList,
  minSize = 1,
  maxSize = 1200,
  nperm = 100000
)

fgsea.res %>%
  arrange(padj)

fgsea.res$ES

# Plot enrichment for AraC_Resistance
plotEnrichment(gene_sets_list[["AraC_Resistance"]], geneList) +
  ggtitle("Enrichment of AraC_Resistance Genes")

# Plot enrichment for AraC_Sensitivity
plotEnrichment(gene_sets_list[["AraC_Sensitivity"]], geneList) +
  ggtitle("Enrichment of AraC_Sensitivity Genes")

```

```{r}
# Load necessary libraries
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(ggrastr)

# Load data
diff <- read.table("../02.data/fpmtb_AraC_Dau_DESeq2_out_R_S_with_normalized_mat.txt", header = TRUE)

# Cap log2FoldChange to [-4, 4] and padj to [1e-6, 1]
diff <- diff %>%
  mutate(
    log2FoldChange = case_when(
      log2FoldChange > 4 ~ 4,
      log2FoldChange < -4 ~ -4,
      TRUE ~ log2FoldChange
    ),
    padj = case_when(
      padj < 1e-6 ~ 1e-6,
      TRUE ~ padj
    )
  )

# Classify genes based on log2FoldChange and adjusted p-value
diff <- diff %>%
  mutate(g = case_when(
    log2FoldChange > 0.5 & padj < 0.05 ~ "up",
    log2FoldChange < -0.5 & padj < 0.05 ~ "down",
    TRUE ~ "stable"
  ))

# Extract significantly upregulated and downregulated genes
diff_up <- diff %>% filter(g == "up")
diff_down <- diff %>% filter(g == "down")

# Load AraC gene sets
araC_gene_sets <- read_csv("../02.data/AraC_TopBottom1000_Genes.csv")

# Extract AraC_Resistance and AraC_Sensitivity gene symbols
AraC_Resistance <- araC_gene_sets %>% filter(category == "AraC_Resistance") %>% pull(Symbol)
AraC_Sensitivity <- araC_gene_sets %>% filter(category == "AraC_Sensitivity") %>% pull(Symbol)

# Classify genes in the differential expression dataset
diff <- diff %>%
  mutate(gene_col = case_when(
    Symbol %in% AraC_Resistance ~ "AraC_Resistance",
    Symbol %in% AraC_Sensitivity ~ "AraC_Sensitivity",
    TRUE ~ "all"
  ))

# Prepare data subsets
data_bg <- diff %>% filter(gene_col == "all")
data_geneset <- diff %>% filter(grepl("AraC_", gene_col))
data_AraC_Resistance <- diff %>% filter(gene_col == "AraC_Resistance")
data_AraC_Sensitivity <- diff %>% filter(gene_col == "AraC_Sensitivity")

# Define the base plot
p <- ggplot(diff, aes(x = log2FoldChange, y = -log10(padj))) +
  # Add shaded regions for significant up- and down-regulated genes
  annotate(
    geom = "rect", fill = "#ea838e", alpha = 0.3,
    xmin = min(diff_up$log2FoldChange), xmax = max(diff_up$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  annotate(
    geom = "rect", fill = "#827fd6", alpha = 0.3,
    xmin = min(diff_down$log2FoldChange), xmax = max(diff_down$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  # Plot background points
  geom_point_rast(
    data = data_bg, shape = 21, color = "black", alpha = 0.2,
    size = 1.3, stroke = 0.6
  ) +
  # Plot gene set and key points
  geom_point_rast(
    data = data_geneset,
    aes(fill = gene_col), shape = 21, color = "black",
    size = 2, stroke = 0.8
  ) +
  # Custom fill colors for gene sets and key genes
  scale_fill_manual(
    values = c(
      "AraC_Resistance" = "#faa400",
      "AraC_Sensitivity" = "#bebebe"
    )
  ) +
  # Add reference lines
  geom_vline(xintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(
    yintercept = -log10(0.05), color = "#ff7d82",
    linewidth = 0.6, linetype = "dotted"
  ) +
  # Customize x-axis
  scale_x_continuous(limits = c(-4, 4), breaks = seq(-4, 4, 1)) +
  # Customize y-axis
  scale_y_continuous(limits = c(-2, 7), breaks = seq(0, 6.5, 2), expand = c(0, 0)) +
  # Apply classic theme and customize
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  ) +
  # Customize axis labels
  xlab("Resistant vs Sensitive (log2FC)") +
  ylab("Significance (-log10 adj.p-val.)")

# Add lineranges and annotations to the plot (p1)
p1 <- p +
  # Add lineranges for AraC_Resistance genes
  geom_linerange(
    data = data_AraC_Resistance,
    aes(x = log2FoldChange, ymin = -1, ymax = -0.35),
    color = "#faa400", size = 0.5, linewidth = 0.1
  ) +
  # Add lineranges for AraC_Sensitivity genes
  geom_linerange(
    data = data_AraC_Sensitivity,
    aes(x = log2FoldChange, ymin = -1.75, ymax = -1.1),
    color = "#bebebe", size = 0.5, linewidth = 0.1
  ) +
  # Add text annotations for AraC_Resistance and AraC_Sensitivity
  annotate(
    "text", x = -3, y = -0.7, label = "AraC_Resistance",
    color = "#faa400", size = 4.8, hjust = 0
  ) +
  annotate(
    "text", x = -3, y = -1.4, label = "AraC_Sensitivity",
    color = "#bebebe", size = 4.8, hjust = 0
  ) +
  # Add enrichment statistics for AraC_Resistance
  annotate(
    "text", x = 3.0, y = -0.7, label = "ES: 0.43\npadj = 0.02",
    color = "#faa400", size = 4.5, lineheight = 0.8, hjust = 1
  ) +
  # Add enrichment statistics for AraC_Sensitivity
  annotate(
    "text", x = 3.0, y = -1.4, label = "ES: -0.51\npadj < 0.001",
    color = "#bebebe", size = 4.5, lineheight = 0.8, hjust = 1
  )

# Display the final plot
p1
ggsave(p1, filename = "../03.out/figure1/05.fpmtb_AraC_dau_valcano.pdf", width = 6.5, height = 6)
```

## Pathway Enrichment via GSEA Using Fold change of Genes between Resistant and Sensitive (Ara-C + Ida)

```{r gsea-7+3, eval=TRUE, message = FALSE, warning = FALSE}
library(clusterProfiler)
library(enrichplot)
library(GseaVis)
library(reactable)
library(tidyverse)

# Define a GSEA plotting function
theme_set(theme_classic())
plot_gseaNb <- function(gsea_obj, gene_set, curve_colors = c('#7582c1', '#f99f1c'), 
                        ht_colors = curve_colors, rank_colors = c('#7582c1', 'white', '#f99f1c')) {
  gseaNb(
    object = gsea_obj,
    geneSetID = gene_set,
    addPval = TRUE,
    subPlot = 3,
    curveCol = curve_colors,
    htCol = ht_colors,
    rankCol = rank_colors
  )
}

# Load correlation results
degene_mat <- read.table("../02.data/fpmtb_AraC_Dau_DESeq2_out_R_S_with_normalized_mat.txt", header = TRUE)

# Prepare ranked gene list for GSEA
DE_list <- degene_mat %>%
  select(ID = 9, Corr = 3) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na() %>%
  mutate(rank = rank(Corr, ties.method = "random"))

geneList <- DE_list$Corr
names(geneList) <- DE_list$ID
geneList <- sort(geneList, decreasing = TRUE)

# Load KEGG gene sets
geneSet <- read.gmt("../02.data/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

# Perform GSEA
set.seed(123456)
gsea.enrich <- GSEA(
  geneList,
  TERM2GENE = geneSet,
  pvalueCutoff = 1,
  nPermSimple = 100000,
  maxGSSize = 800,
  pAdjustMethod = "BH",
  seed = TRUE
)
gsea.out <- gsea.enrich@result
gsea.out <- gsea.out[, -11]

# Display GSEA table
reactable(
  gsea.out[, 1:10],
  pagination = TRUE,
  highlight = TRUE,
  searchable = TRUE,
  defaultPageSize = 10
)

# Dot plot of significant GSEA results
sig.gsea <- gsea.out %>%
  filter(p.adjust < 0.05) %>%
  arrange(NES) %>%
  mutate(Description = factor(gsub("KEGG_", "", Description), levels = gsub("KEGG_", "", Description)))

p <- ggplot(sig.gsea, aes(x = NA, y = Description, color = NES, size = -log10(p.adjust))) +
  geom_point(stroke = 1) +
  labs(x = "", y = "") +
  scale_color_gradient2(low = "#7582c1", mid = "white", high = "#f99f1c") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))

p

# Visualize selected KEGG pathways

plot_gseaNb(gsea.enrich, gene_set = "KEGG_JAK_STAT_SIGNALING_PATHWAY")
plot_gseaNb(gsea.enrich, gene_set = "KEGG_MAPK_SIGNALING_PATHWAY")
plot_gseaNb(gsea.enrich, gene_set = "KEGG_PYRIMIDINE_METABOLISM")
plot_gseaNb(gsea.enrich, gene_set = "KEGG_RIBOSOME")

```

## DFS-Linked Gene Signatures in TCGA-LAML Cohort

To explore whether AraC-associated transcriptional profiles correlate with clinical outcomes, we analyzed RNA-seq and survival data from 141 AML patients treated with Ara-C in the TCGA-LAML cohort. Specifically, we identified genes associated with disease-free survival (DFS) using multivariate Cox proportional hazards regression, and used GSEA to determine pathway-level enrichment of these prognostic signals.

```{r tcga-cox-analysis, eval=FALSE, message = FALSE, warning = FALSE}
# ==============================
# Step 0: Load Required Libraries
# ==============================
# library(survival)
# library(readr)
# library(dplyr)
# library(tibble)
# 
# # ==============================
# # Step 1: Load Expression and Clinical Data
# # ==============================
# # Read expression matrix: must include 'Symbol' column for gene symbols
# tcga_expr <- read.table("../02.data/tcga_all_179_TPM_matrix.txt", 
#                         header = TRUE, sep = "\t")
# 
# # Confirm Symbol column exists
# if (!"Symbol" %in% colnames(tcga_expr)) {
#   stop("No 'Symbol' column found in expression matrix.")
# }
# 
# # Clean up Symbol column
# tcga_expr$Symbol <- trimws(as.character(tcga_expr$Symbol))
# 
# # Load clinical info
# clin_info <- read.csv("../02.data/TCGA_DFS_AraC_clinical.csv", header = TRUE)
# clin_info$sample <- as.character(clin_info$sample)
# 
# # Identify common samples between expression and clinical data
# common_samples <- intersect(colnames(tcga_expr), clin_info$sample)
# common_samples <- sort(common_samples)
# 
# # Subset expression and clinical to common samples
# expr_sub <- tcga_expr[, c("Symbol", common_samples)]
# clin_info <- clin_info[match(common_samples, clin_info$sample), ]
# rownames(clin_info) <- clin_info$sample
# 
# # ==============================
# # Step 2: Filter Genes 
# # ==============================
# 
# # Filter and aggregate expression data by gene Symbol
# expr_filtered <- expr_sub %>%
#   group_by(Symbol) %>%
#   summarise(across(where(is.numeric), mean), .groups = "drop") %>%
#   filter(rowMeans(across(where(is.numeric))) > 0.25) %>%  
#   column_to_rownames("Symbol")
# 
# # Ensure order matches
# expr_filtered <- expr_filtered[, common_samples]
# 
# # Transpose to sample x gene
# expr_transposed <- as.data.frame(t(expr_filtered))
# 
# # ==============================
# # Step 3: Combine with Clinical Data
# # ==============================
# cox_data <- bind_cols(
#   clin_info %>% select(DFS_time, DFS_status, Age, Sex, WBC, cytogenetic_risk),
#   expr_transposed
# )
# 
# # ==============================
# # Step 4: Cox Regression with Progress Bar
# # ==============================
# gene_names <- colnames(expr_transposed)
# cox_results <- data.frame(
#   Gene = gene_names,
#   Coefficient = NA_real_,
#   HR = NA_real_,
#   P.value = NA_real_,
#   stringsAsFactors = FALSE
# )
# 
# pb <- txtProgressBar(min = 0, max = length(gene_names), style = 3)
# 
# for (i in seq_along(gene_names)) {
#   gene <- gene_names[i]
#   formula <- as.formula(paste0("Surv(DFS_time, DFS_status) ~ Age + Sex + WBC + cytogenetic_risk + `", gene, "`"))
#   
#   fit <- tryCatch({
#     summary(coxph(formula, data = cox_data))
#   }, error = function(e) NULL)
# 
#   if (!is.null(fit)) {
#     gene_row <- grep(paste0("^", gene, "$"), rownames(fit$coefficients))
#     if (length(gene_row) == 1) {
#       cox_results[i, "Coefficient"] <- fit$coefficients[gene_row, 1]
#       cox_results[i, "HR"] <- exp(fit$coefficients[gene_row, 1])
#       cox_results[i, "P.value"] <- fit$coefficients[gene_row, 5]
#     }
#   }
# 
#   setTxtProgressBar(pb, i)
# }
# close(pb)
# 
# # Save to file
# write.csv(cox_results, "../02.data/TCGA_AraC_GeneWise_Cox_Results.csv", row.names = FALSE)

```

```{r tcga-dfs-gsea-analysis, eval=TRUE, message = FALSE, warning = FALSE}
# ==============================
# Load Libraries
# ==============================
library(tidyverse)
library(clusterProfiler)
library(GseaVis)
library(cowplot)

# ==============================
# Step 1: Load and Prepare Ranked Gene List (Cox Coefficients)
# ==============================
cox_results <- read.csv("../02.data/TCGA_AraC_GeneWise_Cox_Results.csv")

# Keep only necessary columns and handle duplicates or NAs
ranked_genes <- cox_results %>%
  select(Gene, Coefficient) %>%
  distinct(Gene, .keep_all = TRUE) %>%
  drop_na() %>%
  arrange(desc(Coefficient))  # Sort by effect size

# Create named numeric vector for GSEA
geneList <- setNames(ranked_genes$Coefficient, ranked_genes$Gene)
geneList <- sort(geneList, decreasing = TRUE)

# ==============================
# Step 2: Load KEGG Gene Sets
# ==============================
# Option A: Local KEGG GMT file
kegg_sets <- read.gmt("../02.data/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

# ==============================
# Step 3: Run GSEA
# ==============================
set.seed(123456)
gsea_result <- GSEA(
  geneList,
  TERM2GENE = kegg_sets,
  pvalueCutoff = 1,
  pAdjustMethod = "BH",
  nPermSimple = 1000,
  maxGSSize = 800,
  minGSSize = 10,
  seed = TRUE
)

# Save results
gsea_df <- gsea_result@result
write.csv(gsea_df, "../02.data/TCGA_Cox_GSEA_KEGG.csv", row.names = FALSE)

# ==============================
# Step 4: Plot Selected Enriched Pathways
# ==============================
# List of key gene sets to visualize
pathways_to_plot <- c(
  "KEGG_JAK_STAT_SIGNALING_PATHWAY",
  "KEGG_CHEMOKINE_SIGNALING_PATHWAY",
  "KEGG_RIBOSOME",
  "KEGG_FATTY_ACID_METABOLISM"
)

gsea_plots <- lapply(pathways_to_plot, function(pathway_id) {
  gseaNb(
    object = gsea_result, 
    geneSetID = pathway_id,
    newGsea = F, addPval = TRUE,
    pvalX = 0.55, pvalY = 0.75,
    pCol = "black", pHjust = 0, subPlot = 2,
    curveCol = c('#7582c1', '#f99f1c'),
    htCol = c("#7582c1", "#f99f1c"),
    rankCol = c("#7582c1", "white", "#f99f1c")
  )
})

combined_plot <- cowplot::plot_grid(
  plotlist = gsea_plots, ncol = 4,
  align = "hv"
)
combined_plot

ggsave(
  filename = "../03.out/figure1/TCGA_GSEA_Selected_KEGG.pdf",
  plot = combined_plot, width = 13, height = 3
)

```

