# Figure 2: Development of the ARG24 Signature and Functional Analysis

## Survival Outcomes Based on Induction Response

```{r, eval=TRUE, message = FALSE, warning = FALSE}
# Load necessary libraries
library(readxl)
library(survival)
library(survminer)
library(RColorBrewer)

# Load patient clinical data from RJAML cohort
mat <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
# library(tableone)
# 
# tab1 <- CreateTableOne(
#   vars = c("Age", "Gender","WBC", "HB", "PLT", "BM Blast", "HSCT", "NPM1", "FLT3-ITD"),
#   strata = "CR1_response",
#   data = mat,
#   factorVars = c("Gender", "HSCT", "NPM1", "FLT3-ITD")
# )
# 
# final_table <- print(tab1, showAllLevels = TRUE)
# write.csv(final_table, "../02.data/Table1_rjaml.csv")

# Generate survival plots for OS and EFS grouped by clinical response
plot <- list()
plot[[1]] <- ggsurvplot(
  survfit(Surv(OS_Time, OS_Status) ~ Response, data = mat),
  palette = brewer.pal(8, "Dark2"),
  pval = TRUE,
  pval.size = 4,
  xlab = "Survival time in days",
  conf.int = FALSE,
  risk.table = FALSE,
  surv.median.line = 'hv',
  legend.title = "Responses to induction",
  legend.labs = c(
    paste0("CR1 (N = ", sum(mat$Response == "CR1", na.rm = TRUE), ")"),
    paste0("CR2 (N = ", sum(mat$Response == "CR2", na.rm = TRUE), ")"),
    paste0("NCR (N = ", sum(mat$Response == "NCR", na.rm = TRUE), ")")
  ),
  legend = c(0.65, 0.85)
)

plot[[2]] <- ggsurvplot(
  survfit(Surv(EFS_Time, EFS_Status) ~ Response, data = mat),
  palette = brewer.pal(8, "Dark2"),
  pval = TRUE,
  pval.size = 4,
  xlab = "Survival time in days",
  conf.int = FALSE,
  risk.table = FALSE,
  surv.median.line = 'hv',
  legend.title = "Responses to induction",
  legend.labs = c(
    paste0("CR1 (N = ", sum(mat$Response == "CR1", na.rm = TRUE), ")"),
    paste0("CR2 (N = ", sum(mat$Response == "CR2", na.rm = TRUE), ")"),
    paste0("NCR (N = ", sum(mat$Response == "NCR", na.rm = TRUE), ")")
  ),
  legend = c(0.65, 0.85)
)

# Arrange and print survival plots side-by-side
p <- arrange_ggsurvplots(plot, print = FALSE, ncol = 2, nrow = 1)
p
```

## Differential Expression Analysis Between CR and NCR

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(DESeq2)
library(readxl)
sample <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
tpm <- read.table("../02.data/rjaml_all_TPM_matrix.txt", header=T)
tpm <- subset(tpm, Type %in% c("protein_coding", "miRNA", "lincRNA", "antisense",
                               "processed_transcript",
                               "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene",
                               "transcribed_unprocessed_pseudogene","macro_lncRNA", "TEC",
                               "3prime_overlapping_ncRNA","sense_intronic", "sense_overlapping",
                               "unprocessed_pseudogene"))
tpm <- tpm[, c("ID", "Symbol", "Type", as.character(sample$Tumor_Sample_Barcode))]
# # filter low-express genes 
# tpm <- tpm[rowMeans(tpm[,4:ncol(tpm)]) >= 0.05, ]
# # write.csv(tpm, "RJAML_221_patients_TPM.csv")
# mat <- read.table("../../rjaml_all_count_matrix.txt", header=TRUE, row.names = 1) 
# mat <- mat[tpm$ID, as.character(sample$Tumor_Sample_Barcode)]
# 
# ## run DESEQ2
# source("../01.script/RunDESeq2.R")
# 
# RunDESeq2(count_mat = mat, n.cont = 157, n.treat = 64, 
#           prefix = "1_rjaml_DESeq2_out_NCR_CR", sort.p = FALSE, 
#           merge.normalized = TRUE, 
#           normalized_mat = tpm)

# data_path <- "../02.data/1_rjaml_DESeq2_out_NCR_CR_with_normalized_mat.txt"
# deseq_output <- read.table(data_path, header = TRUE)
# 
# # Select the first 9 columns for plotting
# selected_data <- deseq_output[, 1:9]
# write.csv(selected_data, "../02.data/rjaml_de.out.csv")
```

## Visualization: Volcano Plot of Differentially Expressed Genes

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(DT)
source("plotVolcano.R")
deseq_output <- read.csv("../02.data/rjaml_de.out.csv", header = TRUE)

# Cap extreme log2 fold changes for better visualization
deseq_output$log2FoldChange[deseq_output$log2FoldChange > 5] <- 5
deseq_output$padj[deseq_output$padj < 1e-15] <- 1e-15

degs <- subset(deseq_output, padj < 0.05 & abs(log2FoldChange) > 0.5)
degs$group <- ifelse(degs$log2FoldChange > 0, "Upregulated_in_Resistant", "Upregulated_in_Sensitive")
table(degs$group)

custom_colors <- brewer.pal(4, "PRGn")  # Color palette
selected_genes = data.frame(gsym = character(), Pathway = character())

# Generate volcano plot
volcano_plot <- plotVolcano(
  data = deseq_output, 
  gene.col = "Symbol", 
  logFC.col = "log2FoldChange", pval.col = "padj",
  logFCcut = 0, logFCcut2 = 0.5, logFCcut3 = 1,
  pvalCut = 0.05, pvalCut2 = 0.01, pvalCut3 = 0.001,
  threshold_colors = custom_colors, 
  plot_mode = "advanced", x_range = c(-5, 5), y_range = c(0, 15),
  title = "", 
  label = FALSE,
  selected_genes = selected_genes
)

# Print and save the volcano plot
volcano_plot
ggsave("../03.out/figure2/02.volcano_plot_patients.pdf", plot = volcano_plot, width = 6, height = 5.5)

# Display data table interactively
datatable(selected_data, fillContainer = TRUE, options = list(pageLength = 10))
```
## Ara-C Responsive Gene Overlap

```{r}
library(ggvenn)
library(tidyverse)

patient_df <- read_csv("../02.data/BeatAML_AraC_AUC_Gene_Corr.csv")
rjaml_df <- read.csv("../02.data/rjaml_de.out.csv")
rjaml_df <- rjaml_df[!duplicated(rjaml_df$Symbol), ]
# Rank both
top_bottom_genes <- function(df) {
  df <- df %>% arrange(desc(Z_score))
  top1000 <- df$Symbol[1:1000]
  bottom1000 <- df$Symbol[(nrow(df) - 999):nrow(df)]
  unique(na.omit(c(top1000, bottom1000)))
}

genes_patient <- top_bottom_genes(patient_df)
genes_rjaml <- subset(rjaml_df, padj < 0.05 & abs(log2FoldChange) > 0.5)$Symbol

gene_sets <- list(
  Patient = genes_patient,
  RJAML = genes_rjaml
)

# Create Venn plot with ggvenn
p_venn <- ggvenn(
  gene_sets,
  # fill_color = c("#E69F00", "#56B4E9"),
  stroke_size = 0.5,
  set_name_size = 5,
  text_size = 4
)
p_venn
# Define gene universe = genes that exist in both datasets
gene_universe <- intersect(patient_df$Symbol, rjaml_df$Symbol)

# Logical vectors for membership
in_patient <- gene_universe %in% genes_patient
in_rjaml <- gene_universe %in% genes_rjaml

# Contingency table and p-value
contingency <- table(Patient = in_patient, RJAML = in_rjaml)
fisher_p <- fisher.test(contingency)$p.value
p_label <- paste0("Fisher's Exact Test\np = ", formatC(fisher_p, format = "e", digits = 2))
p_label

ggsave(p_venn, filename = "../03.out/figure1/arac_genes_rjaml_overlap.pdf", width = 5, height = 5)

```

## Feature Selection Using LASSO Regression

```{r, eval=TRUE, message = FALSE, warning = FALSE}
# library(tidyverse)
# library(survival)
# library(survminer)
# library(glmnet)
# library(pROC)

# # Load data
# data_path <- "../02.data/rjaml_intersected_tpm.csv"
# mat <- read.csv(data_path, header = TRUE, row.names = 1)
# 
# # Prepare the response variable and feature matrix
# response <- mat$Group
# features <- log2(mat[,-1] + 1) %>% as.matrix()
# 
# # Initialize variables for cross-validation
# lambda_values <- numeric(10)
# set.seed(123)
# 
# # Perform 10-fold cross-validation to determine optimal lambda
# for (i in seq_along(lambda_values)) {
#   cvfit <- cv.glmnet(features, response, family = "binomial", alpha = 1, nlambda = 1000)
#   lambda_values[i] <- cvfit$lambda.min
# }
# 
# # Calculate the average lambda from cross-validation
# optimal_lambda <- mean(lambda_values)
# 
# # Fit the final model using the determined lambda
# final_fit <- glmnet(features, response, family = "binomial", alpha = 1, lambda = optimal_lambda)
# 
# # Extract non-zero coefficients from the model
# coef_info <- as.matrix(coef(final_fit, s = optimal_lambda))
# significant_genes <- coef_info[coef_info[, 1] != 0, , drop = FALSE]
# rownames(significant_genes) <- rownames(coef_info)[coef_info[, 1] != 0]
# 
# # Identify significant genes and calculate risk scores
# selected_features <- features[, colnames(features) %in% rownames(significant_genes)]
# 
# # Define a function to calculate risk scores based on LASSO results
# calculate_risk_scores <- function(features, coefficients) {
#   scores <- sweep(features, 2, coefficients, `*`)
#   risk_scores <- rowSums(scores)
#   data.frame(risk_score = risk_scores)
# }
# 
# # Calculate patient risk scores
# risk_scores <- calculate_risk_scores(selected_features, significant_genes[, 1])
# 
# # Combine features and risk scores for output
# final_data <- cbind(selected_features, risk_scores)
# 
# # Save significant genes and the final data with risk scores
# write.csv(significant_genes, "../02.data/24-gene_signature.csv")
# write.csv(final_data, "../02.data/24-Gene_RJAML_riskScore.csv")
```

## Visualization of the ARG24 Gene Signature

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(ggpubr)
library(knitr)

# Load data
gene_data_path <- "../02.data/24-gene_signature.csv"
gene_signature <- read.csv(gene_data_path, header = TRUE)

# Categorize coefficients into groups 'Pos' or 'Neg'
gene_signature$group <- ifelse(gene_signature$Coef < 0, "Neg", "Pos")

# Display signature gene table
knitr::kable(gene_signature)

# Create a bar plot of gene coefficients
p1 <- ggbarplot(
  gene_signature, x = "Gene", y = "Coef", fill = "group", color = "white",
  palette = "jco", sort.val = "asc", sort.by.groups = FALSE, x.text.angle = 60,
  xlab = "Gene", ylab = "Coefficient", legend.title = "", rotate = FALSE
) +
  geom_text(aes(label = sprintf("%.2f", Coef)), vjust = -0.5, color = "black") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Print the plot
p1
```

## Expression Levels of ARG24 Genes in CR vs NCR Groups

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(readxl)
library(ggplot2)
library(ggpubr)
library(ggrastr)
library(reshape2)
library(RColorBrewer)
library(pROC)
library(tidyverse)

# Read and transform data
gene_expression_path <- "../02.data/24_gene_tpm_rjaml.csv"
rj_dat <- read.csv(gene_expression_path, header = TRUE, check.names = FALSE) %>%
  melt() %>%
  mutate(value = log2(value + 1))
rj_dat$Group <- gsub("Refractory", "NCR", rj_dat$Group)

# Create a violin plot
p <- ggplot(rj_dat, aes(x = Group, y = value, color = Group)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.4, alpha = 0.75, outlier.shape = NA) +
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  facet_wrap(~ variable, scales = "free", nrow = 3) +
  xlab("") +
  ylab("Gene expression (log2-transformed)") +
  scale_color_manual(values = c("CR" = "#1B9E77", "NCR" = "#7570B3")) +
  theme_pubr(base_size = 12) +
  theme(
    legend.position = "none",
    strip.text.x = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    text = element_text(size = 12)
  ) +
  stat_compare_means(comparisons = list(c("CR", "NCR")), method = 'wilcox.test')

# Print and save the plot
p
# ggsave(p, filename = "../03.out/figure2/02.24genes_violin.pdf", width = 14, height = 9)
```

## ARG24 Score Distributions and ROC Analysis

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(pROC)
library(readxl)
library(RColorBrewer)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(ggrastr)

rj_dat <- read.csv("../02.data/24-Gene_RJAML_riskScore.csv", header = TRUE)

# Violin plot of ARG24 scores by group
p <- ggplot(rj_dat, aes(x = group2, y = risk_score, color = group2)) + 
  geom_violin() +
  geom_boxplot_jitter(width=0.4, alpha=0.75) + 
  geom_quasirandom_rast(width=0.25, alpha=0.5) +
  theme_pubr() + 
  xlab("") + ylab("Risk score") +
  theme(strip.text.x = element_text(size=13)) +
  stat_compare_means(comparisons = list(c('CR1', 'CR2'), c('CR1', 'Refractory'), c('CR2', 'Refractory')), method='wilcox.test') + 
  scale_color_manual(values = c("#1B9E77", "#D95F02", "#7570B3")) +
  theme(legend.position='None')
print(p)

# ROC analysis for CR1
rj_mat <- read.csv("../02.data/rjAML_D5-PBCR.csv", header = TRUE)
auc1 <- auc(rj_mat$CR1_response, rj_mat$risk_score_binary)
roc1 <- roc(rj_mat$CR1_response, rj_mat$risk_score_binary)
auc2 <- auc(rj_mat$CR1_response, rj_mat$risk_score)
roc2 <- roc(rj_mat$CR1_response, rj_mat$risk_score)
auc3 <- auc(rj_mat$CR1_response, rj_mat$D5.PBCR)
roc3 <- roc(rj_mat$CR1_response, rj_mat$D5.PBCR)

plot(roc1, col=brewer.pal(8, "Dark2")[1])
plot(roc2, add=TRUE, col=brewer.pal(8, "Dark2")[2])
plot(roc3, add=TRUE, col=brewer.pal(8, "Dark2")[3])
legend("bottomright", 
       legend = c(paste0("RiskScore-Binary AUC=", round(auc1, 3)),
                  paste0("RiskScore-Continuous AUC=", round(auc2, 3)),
                  paste0("D5-PBCR AUC=", round(auc3, 3))),
       col = brewer.pal(8, "Dark2")[1:3],
       lty = 1)
title(main = "Predictive power in CR1")

# ROC analysis for CR2
auc4 <- auc(rj_mat$CR2_response, rj_mat$risk_score_binary)
roc4 <- roc(rj_mat$CR2_response, rj_mat$risk_score_binary)
auc5 <- auc(rj_mat$CR2_response, rj_mat$risk_score)
roc5 <- roc(rj_mat$CR2_response, rj_mat$risk_score)
auc6 <- auc(rj_mat$CR2_response, rj_mat$D5.PBCR)
roc6 <- roc(rj_mat$CR2_response, rj_mat$D5.PBCR)

plot(roc4, col=brewer.pal(8, "Dark2")[1])
plot(roc5, add=TRUE, col=brewer.pal(8, "Dark2")[2])
plot(roc6, add=TRUE, col=brewer.pal(8, "Dark2")[3])
legend("bottomright", 
       legend = c(paste0("RiskScore-Binary AUC=", round(auc4, 3)),
                  paste0("RiskScore-Continuous AUC=", round(auc5, 3)),
                  paste0("D5-PBCR AUC=", round(auc6, 3))),
       col = brewer.pal(8, "Dark2")[1:3],
       lty = 1)
title(main = "Predictive power in CR2")

# Boxplot by D5-PBCR status
rj_mat$D5.PBCR <- ifelse(rj_mat$D5.PBCR == "1", "Positive", "Negative")
p <- ggplot(rj_mat, aes(x = D5.PBCR, y = risk_score, color = D5.PBCR)) + 
  geom_violin() +
  geom_boxplot_jitter(width=0.4, alpha=0.75) + 
  geom_quasirandom_rast(width=0.25, alpha=0.5) +
  theme_pubr() + 
  xlab("") + ylab("Risk score") +
  theme(strip.text.x = element_text(size=13)) +
  stat_compare_means(comparisons = list(c('Positive', 'Negative')), method='wilcox.test') + 
  scale_color_manual(values = c("#66A61E", "#E6AB02")) +
  theme(legend.position='None')
print(p)

# Fisher test for D5-PBCR vs ARG24 high/low
matrix_data <- matrix(c(53, 35, 33, 54), ncol = 2, byrow = TRUE)
colnames(matrix_data) <- c("High-riskScore", "Low-riskScore")
rownames(matrix_data) <- c("D5-PBCR+", "D5-PBCR-")
fisher_result <- fisher.test(matrix_data)
fisher_result

# Overall AUC for ARG24 on CR1 and CR2 groups
auc1 <- auc(rj_dat$group, rj_dat$risk_score)
roc1 <- roc(rj_dat$group, rj_dat$risk_score)
auc2 <- auc(rj_dat$group3, rj_dat$risk_score)
roc2 <- roc(rj_dat$group3, rj_dat$risk_score)

plot(smooth(roc1), col=brewer.pal(8, "Dark2")[3])
plot(smooth(roc2), add=TRUE, col=brewer.pal(8, "Dark2")[2])
legend("bottomright", 
       legend = c(paste0("CR1 AUC=", round(auc1, 3)),
                  paste0("CR2 AUC=", round(auc2, 3))),
       col = c(brewer.pal(8, "Dark2")[3], brewer.pal(8, "Dark2")[2]),
       lty = 1)
title(main = "RJAML cohort (N=221)")
```

## Comparison of ARG24 with LSC17 and PS29MRC Signatures

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(readxl)
library(dplyr)
library(pROC)
library(RColorBrewer)

calculate_risk_scores <- function(features, coefficients) {
  scores <- sweep(features, 2, coefficients, `*`)
  risk_scores <- rowSums(scores)
  data.frame(risk_score = risk_scores)
}

sample <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
tpm <- read.table("../02.data/rjaml_all_TPM_matrix.txt", header=TRUE)
tpm <- tpm[, c("ID", "Symbol", "Type", as.character(sample$Tumor_Sample_Barcode))]

# LSC17
lsc17 <- read.table("../02.data/LSC17_signatture.txt", header = TRUE)
lsc17_mat <- subset(tpm, Symbol %in% lsc17$gene)
rownames(lsc17_mat) <- lsc17_mat$Symbol
lsc17_mat <- lsc17_mat[, -c(1:3)]
lsc17_mat <- lsc17_mat[rownames(lsc17), ]
lsc17_mat <- log2(t(lsc17_mat) + 1)
lsc17_scores <- calculate_risk_scores(lsc17_mat, lsc17[, 1])
colnames(lsc17_scores) <- "LSC17"

# PS29MRC
ps29mrc <- read.table("../02.data/PS29MRC_signatture.txt", header = TRUE)
ps29mrc_mat <- subset(tpm, Symbol %in% ps29mrc$gene)
rownames(ps29mrc_mat) <- ps29mrc_mat$Symbol
ps29mrc_mat <- ps29mrc_mat[, -c(1:3)]
ps29mrc_mat <- log2(ps29mrc_mat + 1)
ps29mrc_mat["MRC_high", ] <- ifelse(sample$Cytogenetic_risk == "Poor", 1, 0)
ps29mrc_mat["MRC_low", ] <- ifelse(sample$Cytogenetic_risk == "Good", -1, 0)
ps29mrc <- ps29mrc[rownames(ps29mrc_mat), ]
ps29mrc_mat <- t(ps29mrc_mat)
ps29mrc_scores <- calculate_risk_scores(ps29mrc_mat, ps29mrc[, 1])
colnames(ps29mrc_scores) <- "PS29MRC"

lsc17_ps29mrc <- cbind(lsc17_scores, ps29mrc_scores)
auc_dat <- cbind(lsc17_ps29mrc, sample[, c("risk_score", "D5-PBCR", "Response", "CR1_response")])

# ROC analysis - CR1
auc1 <- auc(auc_dat$CR1_response, auc_dat$LSC17)
roc1 <- roc(auc_dat$CR1_response, auc_dat$LSC17)
auc2 <- auc(auc_dat$CR1_response, auc_dat$PS29MRC)
roc2 <- roc(auc_dat$CR1_response, auc_dat$PS29MRC)
auc3 <- auc(auc_dat$CR1_response, auc_dat$risk_score)
roc3 <- roc(auc_dat$CR1_response, auc_dat$risk_score)

# pdf("../03.out/figure2/signatures_CR1_comparision.pdf", width = 4, height = 4)
plot(roc1, col=brewer.pal(8, "Dark2")[1])
plot(roc2, add=TRUE, col=brewer.pal(8, "Dark2")[2])
plot(roc3, add=TRUE, col=brewer.pal(8, "Dark2")[3])
legend("bottomright", 
       legend = c(paste0("LSC17 AUC=", round(auc1, 3)),
                  paste0("PS29MRC AUC=", round(auc2, 3)),
                  paste0("ARG24 AUC=", round(auc3, 3))),
       col = brewer.pal(8, "Dark2")[1:3],
       lty = 1)
title(main = "Predictive power in CR1")
# dev.off()

# ROC analysis - CR2
auc_dat$CR2_response <- ifelse(auc_dat$Response == "NCR", "NCR", "CR")
auc4 <- auc(auc_dat$CR2_response, auc_dat$LSC17)
roc4 <- roc(auc_dat$CR2_response, auc_dat$LSC17)
auc5 <- auc(auc_dat$CR2_response, auc_dat$PS29MRC)
roc5 <- roc(auc_dat$CR2_response, auc_dat$PS29MRC)
auc6 <- auc(auc_dat$CR2_response, auc_dat$risk_score)
roc6 <- roc(auc_dat$CR2_response, auc_dat$risk_score)

# pdf("../03.out/figure2/signatures_CR2_comparision.pdf", width = 4, height = 4)
plot(roc4, col=brewer.pal(8, "Dark2")[1])
plot(roc5, add=TRUE, col=brewer.pal(8, "Dark2")[2])
plot(roc6, add=TRUE, col=brewer.pal(8, "Dark2")[3])
legend("bottomright", 
       legend = c(paste0("LSC17 AUC=", round(auc4, 3)),
                  paste0("PS29MRC AUC=", round(auc5, 3)),
                  paste0("ARG24 AUC=", round(auc6, 3))),
       col = brewer.pal(8, "Dark2")[1:3],
       lty = 1)
title(main = "Predictive power in CR2")
# dev.off()
```

## Comparison of ARG24 with D5-PBCR

```{r, eval=TRUE, message = FALSE, warning = FALSE}
auc_dat1 <- subset(auc_dat, `D5-PBCR` %in% c("Pos", "Neg"))
auc_dat1$`D5-PBCR` <- ifelse(auc_dat1$`D5-PBCR` == "Pos", 1, 0)

# ROC analysis - CR1
auc1 <- auc(auc_dat1$CR1_response, auc_dat1$`D5-PBCR`)
roc1 <- roc(auc_dat1$CR1_response, auc_dat1$`D5-PBCR`)
auc2 <- auc(auc_dat1$CR1_response, auc_dat1$risk_score)
roc2 <- roc(auc_dat1$CR1_response, auc_dat1$risk_score)

# pdf("../03.out/figure2/signatures_CR1_D5PBCR.pdf", width = 4, height = 4)
plot(roc1, col = brewer.pal(8, "Dark2")[1])
plot(roc2, add = TRUE, col = brewer.pal(8, "Dark2")[2])
legend("bottomright",
       legend = c(paste0("D5-PBCR AUC=", round(auc1, 3)),
                  paste0("ARG24 AUC=", round(auc2, 3))),
       col = brewer.pal(8, "Dark2")[1:2],
       lty = 1)
title(main = "Predictive power in CR1")
# dev.off()

# ROC analysis - CR2
auc3 <- auc(auc_dat1$CR2_response, auc_dat1$`D5-PBCR`)
roc3 <- roc(auc_dat1$CR2_response, auc_dat1$`D5-PBCR`)
auc4 <- auc(auc_dat1$CR2_response, auc_dat1$risk_score)
roc4 <- roc(auc_dat1$CR2_response, auc_dat1$risk_score)

# pdf("../03.out/figure2/signatures_CR2_D5PBCR.pdf", width = 4, height = 4)
plot(roc3, col = brewer.pal(8, "Dark2")[1])
plot(roc4, add = TRUE, col = brewer.pal(8, "Dark2")[2])
legend("bottomright",
       legend = c(paste0("D5-PBCR AUC=", round(auc3, 3)),
                  paste0("ARG24 AUC=", round(auc4, 3))),
       col = brewer.pal(8, "Dark2")[1:2],
       lty = 1)
title(main = "Predictive power in CR2")
# dev.off()
```

## Single cell in primary AML cells that survived after cytarabine treatment or Control

```{r}

# GSE146590
library(dplyr)
library(Seurat)
library(patchwork)
library(stringr)
library(tidyr)
library(readr)
library(Matrix)
library(ggplot2)

# 1. Data Loading and Preprocessing --------------------------------
# Get all AML folders (AML1, AML2, etc.)
data_path <- '/dssg/home/acct-medljm/medljm-jqq/projects/jp_projects/chemoSig/02.data/gse146590_arac_scrna'
folders <- list.dirs(path = data_path, full.names = TRUE, recursive = FALSE)

# Read each dataset and create Seurat objects
sceList <- lapply(folders, function(folder) {
  message("Processing ", basename(folder))
  
  # Read components
  barcodes <- read.table(
    file.path(folder, "barcodes.tsv.gz"), 
    header = FALSE,
    stringsAsFactors = FALSE
  )[[1]]  # Get first column as cell names
  
  features <- read.table(
    file.path(folder, "features.tsv.gz"),
    header = FALSE,
    stringsAsFactors = FALSE
  )
  
  counts <- readMM(file.path(folder, "matrix.mtx.gz"))
  
  # Set proper dimnames
  rownames(counts) <- features$V1
  colnames(counts) <- barcodes
  
  # Create Seurat object with basic QC
  CreateSeuratObject(
    counts = counts,
    project = basename(folder),
    min.cells = 3,
    min.features = 200
  )
})

# 2. Data Integration ---------------------------------------------
# Merge all objects with sample identifiers
sample_names <- list.dirs(path = data_path, full.names = FALSE, recursive = FALSE)
gse146590 <- merge(sceList[[1]], y = sceList[-1], add.cell.ids = sample_names)

# Calculate mitochondrial percentage
gse146590[["percent.mt"]] <- PercentageFeatureSet(gse146590, pattern = "^MT-")

# 3. Robust Gene Annotation Update --------------------------------
library(org.Hs.eg.db)

# Clean ENSEMBL IDs and get gene symbols
# 3. Gene Annotation with Duplicate Handling ----------------------
library(org.Hs.eg.db)

# Clean ENSEMBL IDs and get gene symbols
ensembl_ids <- sub("\\..*", "", rownames(gse146590))
symbols <- mapIds(org.Hs.eg.db,
                 keys = ensembl_ids,
                 column = "SYMBOL",
                 keytype = "ENSEMBL",
                 multiVals = "first")

# Filter genes
keep_genes <- !is.na(symbols) & 
             !grepl("^LOC", symbols) & 
             symbols != ""

# Create new gene names vector
new_gene_names <- symbols[keep_genes]

# Fast duplicate detection and handling
if(anyDuplicated(new_gene_names)) {
  dup_genes <- unique(new_gene_names[duplicated(new_gene_names)])
  message("Found ", length(dup_genes), " duplicate gene symbols")
  message("Examples: ", paste(head(dup_genes), collapse = ", "))
  
  # Get counts matrix (sparse format)
  counts_matrix <- GetAssayData(gse146590, layer = "counts")[keep_genes, ]
  
  # FAST AGGREGATION (using sparse matrix operations)
  # Create mapping matrix
  unique_genes <- unique(new_gene_names)
  mapping_matrix <- sparseMatrix(
    i = match(new_gene_names, unique_genes),
    j = seq_along(new_gene_names),
    x = 1,
    dims = c(length(unique_genes), length(new_gene_names))
  )
}
  
# Aggregate counts (matrix multiplication)
aggregated_counts <- mapping_matrix %*% counts_matrix
rownames(aggregated_counts) <- unique_genes
colnames(aggregated_counts) <- colnames(counts_matrix)
  
# Create new assay object
new_assay <- CreateAssayObject(counts = aggregated_counts)
if(exists("aggregated_data")) new_assay@data <- aggregated_data

# Replace the assay
gse146590[["RNA"]] <- new_assay

# 4. Standard Seurat Pipeline -------------------------------------
# Basic QC filtering (adjust thresholds as needed)
gse146590 <- subset(gse146590, 
                   subset = nFeature_RNA > 200 & 
                           nFeature_RNA < 6000 & 
                           percent.mt < 15)

# Normalization and feature selection
gse146590 <- NormalizeData(gse146590)
gse146590 <- FindVariableFeatures(gse146590, selection.method = "vst", nfeatures = 2000)
gse146590 <- ScaleData(gse146590, vars.to.regress = c("nFeature_RNA", "percent.mt"))

# Dimensionality reduction
gse146590 <- RunPCA(gse146590, npcs = 50)
gse146590 <- RunUMAP(gse146590, dims = 1:15)

# Clustering
gse146590 <- FindNeighbors(gse146590, dims = 1:30)
gse146590 <- FindClusters(gse146590, resolution = 1)


# 1. Load 24-gene signature and filter to available genes
gene.min <- read.csv("../02.data/24-gene_signature.csv", 
                    header = TRUE, 
                    row.names = 1, 
                    check.names = FALSE) %>%
  {.[rownames(.) %in% rownames(gse146590), , drop = FALSE]} 

# Check filtered genes
message("Using ", nrow(gene.min), " of 24 genes:")
print(rownames(gene.min))

# 2. Updated risk score function (matrix-safe)
risk_score <- function(data, lasso_coef) {
  genes <- intersect(rownames(data), rownames(lasso_coef))
  expr_matrix <- as.matrix(data[genes, ]) # Ensure matrix format
  score_mat <- sweep(t(expr_matrix), 2, lasso_coef[genes, 1], `*`)
  rowSums(score_mat, na.rm = TRUE)
}

# 3. Calculate scores using updated Seurat syntax
gse146590@meta.data$sc_ARG24 <- risk_score(
  GetAssayData(gse146590, assay = "RNA", layer = "data"),
  gene.min
)

# save 
# saveRDS(gse146590, "../02.data/gse146590_arac_scrna/gse146590.rds")
```

```{r}
# 5. Visualization ------------------------------------------------
# UMAP plot colored by sample origin
gse146590 <- readRDS("../02.data/gse146590_arac_scrna/gse146590.rds")

mycolors <- c('#DCC1DD', '#CCE0F5', '#712820', '#CCC9E6')

umap_sample <- DimPlot(gse146590, reduction = "umap", 
                       group.by = "orig.ident", 
                       cols = mycolors) 

# UMAP colored by ARG24 score
umap_score <- FeaturePlot(gse146590, 
                         features = "sc_ARG24",
                         reduction = "umap",
                         pt.size = 0.5) +
              scale_color_gradient2(low = "blue", 
                                   mid = "lightgrey", 
                                   high = "red",
                                   midpoint = median(gse146590$sc_ARG24)) +
              ggtitle("ARG24 Signature Score") 
umap_score


library(ggplot2)
library(dplyr)
library(ggpubr)

# 1. Compute mean score and p-value
meta <- gse146590@meta.data

mean_scores <- meta %>%
  group_by(orig.ident) %>%
  summarise(mean_ARG24 = mean(sc_ARG24, na.rm = TRUE), .groups = "drop")

# Wilcoxon test between the two groups
pval <- wilcox.test(sc_ARG24 ~ orig.ident, data = meta)$p.value
mean_scores$pval <- pval

# Calculate min and max values for consistent scaling
min_val <- min(mean_scores$mean_ARG24, na.rm = TRUE)
max_val <- max(mean_scores$mean_ARG24, na.rm = TRUE)

# 2. Plot
p <- ggplot(mean_scores, aes(x = orig.ident, y = 0)) +
  geom_point(
    aes(size = mean_ARG24, fill = mean_ARG24),
    shape = 21, color = "black", stroke = 0.6
  ) +
  geom_text(aes(label = paste0("p = ", signif(pval, 2))), 
            x = 1.5, y = 0.01, size = 4.5) +
  scale_fill_gradientn(
      colors = c("#FFEDA0", "#F03B20"),
    name = "Mean ARG24",
    limits = c(min_val, max_val),  # Ensure consistent color mapping
    guide = guide_colorbar(frame.colour = "black")
  ) +
  scale_size_continuous(
    range = c(2, 6), 
    name = "Mean ARG24",
    limits = c(min_val, max_val)   # Ensure size matches color scale
  ) +
  labs(
    x = "",
    y = "",
    title = "Mean ARG24 Scores by Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 12),
    legend.position = "right",
    legend.box = "vertical",
    legend.spacing = unit(0.5, "cm")
  ) +
  guides(
    fill = guide_colorbar(order = 1),
    size = guide_legend(order = 2)
  )

# Save and display
p

plots <- umap_sample + umap_score + p

ggsave("../03.out/figure2/ARG24_dot_mean_colored_sized.pdf", plots, width = 11, height = 3)

ggsave("../03.out/figure2/ARG24_dot_mean.pdf", p, width = 4, height = 3)

```

## Resistant K562 cells

```{r}
library(ggpubr)
# Load data
dat <- read.csv("../02.data/gse131823_tpm_k562_resistant.csv", header = TRUE, row.names = 1)
gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE, row.names = 1)

# Log2 transform TPM values (add pseudocount to avoid -Inf)
log2tpm <- log2(dat + 1)

# Create sample groups
sample_groups <- ifelse(grepl("_R", colnames(log2tpm)), "Resistant", "Sensitive")

# Filter to signature genes available in dataset
sig_genes <- intersect(rownames(gene.min), rownames(log2tpm))
gene.min_filtered <- gene.min[sig_genes, , drop = FALSE]

message("Using ", length(sig_genes), " of 24 signature genes:")
print(sig_genes)

# Calculate scores (weighted sum)
calculate_arg24 <- function(expr_mat, coef_df) {
  genes <- intersect(rownames(expr_mat), rownames(coef_df))
  expr_mat <- as.matrix(expr_mat[genes, ])
  scores <- colSums(expr_mat * coef_df[genes, 1], na.rm = TRUE)
  return(scores)
}

arg24_scores <- calculate_arg24(log2tpm, gene.min_filtered)

# 6. Create results DataFrame
results <- data.frame(
  Sample = names(arg24_scores),
  ARG24_Score = arg24_scores,
  Group = sample_groups
)

# 7. Add pair information based on sample naming convention
results <- results %>%
  mutate(
    Pair = str_extract(Sample, "K562_[SR]\\d+"),
    Pair = str_replace(Pair, "K562_[SR]", "Pair_")
  )

# Enhanced paired dot plot
paired_plot <- ggpaired(
  results,
  x = "Group",
  y = "ARG24_Score",
  id = "Pair",
  color = "Group",
  line.color = "gray60",
  line.size = 0.5,
  point.size = 4,
  palette = c("#0072B2", "#D55E00")  # color-blind friendly blue/red
) +
  labs(
    title = "Paired Comparison of ARG24 Score in Chemoresistance",
    y = "ARG24 Score",
    x = NULL
  ) +
  stat_compare_means(
    paired = FALSE,
    label.y = max(results$ARG24_Score, na.rm = TRUE) * 1.1,
    size = 5
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 12),
    legend.position = "none"
  )

# Display
paired_plot

combined_plots <- (umap_sample | umap_score | p | paired_plot) +
  plot_layout(nrow = 1, widths = c(1.25, 1.25, 1, 0.75))  

ggsave("../03.out/figure2/ARG24_paired_comparisons_K562_Arac.pdf", combined_plots, width = 20, height = 4)

```

## DFS and EFS analysis of ARG24

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(survival)
library(survminer)
library(pROC)
library(corrplot)
library(RColorBrewer)

dat <- read.csv("../02.data/4_RJAML_24Gene_Clinical_tpm.csv", header = TRUE, check.names = FALSE)

# Initialize empty data frame for statistics
stat <- data.frame()

# Calculate statistics for each gene
for (i in 7:ncol(dat)) {
  tmp <- dat[, c(6, i)]
  tmp_data <- data.frame(matrix(NA, nrow = 1, ncol = 5))
  tmp_data[1, 1] <- colnames(tmp)[2]
  colnames(tmp)[2] <- "Gene"

  # Calculate AUC
  auc_value <- auc(tmp$Response, tmp$Gene)
  tmp_data[1, 2] <- as.numeric(auc_value)

  # Subset data for CR and NCR
  tmp_cr <- subset(tmp, Response == "CR")
  tmp_ncr <- subset(tmp, Response == "NCR")

  # Perform Wilcoxon test
  testout <- wilcox.test(tmp_cr$Gene, tmp_ncr$Gene, alternative = "two.sided")

  # Store statistics
  tmp_data[1, 3:5] <- c(median(tmp_cr$Gene), median(tmp_ncr$Gene), testout$p.value)
  stat <- rbind(tmp_data, stat)
}

# Rename columns and calculate log2 fold change
colnames(stat) <- c("Genes", "AUC", "CR_median", "NCR_median", "P.value")
stat$log2FC <- log2(stat$NCR_median / stat$CR_median)
stat <- stat[order(stat$AUC, decreasing = TRUE), ]
row.names(stat) <- stat[, 1]
stat <- stat[, -1]
# Reorder and prioritize DPP4 gene
stat <- rbind(stat["DPP4", ], stat[!row.names(stat) %in% "DPP4", ])

# Prepare data for corrplot
corr_dat <- data.frame(auc = stat[, "AUC"])
row.names(corr_dat) <- row.names(stat)
p_dat <- data.frame(p.value = stat[, "P.value"])
row.names(p_dat) <- row.names(stat)

# Define color palette
colors <- colorRampPalette(brewer.pal(9, 'YlGn'))
color_continuous <- colors(100)

# Create a correlation plot
# pdf("../03.out/figure2/03.auc_corrplot.pdf", width = 2, height = 6)
corrplot(
  data.matrix(corr_dat),
  method = "circle",
  tl.cex = 1,
  p.mat = data.matrix(p_dat),
  sig.level = 0.05,
  insig = "label_sig",
  pch.col = "black",
  pch.cex = 1,
  is.corr = FALSE,
  tl.col = 'black',
  cl.ratio = 1,
  col = color_continuous,
  col.lim = c(0.55, 0.75)
)
# dev.off()
```

### EFS in RJAML

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(survival)
library(survminer)

dat <- read.csv("../02.data/4_RJAML_24Gene_Clinical_tpm.csv", header = TRUE, check.names = FALSE, row.names = 1)

# Univariate survival analysis for all genes
uni_surv_anal <- function(exprSet, clinicData, dataset) {
  mySurv <- with(clinicData, Surv(Time, Status))

  log_rank_p <- apply(exprSet, 1, function(gene) {
    clinicData$group <- ifelse(gene >= median(gene), 'high', 'low')
    data.survdiff <- survdiff(mySurv ~ group, data = clinicData)
    p.val <- 1 - pchisq(data.survdiff$chisq, length(data.survdiff$n) - 1)
    return(p.val)
  })

  cox_results <- apply(exprSet, 1, function(gene) {
    group <- ifelse(gene >= median(gene), 'high', 'low')
    survival_dat <- data.frame(group = group, stringsAsFactors = FALSE)
    m <- coxph(mySurv ~ group, data = survival_dat)

    beta <- -coef(m)
    se <- sqrt(diag(vcov(m)))
    HR <- exp(beta)
    HRse <- HR * se

    tmp <- round(cbind(
      coef = beta,
      z = beta / se,
      HR = HR,
      HRse = HRse,
      HRz = (HR - 1) / HRse,
      HRp = 1 - pchisq(((HR - 1) / HRse)^2, 1),
      Lower = exp(beta - qnorm(.975, 0, 1) * se),
      Upper = exp(beta + qnorm(.975, 0, 1) * se)
    ), 3)
    return(tmp['grouplow', ])
  })

  cox_results <- t(cox_results)
  cox_results <- as.data.frame(cox_results)
  log_rank_p <- as.data.frame(log_rank_p)
  cox_results$logrank.P.value <- log_rank_p$log_rank_p
  write.csv(cox_results, paste0(dataset, "_Uni.Cox.out.csv"))
}

mat <- dat[, 6:29]
mat <- data.frame(t(mat))
clinicData <- dat[, c("EFS_Time", "EFS_Status")]
colnames(clinicData) <- c("Time", "Status")

uni_surv_anal(mat, clinicData, "../03.out/figure2/01.RJAML_EFS")

cox_result <- read.csv("../03.out/figure2/01.RJAML_EFS_Uni.Cox.out.csv", header = TRUE, row.names = 1, check.names = FALSE)
cox_result <- cox_result[order(row.names(cox_result)), ]
```

### DFS in TCGA

```{r, eval=TRUE, message = FALSE, warning = FALSE}
dat <- read.csv("../02.data/4_TCGA_24Gene_Clinical_tpm.csv", header = TRUE, check.names = FALSE, row.names = 1)
mat <- dat[, 5:28]
mat <- data.frame(t(mat))
clinicData <- dat[, c("DFS_time", "DFS_status")]
colnames(clinicData) <- c("Time", "Status")

uni_surv_anal(mat, clinicData, "../03.out/figure2/02.TCGA_DFS")

cox_result1 <- read.csv("../03.out/figure2/02.TCGA_DFS_Uni.Cox.out.csv", header = TRUE, row.names = 1, check.names = FALSE)
cox_result1 <- cox_result1[order(row.names(cox_result1)), ]
```

```{r}
# Prepare data for EFS vs DFS corrplot
corr_dat <- cbind(cox_result[, "HR"], cox_result1[, "HR"])
row.names(corr_dat) <- row.names(cox_result)
p_dat <- data.frame(cox_result[, "logrank.P.value"], cox_result1[, "logrank.P.value"])
row.names(p_dat) <- row.names(cox_result)

# Reorder rows to match previous ordering
corr_dat <- corr_dat[row.names(stat), ]
p_dat <- p_dat[row.names(stat), ]

colors <- colorRampPalette(brewer.pal(11, 'RdBu'))
color_continuous <- rev(colors(100))

corrplot(
  data.matrix(corr_dat),
  method = "circle",
  tl.cex = 1,
  p.mat = data.matrix(p_dat),
  sig.level = 0.05,
  insig = "label_sig",
  pch.col = "black",
  pch.cex = 1,
  is.corr = FALSE,
  tl.col = 'black',
  cl.ratio = 1,
  col = color_continuous,
  col.lim = c(0.4, 2.2)
)
```

## 2.11 Select the DPP4 for Further Validation

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(ggplot2)
library(ggpubr)
library(ggrastr)
library(RColorBrewer)
library(tidyverse)

# Load and transform DPP4 expression data
rj_dat <- read.csv("../02.data/4_RJAML_24Gene_Clinical_tpm.csv", header = TRUE, check.names = FALSE) %>%
  mutate(DPP4 = log2(DPP4 + 1))

# Generate violin plot
p <- ggplot(rj_dat, aes(x = Response, y = DPP4, color = Response)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.4, alpha = 0.75, outlier.shape = NA) +
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  xlab("") +
  ylab("DPP4 expression (log2-transformed)") +
  scale_color_manual(values = c("CR" = "#1B9E77", "NCR" = "#7570B3")) +
  theme_pubr(base_size = 12) +
  theme(
    legend.position = "none",
    strip.text.x = element_text(size = 13),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1),
    text = element_text(size = 12)
  ) +
  stat_compare_means(comparisons = list(c("CR", "NCR")), method = 'wilcox.test')

p
# ggsave(p, filename = "../03.out/figure2/04.DPP4_violin.pdf", width = 3, height = 3)
```

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(survival)
library(survminer)
library(tidyverse)
library(patchwork)

# --------- RJAML EFS Survival Analysis ---------
# Load and preprocess data
data_rjaml <- read.csv("../02.data/4_RJAML_24Gene_Clinical_tpm.csv", header = TRUE, check.names = FALSE, row.names = 1)
dat1 <- data_rjaml[, c("EFS_Time", "EFS_Status", "DPP4")]
colnames(dat1) <- c("Time", "Status", "DPP4")
dat1$group <- ifelse(dat1$DPP4 >= median(dat1$DPP4, na.rm = TRUE), "High", "Low")
dat1$group <- factor(dat1$group, levels = c("High", "Low"))

# Survival stats
surv_diff1 <- survdiff(Surv(Time, Status) ~ group, data = dat1)
p.val1 <- 1 - pchisq(surv_diff1$chisq, length(surv_diff1$n) - 1)
HR1 <- (surv_diff1$obs[1] / surv_diff1$exp[1]) / (surv_diff1$obs[2] / surv_diff1$exp[2])
CI_lower1 <- exp(log(HR1) - qnorm(0.975) * sqrt(1 / surv_diff1$exp[2] + 1 / surv_diff1$exp[1]))
CI_upper1 <- exp(log(HR1) + qnorm(0.975) * sqrt(1 / surv_diff1$exp[2] + 1 / surv_diff1$exp[1]))

# Plot
p1 <- ggsurvplot(
  survfit(Surv(Time, Status) ~ group, data = dat1),
  palette = "npg",
  pval = paste0(
    ifelse(p.val1 < 0.001, "p < 0.001", paste0("p = ", round(p.val1, 4))),
    "\nHazard Ratio = ", round(HR1, 2),
    "\n95% CI: ", round(CI_lower1, 2), " - ", round(CI_upper1, 2)
  ),
  pval.size = 5,
  xlab = "Survival time in days",
  conf.int = FALSE,
  combine = TRUE,
  risk.table = FALSE,
  title = "RJAML EFS",
  legend.title = "DPP4 score",
  legend.labs = c(
    paste0("High (N = ", table(dat1$group)["High"], ")"),
    paste0("Low  (N = ", table(dat1$group)["Low"], ")")
  ),
  legend = c(0.75, 0.85)
)

# --------- TCGA DFS Survival Analysis ---------
# Load and preprocess data
data_tcga <- read.csv("../02.data/4_TCGA_24Gene_Clinical_tpm.csv", header = TRUE, check.names = FALSE, row.names = 1)
dat2 <- data_tcga[, c("DFS_time", "DFS_status", "DPP4")]
colnames(dat2) <- c("Time", "Status", "DPP4")
dat2$group <- ifelse(dat2$DPP4 >= median(dat2$DPP4, na.rm = TRUE), "High", "Low")
dat2$group <- factor(dat2$group, levels = c("High", "Low"))

# Survival stats
surv_diff2 <- survdiff(Surv(Time, Status) ~ group, data = dat2)
p.val2 <- 1 - pchisq(surv_diff2$chisq, length(surv_diff2$n) - 1)
HR2 <- (surv_diff2$obs[1] / surv_diff2$exp[1]) / (surv_diff2$obs[2] / surv_diff2$exp[2])
CI_lower2 <- exp(log(HR2) - qnorm(0.975) * sqrt(1 / surv_diff2$exp[2] + 1 / surv_diff2$exp[1]))
CI_upper2 <- exp(log(HR2) + qnorm(0.975) * sqrt(1 / surv_diff2$exp[2] + 1 / surv_diff2$exp[1]))

# Plot
p2 <- ggsurvplot(
  survfit(Surv(Time, Status) ~ group, data = dat2),
  palette = "npg",
  pval = paste0(
    ifelse(p.val2 < 0.001, "p < 0.001", paste0("p = ", round(p.val2, 4))),
    "\nHazard Ratio = ", round(HR2, 2),
    "\n95% CI: ", round(CI_lower2, 2), " - ", round(CI_upper2, 2)
  ),
  pval.size = 5,
  xlab = "Survival time in days",
  conf.int = FALSE,
  combine = TRUE,
  risk.table = FALSE,
  title = "TCGA DFS",
  legend.title = "DPP4 score",
  legend.labs = c(
    paste0("High (N = ", table(dat2$group)["High"], ")"),
    paste0("Low  (N = ", table(dat2$group)["Low"], ")")
  ),
  legend = c(0.75, 0.85)
)

# Arrange plots
combined_plot <- arrange_ggsurvplots(list(p1, p2), print = FALSE, ncol = 1, nrow = 2)
combined_plot

# Save to PDF (optional)
# ggsave("../03.out/figure6/13.DPP4_survival_combined.pdf", combined_plot, width = 5.5, height = 8)
```

## 2.12 DE Analysis for RNA-seq with DPP4 Knockdown

```{r, eval=TRUE, message = FALSE, warning = FALSE}
# library(DESeq2)
# source("../01.script/RunDESeq2.R")
# 
# # Load TPM matrix and filter by gene biotypes
raw_tpm <- read.table("../02.data/shDPP4_tpm_matrix.txt", header = TRUE, check.names = FALSE)
tpm <- subset(raw_tpm, Type %in% c(
  "protein_coding", "miRNA", "lincRNA", "antisense", "processed_transcript",
  "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene",
  "transcribed_unprocessed_pseudogene", "macro_lncRNA", "TEC",
  "3prime_overlapping_ncRNA", "sense_intronic", "sense_overlapping", "unprocessed_pseudogene"
))

# Filter out lowly expressed genes
expr_filter <- rowMeans(tpm[, 4:ncol(tpm)]) >= 0.5
tpm <- tpm[expr_filter, ]
# 
# # Load count matrix and match filtered genes
# mat <- read.table("../02.data/shDPP4_count_matrix.txt", header = TRUE, row.names = 1, check.names = FALSE)
# mat <- mat[tpm$ID, -c(1:2)]
# 
# # Differential expression analysis
# ## shDPP4-1 vs Control
# RunDESeq2(
#   count_mat = mat[, 1:6], n.cont = 3, n.treat = 3,
#   prefix = "../02.data/shDPP4_1_DESeq2_out", sort.p = FALSE,
#   merge.normalized = TRUE, normalized_mat = tpm[, 1:9]
# )
# 
# ## shDPP4-2 vs Control
# RunDESeq2(
#   count_mat = mat[, c(1:3, 7:9)], n.cont = 3, n.treat = 3,
#   prefix = "../02.data/shDPP4_2_DESeq2_out", sort.p = FALSE,
#   merge.normalized = TRUE, normalized_mat = tpm[, c(1:6, 10:12)]
# )
# 
# ## Combined shDPP4 vs Control
# RunDESeq2(
#   count_mat = mat, n.cont = 3, n.treat = 6,
#   prefix = "../02.data/shDPP4_combined_DESeq2_out", sort.p = FALSE,
#   merge.normalized = TRUE, normalized_mat = tpm
# )
```

## 2.13 PCA Analysis

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(ggfortify)
library(scatterplot3d)

# Prepare scaled matrix
dat <- tpm[, 4:12]
rownames(dat) <- tpm$ID
scaled_matrix <- scale(t(dat))
pca_result <- prcomp(scaled_matrix, scale. = TRUE)

# Define groups and colors
groups <- c(rep("Scramble", 3), rep("shDPP4-1", 3), rep("shDPP4-2", 3))
pca_3d_df <- as.data.frame(pca_result$x)
pca_3d_df$Group <- groups
group_colors <- c("Scramble" = "red", "shDPP4-1" = "green", "shDPP4-2" = "blue")

# Labels with variance explained
expl_var <- summary(pca_result)$importance[2, ]
xlab <- paste0("PC1 (", round(expl_var[1] * 100, 2), "%)")
ylab <- paste0("PC2 (", round(expl_var[2] * 100, 2), "%)")
zlab <- paste0("PC3 (", round(expl_var[3] * 100, 2), "%)")

# pdf("../03.out/figure2/05.shDPP4_3D_PCA.pdf", width = 6, height = 6)
scatterplot3d(
  pca_3d_df$PC1, pca_3d_df$PC2, pca_3d_df$PC3,
  color = group_colors[pca_3d_df$Group],
  pch = 19, angle = 55, grid = TRUE, box = TRUE,
  xlab = xlab, ylab = ylab, zlab = zlab, main = "3D PCA of TPM Matrix"
)
legend("top", legend = names(group_colors), col = group_colors, pch = 19, title = "Groups")
# dev.off()
```

## 2.14 Volcano Plots

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(patchwork)
source("plotVolcano.R")
selected_genes <- read.csv("../02.data/shDPP4_input_selected.csv", header = TRUE)

# Load DE results
deseq1 <- read.table("../02.data/shDPP4_1_DESeq2_out_with_normalized_mat.txt", header = TRUE)
deseq2 <- read.table("../02.data/shDPP4_2_DESeq2_out_with_normalized_mat.txt", header = TRUE)

# Preprocess DE results
deseq1 <- deseq1[, 1:9]
deseq1$log2FoldChange <- pmin(deseq1$log2FoldChange, 5)
deseq1$padj[deseq1$padj < 1e-60] <- 1e-60

deseq2 <- deseq2[, 1:9]
deseq2$log2FoldChange <- pmax(pmin(deseq2$log2FoldChange, 5), -5)
deseq2$padj[deseq2$padj < 1e-60] <- 1e-60

# Plot volcano plots
custom_colors <- brewer.pal(4, "BrBG")
vp1 <- plotVolcano(
  data = deseq1, gene.col = "Symbol", logFC.col = "log2FoldChange", pval.col = "padj",
  selected_genes = selected_genes, logFCcut = 0.5, logFCcut2 = 1, logFCcut3 = 2,
  pvalCut = 0.05, pvalCut2 = 0.0001, pvalCut3 = 0.00001,
  threshold_colors = custom_colors, plot_mode = "advanced",
  x_range = c(-5, 5), y_range = c(0, 60), title = "shDPP4-1 vs. Scramble"
)

vp2 <- plotVolcano(
  data = deseq2, gene.col = "Symbol", logFC.col = "log2FoldChange", pval.col = "padj",
  selected_genes = selected_genes, logFCcut = 0.5, logFCcut2 = 1, logFCcut3 = 2,
  pvalCut = 0.05, pvalCut2 = 0.0001, pvalCut3 = 0.00001,
  threshold_colors = custom_colors, plot_mode = "advanced",
  x_range = c(-5, 5), y_range = c(0, 60), title = "shDPP4-2 vs. Scramble"
)

combined_plot <- vp1 / vp2

combined_plot
# ggsave("../03.out/figure2/06.volcano_plot_with_pathways.pdf", plot = combined_plot, width = 18, height = 5.8)
```

## 2.15 GSEA Analysis

```{r, eval=TRUE, message = FALSE, warning = FALSE}
library(clusterProfiler)
library(enrichplot)
library(GseaVis)
library(tidyverse)

# Prepare gene list
deg <- read.table("../02.data/shDPP4_combined_DESeq2_out_with_normalized_mat.txt", header = TRUE)
geneList <- deg %>%
  select(Symbol, log2FoldChange) %>%
  distinct(Symbol, .keep_all = TRUE) %>%
  drop_na() %>%
  arrange(desc(log2FoldChange))

gene_list <- geneList$log2FoldChange
names(gene_list) <- geneList$Symbol

# Load gene sets
gmt <- read.gmt("../02.data/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

# Run GSEA
set.seed(123456)
gsea_result <- GSEA(
  gene_list, TERM2GENE = gmt, pvalueCutoff = 1, nPermSimple = 10000,
  maxGSSize = 800, pAdjustMethod = "BH", seed = TRUE
)

# Plot top pathways
gsea_result@result$Description <- gsub("KEGG_", "", gsea_result@result$Description)

# pdf("../03.out/figure2/07.shDPP4_GSEA_up.pdf", width = 6, height = 4)
gseaplot2(
  gsea_result, subplots = 1:2, pvalue_table = TRUE,
  geneSetID = c("KEGG_RIBOSOME", "KEGG_OXIDATIVE_PHOSPHORYLATION", "KEGG_DNA_REPLICATION"),
  title = "Pathways activated by DPP4 knockdown"
)
# dev.off()

# pdf("../03.out/figure2/07.shDPP4_GSEA_down.pdf", width = 6, height = 4)
gseaplot2(
  gsea_result, subplots = 1:2, pvalue_table = TRUE,
  geneSetID = c("KEGG_JAK_STAT_SIGNALING_PATHWAY", "KEGG_CHEMOKINE_SIGNALING_PATHWAY"),
  title = "Pathways repressed by DPP4 knockdown"
)
# dev.off()
```

