# Figure 3: Validation of the ARG24 Signature

To independently validate the association of the ARG24 score with patient response to chemotherapy at diagnosis, we integrated the transcriptomic and clinical data from six cohorts across different analysis platforms, consisting of 1,512 patients. The bulk RNA-seq cohorts (BeatAML, Leucegene, TCVGH, GSE165656 and GSE146173) and microarray cohort (HOVON). We calculated the ARG24 scores using each patients' expression profiles. 

## ARG24 Score Validation in Bulk RNA-Seq Cohorts

```{r fig3_bulk_validation, eval=TRUE, message=FALSE, warning=FALSE}
library(pROC)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)

# Load validation dataset
bulk_data <- read.csv("../02.data/24-Gene_MultiCohorts_riskScore.csv", header = TRUE)
bulk_data <- subset(bulk_data, center != "RJAML")
bulk_data$center <- factor(bulk_data$center, levels = c("BeatAML", "Leucegene", "TCVGH", "GSE165656", "GSE146173", "HOVON"))

# Plot violin + box + raster
p <- ggplot(bulk_data, aes(x = group, y = risk_score, color = group)) + 
  geom_violin() +
  geom_boxplot_jitter(width = 0.4, alpha = 0.75) + 
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  facet_wrap(. ~ center, ncol = 3, scales = 'free') + 
  theme_pubr() + 
  xlab("") + ylab("ARG24 Score") +
  stat_compare_means(comparisons = list(c("CR", "NCR")), method = "wilcox.test") + 
  scale_color_manual(values = c("#4292C6", "#A50F15")) +
  theme(strip.text.x = element_text(size = 13), legend.position = "none")
p

# ggsave("../03.out/figure3/01.bulk_violin.pdf", p, width = 10, height = 4.2)

## validation datasets
data <- bulk_data[, 1:4]
beat_dat <- subset(data, center == "BeatAML")
leu_dat <- subset(data, center == "Leucegene")
gse146173_dat <- subset(data, center == "GSE146173")
tcv_dat <- subset(data, center == "TCVGH")
gse165656_dat <- subset(data, center == "GSE165656")
hovon_dat <- subset(data, center == "HOVON")

auc3 <- auc(beat_dat$group, beat_dat$risk_score)
roc3 <- roc(beat_dat$group, beat_dat$risk_score)
auc4 <- auc(leu_dat$group, leu_dat$risk_score)
roc4 <- roc(leu_dat$group, leu_dat$risk_score)
auc5 <- auc(tcv_dat$group, tcv_dat$risk_score)
roc5 <- roc(tcv_dat$group, tcv_dat$risk_score)
auc6 <- auc(gse165656_dat$group, gse165656_dat$risk_score)
roc6 <- roc(gse165656_dat$group, gse165656_dat$risk_score)
auc7 <- auc(gse146173_dat$group, gse146173_dat$risk_score)
roc7 <- roc(gse146173_dat$group, gse146173_dat$risk_score)
auc8 <- auc(hovon_dat$group, hovon_dat$risk_score)
roc8 <- roc(hovon_dat$group, hovon_dat$risk_score)

plot(smooth(roc3),  col=brewer.pal(8, "Dark2")[1])
plot(smooth(roc4), add=TRUE, col=brewer.pal(8, "Dark2")[2])
plot(smooth(roc5), add=TRUE, col=brewer.pal(8, "Dark2")[3])
plot(smooth(roc6), add=TRUE, col=brewer.pal(8, "Dark2")[4])
plot(smooth(roc7), add=TRUE, col=brewer.pal(8, "Dark2")[5])
plot(smooth(roc8), add=TRUE, col=brewer.pal(8, "Dark2")[6])
legend("bottomright", 
       legend = c(paste0("BeatAML AUC=", round(auc3, 3)),
                  paste0("Leucegene AUC=", round(auc4, 3)),
                  paste0("TCVGH AUC=", round(auc5, 3)),
                  paste0("GSE165656 AUC=", round(auc6, 3)),
                  paste0("GSE146173 AUC=", round(auc7, 3)),
                  paste0("HOVON AUC=", round(auc8, 3))),
       col = c(brewer.pal(8, "Dark2")[1:6]),
       lty = 1)
title(main = "Validation datasets")
```

## Integrate single cell datasets

```{r}
# ==============================================
# AML Single-Cell Data Integration Pipeline
# Description: Subsets and merges multiple AML scRNA-seq datasets
# ==============================================

library(Seurat)
library(dplyr)
library(future)
library(purrr) 

# Enable parallel processing if needed
plan("multicore", workers = 8)
options(future.globals.maxSize = 8000 * 1024^2) # 8GB

# ======================
# Data Loading Functions
# ======================

load_and_subset <- function(path, target_samples, id_col = "orig.ident") {
  #' Load and subset Seurat object
  #'
  #' @param path Path to RDS file
  #' @param target_samples Vector of sample IDs to keep
  #' @param id_col Column containing sample IDs
  #' @return Subsetted Seurat object
  
  message("\nLoading: ", basename(path))
  obj <- tryCatch(
    {
      sobj <- readRDS(path)
      
      # Handle special cases for orig.ident
      if (grepl("GSE227903", path)) {
        sobj$orig.ident <- sobj$Sample
      } else if (grepl("GSE239721", path)) {
        sobj$orig.ident <- paste(sobj$patient, sobj$day, sep = "_")
      }
      
      # Subset the object
      cells_keep <- colnames(sobj)[sobj[[id_col]][[1]] %in% target_samples]
      if (length(cells_keep) == 0) {
        stop("No cells matched the target samples")
      }
      
      subset(sobj, cells = cells_keep)
    },
    error = function(e) {
      message("Error processing ", path, ": ", e$message)
      return(NULL)
    }
  )
  
  if (!is.null(obj)) {
    message("Retained ", ncol(obj), " cells from ", 
            length(unique(obj$orig.ident)), " samples")
  }
  
  return(obj)
}

# ====================
# Dataset Configuration
# ====================

dataset_config <- list(
  "GSE116256" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/AML_plasticity/jp_2024/scAML_GSE116256_30712Cells.rds",
    samples = c("AML314-D0", "AML329-D0", "AML371-D0", "AML420B-D0", "AML722B-D0",
                "AML475-D0", "AML556-D0", "AML707B-D0", "AML870-D0", "AML997-D0")
  ),
  "GSE185991" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/AML_plasticity/jp_2024/scAML_GSE185991_86446Cells.rds",
    samples = c("PT01_DX", "PT02_DX", "PT06_DX", "PT07_DX", "PT08_DX",
                "PT09_DX", "PT10_DX", "PT11_DX", "PT12_DX", "PT13_DX",
                "PT15_DX", "PT17_DX", "PT18_DX")
  ),
  "GSE227903" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/AML_plasticity/jp_2024/scAML_GSE227903_111333Cells.rds",
    samples = c("1216_Diagnosis", "1886_Diagnosis", "3822_Diagnosis", "3853_Diagnosis", 
                "3904_Diagnosis", "40389_Diagnosis", "4978_Diagnosis", "5143_Diagnosis",
                "5750_Diagnosis", "6323_Diagnosis")
  ),
  "GSE239721" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/AML_plasticity/jp_2024/GSE239721_101767Cells.rds",
    samples = c("A.2_d0", "A.3_d0", "A.4_d0", "A.5_d0", "A.6_d0", "A.7_d0",
                "A.8_d0", "A.9_d0", "A.10_d0", "A.11_d0", "A.12_d0", "A.13_d0",
                "A.14_d0", "A.15_d0", "B.2_d0", "B.4_d0")
  ),
  "GSE241989" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/AML_plasticity/jp_2024/scAML_GSE241989_103690Cells.rds",
    samples = c("03H112", "06H088", "09H010", "09H032",
                "05H034", "05H066", "07H134", "14H007")
  ),
  "GSE272758" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/jp_projects/neoantigen/rjaml_7samples_61902_cells.rds",
    samples = c("R0132", "R0133", "R0134", "R0135")
  ),
  "GSE240618" = list(
    path = "/dssg/home/acct-medljm/medljm-jqq/projects/AML_plasticity/jp_2024/scAML_GSE240618.rds",
    samples = c("AML1",	"AML2",	"AML3",	"AML4",	"AML5",	"AML6",	"AML7",	"AML8",	"AML9",	"AML10", "AML11",	
                "AML12",	"AML13",	"AML14",	"AML15",	"AML16",	"AML17",	"AML18", "AML19",	"AML20") 
  )
)

# ====================
# Data Processing
# ====================

# Load and subset all datasets
seurat_list <- lapply(names(dataset_config), function(dataset) {
  config <- dataset_config[[dataset]]
  if (!is.null(config$samples)) {
    load_and_subset(config$path, config$samples)
  } else {
    readRDS(config$path)
  }
}) %>% 
  setNames(names(dataset_config)) %>% 
  compact() # Remove NULL elements

# =============
# Data Merging
# =============

# Add dataset origin metadata
for (dataset in names(seurat_list)) {
  seurat_list[[dataset]]$Dataset <- dataset
}

# Merge all datasets
message("\nMerging datasets...")
merged_aml <- merge(
  x = seurat_list[[1]],
  y = seurat_list[-1],
  add.cell.ids = names(seurat_list),
  project = "AML_Merged"
)

# Clean up metadata
merged_aml$orig.ident <- factor(merged_aml$orig.ident)
merged_aml$Dataset <- factor(merged_aml$Dataset)

# Create a named vector of response status
response_data <- c(
  # GSE116256 samples
  "AML314-D0" = "CR", "AML329-D0" = "CR", "AML371-D0" = "CR", 
  "AML420B-D0" = "CR", "AML475-D0" = "CR", "AML556-D0" = "CR",
  "AML707B-D0" = "NCR", "AML870-D0" = "CR", "AML997-D0" = "CR",
  "AML722B-D0" = "CR",
  
  # GSE185991 samples
  "PT01_DX" = "NCR", "PT02_DX" = "NCR", "PT06_DX" = "CR",
  "PT07_DX" = "CR", "PT08_DX" = "CR", "PT09_DX" = "CR",
  "PT10_DX" = "CR", "PT11_DX" = "NCR", "PT12_DX" = "CR",
  "PT13_DX" = "NCR", "PT15_DX" = "CR", "PT17_DX" = "NCR", 
  "PT18_DX" = "CR",
  
  # GSE227903 samples
  "1216_Diagnosis" = "CR", "1886_Diagnosis" = "CR",
  "3822_Diagnosis" = "CR", "3853_Diagnosis" = "CR",
  "3904_Diagnosis" = "CR", "40389_Diagnosis" = "CR",
  "4978_Diagnosis" = "CR", "5143_Diagnosis" = "CR",
  "5750_Diagnosis" = "CR", "6323_Diagnosis" = "CR",
  
  # GSE239721 samples
  "A.2_d0" = "NCR", "A.3_d0" = "NCR", "A.4_d0" = "NCR",
  "A.5_d0" = "NCR", "A.6_d0" = "CR", "A.7_d0" = "CR",
  "A.8_d0" = "NCR", "A.9_d0" = "CR", "A.10_d0" = "CR",
  "A.11_d0" = "CR", "A.12_d0" = "CR", "A.13_d0" = "CR",
  "A.14_d0" = "NCR", "A.15_d0" = "NCR", "B.2_d0" = "NCR", 
  "B.4_d0" = "CR",
  
  # GSE241989 samples
  "03H112" = "CR", "06H088" = "CR", "09H010" = "CR",
  "09H032" = "CR", "05H034" = "NCR", "05H066" = "NCR",
  "07H134" = "NCR", "14H007" = "NCR",
  
  # GSE272758 samples
  "R0132" = "CR", "R0133" = "CR", "R0134" = "CR", 
  "R0135" = "NCR",
  
  # GSE240618 samples
  "AML1" = "NCR", "AML2" = "NCR", "AML3" = "NCR",
  "AML4" = "NCR", "AML5" = "NCR", "AML6" = "CR",
  "AML7" = "CR", "AML8" = "CR", "AML9" = "CR",
  "AML10" = "CR", "AML11" = "CR", "AML12" = "CR",
  "AML13" = "CR", "AML14" = "CR", "AML15" = "CR",
  "AML16" = "CR", "AML17" = "CR", "AML18" = "CR",
  "AML19" = "CR", "AML20" = "CR"
)

# Convert to data frame
response_df <- data.frame(
  orig.ident = names(response_data),
  Response = unname(response_data),
  row.names = NULL
)

# Create response vector with proper names
response_vec <- setNames(response_df$Response, response_df$orig.ident)

# Check coverage
coverage <- mean(unique(merged_aml$orig.ident) %in% names(response_vec))
message(sprintf("Response data covers %.1f%% of samples", coverage*100))

# Direct assignment (most reliable)
merged_aml$Response <- ifelse(
  merged_aml$orig.ident %in% names(response_vec),
  response_vec[as.character(merged_aml$orig.ident)],
  "Unknown"  # Default for unmatched samples
)

merged_aml@meta.data <- merged_aml@meta.data[, c("orig.ident", "nCount_RNA", "nFeature_RNA", "percent.mt", "Dataset", "Response")]

# =============
# Quality Report
# =============

message("\nFinal merged object summary:")
print(merged_aml)
message("\nCell counts by dataset:")
print(table(merged_aml$Dataset))
message("\nCell counts by sample:")
print(table(merged_aml$orig.ident))

# =============
# Save Results
# =============

output_path <- "/dssg/home/acct-medljm/medljm-jqq/projects/jp_projects/chemoSig/02.data/merged_scaml_response.rds"
saveRDS(merged_aml, file = output_path)
message("\nSaved merged object to: ", output_path)

```

### Single-cell RNA-seq Reference Mapping

```{r fig3_scrna_refmap, eval=TRUE, message=FALSE, warning=FALSE}
library(Seurat)
library(dplyr)
library(ggplot2)
library(BoneMarrowMap)

# Load Symphony reference
ref <- readRDS("../02.data/BM_reference/BoneMarrow_RefMap_SymphonyRef.rds")
# Set uwot path for UMAP projection
ref$save_uwot_path <- "../02.data/BM_reference/BoneMarrow_RefMap_uwot_model.uwot"

# Visualize Bone Marrow reference
ReferenceSeuratObj <- create_ReferenceObject(ref)

###########################################################
# Map Query data and evaluate mapping QC metrics (< 1min) #
###########################################################
query <- readRDS("../02.data/merged_scaml_response.rds")

# batch variable to correct in the query data, set as NULL if no batches in query
batchvar <- 'orig.ident'

# Map query dataset using Symphony (Kang et al 2021)
query <- map_Query(
    exp_query = query[['RNA']]@counts, 
    metadata_query = query@meta.data,
    ref_obj = ref,
    vars = batchvar
)

# Run QC based on mapping error score, flag cells with mapping error >= 2.5 MADs above median
query <- query %>% calculate_MappingError(., reference = ref, MAD_threshold = 2.5, 
                                          threshold_by_donor = TRUE, donor_key = batchvar)

QC_plots <- plot_MappingErrorQC(query)

# Plot together - If this is too crowded, can also just call "QC_plots" aloneto display one by one
p <- patchwork::wrap_plots(QC_plots, ncol = 4, widths = c(0.8, 0.3, 0.8, 0.3))

# Predict Hematopoietic Cell Types by KNN classification
query <- predict_CellTypes(
  query_obj = query, 
  ref_obj = ref, 
  initial_label = 'initial_CellType', # celltype assignments before filtering on mapping QC
  final_label = 'predicted_CellType'  # celltype assignments with map QC failing cells assigned as NA
) 

# Broad cell type mapping 
celltype_mapping <- c(
  "HSC" = "HSC MPP",
  "MPP-MkEry" = "HSC MPP",
  "MPP-MyLy" = "HSC MPP",
  "MEP" = "MEP",
  "BFU-E" = "MEP",
  "Megakaryocyte Precursor" = "Megakaryocyte Precursor",
  "Megakaryocyte" = "Megakaryocyte Precursor",
  "EoBasoMast Precursor" = "EoBasoMast Precursor",
  "CFU-E" = "Early Erythroid",
  "Pro-Erythroblast" = "Early Erythroid",
  "Basophilic Erythroblast" = "Early Erythroid",
  "Orthochromatic Erythroblast" = "Late Erythroid",
  "Polychromatic Erythroblast" = "Late Erythroid",
  "LMPP" = "LMPP",
  "Cycling Progenitor" = "Cycling Progenitor",
  "Early GMP" = "Early GMP",
  "GMP-Cycle" = "Late GMP",
  "GMP-Neut" = "Late GMP",
  "GMP-Mono" = "Late GMP",
  "Early ProMono" = "Pro-Monocyte",
  "Late ProMono" = "Pro-Monocyte",
  "CD14 Mono" = "Monocyte",
  "CD16 Mono" = "Monocyte",
  "Pre-cDC" = "cDC",
  "cDC2" = "cDC",
  "cDC1" = "cDC",
  "ASDC" = "cDC",
  "MLP" = "Early Lymphoid",
  "CLP" = "Early Lymphoid",
  "Pre-ProB" = "Early Lymphoid",
  "MLP-II" = "Early Lymphoid",
  "Pre-pDC" = "pDC",
  "Pre-pDC Cycling" = "pDC",
  "pDC" = "pDC",
  "Pro-B VDJ" = "Pro-B",
  "Pro-B Cycling" = "Pro-B",
  "Large Pre-B" = "Pre-B",
  "Small Pre-B" = "Pre-B",
  "Immature B" = "B",
  "Mature B" = "B",
  "CD4 Naive" = "Naïve T",
  "CD8 Naive" = "Naïve T",
  "CD4 Regulatory" = "CD4 Memory T",
  "CD4 Central Memory" = "CD4 Memory T",
  "CD4 Effector Memory" = "CD4 Memory T",
  "CD8 Effector Memory 1" = "CD8 Memory T",
  "CD8 Effector Memory 2" = "CD8 Memory T",
  "CD8 Central Memory" = "CD8 Memory T",
  "CD8 Tissue Resident Memory" = "CD8 Memory T",
  "T Proliferating" = "CD8 Memory T",
  "NK" = "NK",
  "NK CD56high" = "NK",
  "NK Proliferating" = "NK",
  "Plasma Cell" = "Plasma Cell",
  "Stromal" = "Stromal"
)

# Then add the new column to your metadata
query@meta.data$predicted_CellType_Broad <- celltype_mapping[as.character(query@meta.data$predicted_CellType)]

# Check if any NAs were introduced (for cell types not in your mapping)
if (any(is.na(query@meta.data$predicted_CellType_Broad))) {
  warning("Some cell types didn't have a mapping. They were assigned NA.")
  print(table(query@meta.data$predicted_CellType[is.na(query@meta.data$predicted_CellType_Broad)]))
}

saveRDS(query, "../02.data/merged_scaml_response.rds")
```

### Single-cell Mapping

```{r fig3_scrna_map_riskscore, eval=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(Seurat)
library(dplyr)

scaml <- readRDS("../02.data/merged_scaml_response.rds")

scaml <- NormalizeData(scaml, normalization.method = "LogNormalize")

# Fail   Pass 
# 12021 273910
  
# Color palette
my36colors <- c('#712820', '#DCC1DD', '#CCE0F5', '#CCC9E6', '#625D9E', '#68A180', '#3A6963', 
                '#E5D2DD', '#53A85F', '#F1BB72', '#F3B1A0', '#D6E7A3', '#57C3F3', '#476D87',
                '#E95C59', '#E59CC4', '#AB3282', '#23452F', '#BD956A', '#8C549C', '#585658', 
                '#9FA3A8', '#E0D4CA', '#5F3D69', '#58A4C3', "#b20000",'#E4C755', '#F7F398',
                '#AA9A59', '#E63863', '#E39A35', '#C1E6F3', '#6778AE', '#91D0BE', '#B53E2B','#968175')

scaml.filter <- subset(scaml, mapping_error_QC == 'Pass')

# DimPlot: AML cells projected onto reference map
p1 <- DimPlot(scaml.filter, group.by = "predicted_CellType_Broad", cols = my36colors, 
              label = FALSE, repel = TRUE, raster = FALSE) 
p2 <- DimPlot(scaml.filter, group.by = "predicted_CellType_Broad", cols = my36colors, 
              split.by = "Response", label = FALSE, repel = TRUE, raster = FALSE) 

# Save plots
pdf("../03.out/figure3/02.AML_mapTo_refMap_cellClass.pdf", width = 7, height = 4.2)
p1
dev.off()

pdf("../03.out/figure3/02.AML_mapTo_refMap_Response.pdf", width = 11, height = 4.2)
p2
dev.off()

```

### Pseudobulk scoring

```{r fig3_scrna_pseudobulk_score, eval=TRUE, message=FALSE, warning=FALSE}
# Process data

library(Seurat)
library(dplyr)
library(Matrix)

# 1. Load filtered data
scaml <- readRDS("../02.data/merged_scaml_response.rds")
scaml.filter <- subset(scaml, mapping_error_QC == 'Pass')

# 2. Load gene signature
gene.min <- read.csv("../02.data/24-gene_signature.csv", 
                    header = TRUE, 
                    row.names = 1, 
                    check.names = FALSE) %>%
  {.[rownames(.) %in% rownames(scaml.filter), , drop = FALSE]}

# 3. Create pseudobulk matrix (unchanged)
pseudobulk_creation <- function(seurat_obj, assay = "RNA", sample_col = "orig.ident") {
  counts <- GetAssayData(seurat_obj, assay = assay, layer = "counts")
  sample_list <- unique(seurat_obj@meta.data[[sample_col]])
  pb_counts <- matrix(NA, nrow = nrow(counts), ncol = length(sample_list),
                     dimnames = list(rownames(counts), sample_list))
  
  for (sample in sample_list) {
    cells <- colnames(seurat_obj)[seurat_obj@meta.data[[sample_col]] == sample]
    pb_counts[, sample] <- if(length(cells) > 1) rowSums(counts[, cells]) else counts[, cells]
  }
  
  log2(edgeR::cpm(pb_counts) + 1)
}

pb_matrix <- pseudobulk_creation(scaml.filter)

# 4. Calculate scores (unchanged)
calculate_pb_score <- function(pb_matrix, lasso_coef) {
  genes <- intersect(rownames(pb_matrix), rownames(lasso_coef))
  colSums(pb_matrix[genes, ] * lasso_coef[genes, 1], na.rm = TRUE)
}

pb_scores <- calculate_pb_score(pb_matrix, gene.min)

# 5. Create results with FIXED response mapping
response_mapping <- scaml.filter@meta.data %>%
  group_by(orig.ident) %>%
  summarise(
    Response = if("Response" %in% colnames(.)) Response[1] else NA_character_,
    .groups = 'drop'
  )

pb_results <- data.frame(
  sample = names(pb_scores),
  ARG24_score = pb_scores,
  n_cells = as.integer(table(scaml.filter@meta.data$orig.ident)[names(pb_scores)]),
  row.names = NULL
) %>%
  left_join(response_mapping, by = c("sample" = "orig.ident")) %>%
  dplyr::select(sample, Response, everything()) %>%
  arrange(desc(ARG24_score))

# 6. Save results
write.csv(pb_results, "../02.data/pseudobulk_ARG24_scores.csv", row.names = FALSE)
```

```{r}
# Visualize results

library(pROC)
library(ggpubr)
library(dplyr)
library(RColorBrewer)

# 1. Load and prepare data
pb_results <- read.csv("../02.data/pseudobulk_ARG24_scores.csv", header = TRUE) %>%
  mutate(
    Response = factor(Response, levels = c("CR", "NCR")), # Ensure consistent ordering
    size_label = ifelse(n_cells > 10000, format(n_cells, big.mark = ","), NA) # Format large numbers
  )

# 2. ARG24 Score by Sample (Bar plot)
p1 <- ggplot(pb_results, aes(x = reorder(sample, ARG24_score), y = ARG24_score)) +
  geom_col(aes(fill = ARG24_score)) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "ARG24 Score") +
  labs(
    x = "Sample",
    y = "Pseudobulk ARG24 Score",
    title = "ARG24 Score Distribution Across Samples"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major.x = element_blank()
  )

# 3. Response Comparison (Violin + Boxplot)
p2 <- ggplot(pb_results, aes(x = Response, y = ARG24_score, color = Response)) + 
  geom_violin() +
  geom_boxplot(width = 0.4, alpha = 0.75) + 
  geom_jitter(aes(size = n_cells), width = 0.2, alpha = 0.7) +
  theme_pubr() + 
  xlab("") + ylab("ARG24 Score") +
  stat_compare_means(comparisons = list(c("CR", "NCR")), method = "wilcox.test") +
  scale_color_manual(values = c("#4292C6", "#A50F15")) +
  theme(strip.text.x = element_text(size = 13), legend.position = "right")

# 4. ROC Analysis

roc_obj <- roc(pb_results$Response, pb_results$ARG24_score)

pdf("../03.out/figure3/03.scrna_roc.pdf", width = 4.2, height = 4)
plot(smooth(roc_obj), col = brewer.pal(8, "Dark2")[3])
legend("bottomright", legend = paste0("AUC = ", round(auc(roc_obj), 3)), col = brewer.pal(8, "Dark2")[3], lty = 1)
title(main = "Pseudobulk Validation (N = 81)")
dev.off()

# Save plots
ggsave("../03.out/figure3/03.pseudobulk_barplot.pdf", p1, width = 10, height = 5)
ggsave("../03.out/figure3/03.pseudobulk_violin.pdf", p2, width = 5.5, height = 5)

```

### Single-cell ARG24 score mapping

```{r}
library(Seurat)
library(dplyr)

scaml <- readRDS("../02.data/merged_scaml_response.rds")
scaml.filter <- subset(scaml, mapping_error_QC == 'Pass')
Idents(scaml.filter) <- scaml.filter$predicted_CellType_Broad

# Set color palette
my36colors <- c('#E5D2DD', '#53A85F', '#F1BB72', '#F3B1A0', '#D6E7A3', '#57C3F3', '#476D87',
                '#E95C59', '#E59CC4', '#AB3282', '#23452F', '#BD956A', '#8C549C', '#585658', 
                '#9FA3A8', '#E0D4CA', '#5F3D69', '#58A4C3', "#b20000",'#E4C755', '#F7F398',
                '#AA9A59', '#E63863', '#E39A35', '#C1E6F3', '#6778AE', '#91D0BE', '#B53E2B',
                '#712820', '#DCC1DD', '#CCE0F5', '#CCC9E6', '#625D9E', '#68A180', '#3A6963', '#968175')

pdf("../03.out/figure3/05.scaml.filter_umap.pdf", width = 9, height = 4)
DimPlot(scaml.filter, group.by = "predicted_CellType_Broad", label = TRUE, repel = TRUE, raster = FALSE, cols = my36colors)
dev.off()

# 1. Load 24-gene signature and filter to available genes
gene.min <- read.csv("../02.data/24-gene_signature.csv", 
                    header = TRUE, 
                    row.names = 1, 
                    check.names = FALSE) %>%
  {.[rownames(.) %in% rownames(scaml.filter), , drop = FALSE]} # Keep only genes present in scaml

# Check filtered genes
message("Using ", nrow(gene.min), " of 24 genes:")
print(rownames(gene.min))

# 2. Updated risk score function (matrix-safe)
risk_score <- function(data, lasso_coef) {
  genes <- intersect(rownames(data), rownames(lasso_coef))
  expr_matrix <- as.matrix(data[genes, ]) # Ensure matrix format
  score_mat <- sweep(t(expr_matrix), 2, lasso_coef[genes, 1], `*`)
  rowSums(score_mat, na.rm = TRUE)
}

# 3. Calculate scores using updated Seurat syntax
scaml.filter@meta.data$sc_ARG24 <- risk_score(
  GetAssayData(scaml.filter, assay = "RNA", layer = "data"),
  gene.min
)

# 4. Verify results
summary(scaml.filter@meta.data$sc_ARG24)

# DotPlot plot for ARG24 score
library(ggplot2)
library(Seurat)
library(patchwork)

# Subset DotPlots
p1_data <- DotPlot(subset(scaml.filter, Response == "CR"), features = "sc_ARG24")$data
p2_data <- DotPlot(subset(scaml.filter, Response == "NCR"), features = "sc_ARG24")$data

# Combine to get common limits
all_data <- rbind(p1_data, p2_data)
color_limits <- range(all_data$avg.exp.scaled, na.rm = TRUE)
size_limits <- range(all_data$pct.exp, na.rm = TRUE)

# Rebuild plots with unified scales
p1 <- ggplot(p1_data, aes(x = features.plot, y = id)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_color_gradientn(colors = c("grey90", "blue"), limits = color_limits) +
  scale_size(range = c(1, 8), limits = size_limits) +
  theme_bw() +
  labs(title = "CR", x = "sc_ARG24", y = "") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

p2 <- ggplot(p2_data, aes(x = features.plot, y = id)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_color_gradientn(colors = c("grey90", "blue"), limits = color_limits) +
  scale_size(range = c(1, 8), limits = size_limits) +
  theme_bw() +
  labs(title = "NCR", x = "sc_ARG24", y = "") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Combine
plots <- p1 + p2 + plot_layout(guides = "collect")

ggsave(plots, filename = "../03.out/figure3/dotplot_arg24_response.pdf", width = 6, height = 7.5)

# Extract metadata
meta <- scaml.filter@meta.data

# 5. Summary and Wilcoxon p-values
summary_stats <- meta %>%
  group_by(predicted_CellType_Broad, Response) %>%
  summarise(mean_score = mean(sc_ARG24, na.rm = TRUE), .groups = 'drop') %>%
  tidyr::pivot_wider(names_from = Response, values_from = mean_score, names_prefix = "mean_") %>%
  left_join(
    meta %>%
      group_by(predicted_CellType_Broad) %>%
      summarise(p_val = wilcox.test(sc_ARG24 ~ Response)$p.value, .groups = 'drop'),
    by = "predicted_CellType_Broad"
  ) %>%
  mutate(
    signif = case_when(
      p_val < 0.001 ~ "***",
      p_val < 0.01 ~ "**",
      p_val < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

# 6. Cell counts and filtering
cell_counts <- meta %>%
  group_by(predicted_CellType_Broad, Response) %>%
  summarise(n = n(), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = Response, values_from = n, values_fill = 0) %>%
  mutate(total = CR + NCR)

# write.csv(meta, "../02.data/scaml.filter_meta.csv")
# write.csv(summary_stats, "../02.data/scaml.filter_summary_stats.csv")
# write.csv(cell_counts, "../02.data/scaml.filter_cell_counts.csv")

```

```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(patchwork) # For combining plots

meta <- read.csv("../02.data/scaml.filter_meta.csv", header = T)
summary_stats <- read.csv("../02.data/scaml.filter_summary_stats.csv", header = T)
cell_counts <- read.csv("../02.data/scaml.filter_cell_counts.csv", header = T)

# # Calculate filtering threshold (2‰ of total cells)
# filter_threshold <- dim(meta)[1] * 0.002
# 
# # Keep only cell types with ≥20 cells in each group
# valid_cells <- cell_counts %>%
#   filter(total >= filter_threshold) %>%
#   pull(predicted_CellType_Broad)

# summary_stats <- summary_stats %>%
#   filter(predicted_CellType_Broad %in% valid_cells)
# cell_counts <- cell_counts %>%
#   filter(predicted_CellType_Broad %in% valid_cells)

# Create scatter plot
scatter_plot <- ggplot(summary_stats, aes(x = mean_CR, y = mean_NCR)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", 
              color = "grey40", linewidth = 0.8, alpha = 0.7) +
  # Points colored by significance and sized by p-value
  geom_point(aes(size = -log10(p_val), 
             color = "white", stroke = 0.3)) + # White outline for better visibility
  geom_point(aes(size = -log10(p_val), color = signif), 
             alpha = 0.8) +
  # Text labels with improved formatting
  geom_text_repel(aes(label = paste0(predicted_CellType_Broad, "\n", signif)),
                  size = 3.2, 
                  box.padding = 0.4,
                  point.padding = 0.2,
                  min.segment.length = 0.2,
                  max.overlaps = Inf,
                  segment.color = "grey70") +
  # Size scale for p-values
  scale_size_continuous(range = c(3, 8),
                       breaks = c(20, 50, 100, 150),
                       name = "-log10(p-value)") +
  # Axis labels and title
  labs(x = "Mean sc_ARG24 in CR cells",
       y = "Mean sc_ARG24 in NCR cells",
       title = "Comparison of sc_ARG24 Scores in CR vs NCR by Cell Type",
       subtitle = "Diagonal line represents equal expression in both groups") +
  # Theme improvements
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey40"),
    legend.position = "right",
    legend.box = "vertical",
    legend.spacing.y = unit(0.2, "cm"),
    panel.grid.minor = element_blank(),
    aspect.ratio = 1  # Make plot square
  ) +
  # Equal axis scaling
  coord_fixed(ratio = 1) +
  # Add a small margin around points
  expand_limits(x = c(min(summary_stats$mean_CR, summary_stats$mean_NCR) * 0.95),
                y = c(min(summary_stats$mean_CR, summary_stats$mean_NCR) * 0.95))

scatter_plot

ggsave(scatter_plot, filename = "../03.out/figure3/05.scaml_celltype_scatterplot.pdf", width = 7, height = 6.7)

```

## Cell Type Ratios Between CR and NCR Groups

```{r fig3_celltype_ratios, eval=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggpubr)
library(ggbeeswarm)
library(ggrastr)
library(RColorBrewer)
library(dplyr)
library(ggrepel)

scaml <- readRDS("../02.data/merged_scaml_response.rds")
scaml.filter <- subset(scaml, mapping_error_QC == 'Pass')

# Set color palette
my36colors <- c('#E5D2DD', '#53A85F', '#F1BB72', '#F3B1A0', '#D6E7A3', '#57C3F3', '#476D87',
                '#E95C59', '#E59CC4', '#AB3282', '#23452F', '#BD956A', '#8C549C', '#585658', 
                '#9FA3A8', '#E0D4CA', '#5F3D69', '#58A4C3', "#b20000",'#E4C755', '#F7F398',
                '#AA9A59', '#E63863', '#E39A35', '#C1E6F3', '#6778AE', '#91D0BE', '#B53E2B',
                '#712820', '#DCC1DD', '#CCE0F5', '#CCC9E6', '#625D9E', '#68A180', '#3A6963', '#968175')


# Compute cell counts by type and response
num_cells <- as.data.frame(table(scaml.filter$predicted_CellType_Broad, scaml.filter$Response))
colnames(num_cells) <- c("Cell_type", "Response", "Number")

# Barplot of cell counts
p_count <- ggplot(num_cells, aes(x = reorder(Cell_type, -Number), y = Number, fill = Number)) +
  geom_bar(stat = "identity") +
  facet_grid(Response ~ .) +
  theme_classic() +
  labs(x = "Cell Type", y = "Cell Number (log scale)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_y_log10()
p_count
ggsave(p_count, filename = "../03.out/figure3/04.cellType_counts.pdf", width = 8, height = 5)

# Compute cell ratios by patient
cellratio <- prop.table(table(scaml.filter$predicted_CellType_Broad, scaml.filter$orig.ident), margin = 2)
cellratio <- data.frame(cellratio)
colnames(cellratio) <- c("Cell_type", "orig.ident", "ratio")

# Merge with patient metadata
meta <- scaml.filter@meta.data

sample <- meta[, c("orig.ident", "Response")]
sample <- sample[!duplicated(sample), ]
rownames(sample) <- 1:nrow(sample)

cellratio <- merge(cellratio, sample, by = "orig.ident")

# Calculate filtering threshold (2‰ of total cells)

# filter_threshold <- dim(meta)[1] * 0.002
# 
# # Filter to major cell types (>=300 cells total)
# total_cells <- as.data.frame(table(scaml.filter$predicted_CellType_Broad))
# major_types <- total_cells$Var1[total_cells$Freq >= filter_threshold]
# cellratio <- cellratio[cellratio$Cell_type %in% major_types, ]

# Statistical comparison by cell type
stats <- lapply(unique(cellratio$Cell_type), function(ct) {
  tmp <- cellratio[cellratio$Cell_type == ct, ]
  test <- wilcox.test(ratio ~ Response, data = tmp)
  data.frame(
    Cell_type = ct,
    Ratio_CR = mean(tmp$ratio[tmp$Response == "CR"]),
    Ratio_NCR = mean(tmp$ratio[tmp$Response == "NCR"]),
    P.value = test$p.value
  )
}) %>% bind_rows()

# Annotate significance
stats <- stats %>%
  mutate(
    diff_abundance = Ratio_NCR - Ratio_CR,
    log2fc = log2(Ratio_NCR / Ratio_CR),
    p_label = case_when(
      P.value < 0.001 ~ "***",
      P.value < 0.01 ~ "**",
      P.value < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    negLog10P = -log10(P.value)
  ) %>%
  arrange(desc(log2fc))

stats$Cell_type <- factor(stats$Cell_type, levels = rev(stats$Cell_type))

# Create scatter plot
scatter_plot <- ggplot(stats, aes(x = Ratio_CR, y = Ratio_NCR)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", 
              color = "grey40", linewidth = 0.8, alpha = 0.7) +
  # Points colored by significance and sized by p-value
  geom_point(aes(size = -log10(P.value), 
             color = "white", stroke = 0.3)) + # White outline for better visibility
  geom_point(aes(size = -log10(P.value), color = p_label), 
             alpha = 0.8) +
  # Text labels with improved formatting
  geom_text_repel(aes(label = paste0(Cell_type, "\n", p_label)),
                  size = 3.2, 
                  box.padding = 0.4,
                  point.padding = 0.2,
                  min.segment.length = 0.2,
                  max.overlaps = Inf,
                  segment.color = "grey70") +
  # Axis labels and title
  labs(x = "Mean cell ratio in CR group",
       y = "Mean cell ratio in NCR group") +
  # Theme improvements
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey40"),
    legend.position = "right",
    legend.box = "vertical",
    legend.spacing.y = unit(0.2, "cm"),
    panel.grid.minor = element_blank(),
    aspect.ratio = 1  # Make plot square
  ) +
  # Equal axis scaling
  coord_fixed(ratio = 1) +
  # Add a small margin around points
  expand_limits(x = c(min(stats$Ratio_CR, stats$Ratio_NCR) * 0.95),
                y = c(min(stats$Ratio_CR, stats$Ratio_NCR) * 0.95))

scatter_plot

ggsave(scatter_plot, filename = "../03.out/figure3/05.cellular_abundance.pdf", width = 7, height = 6.7)

```

## Cell Type Distribution Across Patients

```{r fig3_celltype_bar_violin, eval=TRUE, message=FALSE, warning=FALSE}
# Bar plot by patient
p_bar <- ggplot(cellratio) + 
  geom_bar(aes(x = orig.ident, y = ratio, fill = Cell_type), stat = "identity", width = 0.7, size = 0.5, colour = '#222222') + 
  # coord_flip() +
  scale_fill_manual(values = my36colors) +
  labs(x = 'Patient', y = 'Cell Ratio') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.border = element_rect(fill = NA, color = "black", size = 0.5))
p_bar
ggsave(p_bar, filename = "../03.out/figure3/04.cellType_ratio.pdf", width = 18, height = 8)

# Violin plots: all cells
p_violin_all <- ggplot(cellratio, aes(x = Response, y = ratio, fill = Response)) + 
  geom_violin(alpha = 0.5) +
  geom_boxplot_jitter(width = 0.4, alpha = 0.75) + 
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  facet_wrap(. ~ Cell_type, scales = "free", nrow = 3) +
  labs(y = "Cell Ratio") +
  scale_fill_manual(values = c("#4292C6", "#A50F15")) +
  stat_compare_means(comparisons = list(c("CR", "NCR")), method = "wilcox.test", label = "p.signif") +
  theme_pubr() +
  theme(legend.position = "none")
p_violin_all
# ggsave(p_violin_all, filename = "../03.out/figure3/04.cellType_diff_all.pdf", width = 16, height = 8)

```

## ARG24 Scores Across Cell Subtypes

```{r fig3_riskscore_celltypes, eval=TRUE, message=FALSE, warning=FALSE}
library(Seurat)
library(ggplot2)
library(ggpubr)
library(ggbeeswarm)

meta <- read.csv("../02.data/scaml.filter_meta.csv", header = T)
# Define comparison
my_comparisons <- list(c("CR", "NCR"))

# Violin + boxplot with swarm
p_risk_by_celltype <- ggplot(meta, aes(x = Response, y = sc_ARG24, fill = Response)) +
  geom_violin(alpha = 0.6, width = 0.9, color = NA, trim = TRUE) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.4, color = "black") +
  geom_quasirandom(width = 0.2, size = 0.4, alpha = 0.7, aes(color = Response), groupOnX = TRUE) +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 2, fill = "white", color = "black") +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", size = 3) +
  scale_fill_manual(values = c("CR" = "#4292C6", "NCR" = "#A50F15")) +
  scale_color_manual(values = c("CR" = "#4292C6", "NCR" = "#A50F15")) +
  facet_wrap(~ predicted_CellType_Broad, scales = "free_y", nrow = 3) +
  labs(x = NULL, y = "ARG24 Risk Score") +
  theme_classic(base_size = 10) +
  theme(strip.text = element_text(face = "bold", size = 9),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

p_risk_by_celltype

# ggsave(p_risk_by_celltype, filename = "../03.out/figure3/06.ARG24_diff_by_celltype.pdf", width = 16, height = 8)

```

## PCR Validation of ARG24 Score

```{r fig3_pcr_riskscore_calc, eval=TRUE, message=FALSE, warning=FALSE}
dat <- read.csv("../02.data/rjaml_qPCR_39.csv", header = TRUE, row.names = 1, check.names = FALSE)
dat <- dat[, -1]

gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE, row.names = 1, check.names = FALSE)
genes <- row.names(gene.min)
dat <- dat[, genes]

# Compute LASSO-based risk scores
calculate_risk_scores <- function(features, coefficients) {
  scores <- sweep(features, 2, coefficients, `*`)
  risk_scores <- rowSums(scores)
  data.frame(risk_score = risk_scores)
}

rjaml_pcr <- calculate_risk_scores(dat, gene.min[, 1])
clinidf <- read.csv("../02.data/rjaml_qPCR_39_clinical.csv", header = TRUE, row.names = 1, check.names = FALSE)
clinidf <- clinidf[row.names(rjaml_pcr), ]

df <- cbind(clinidf, rjaml_pcr)
df$group <- ifelse(df$CR1_Response == "1", "NCR", "CR")

```

### Correlation Between RT-qPCR and RNA-seq ARG24 Scores

```{r fig3_pcr_vs_rnaseq_correlation, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)

p_corr <- ggscatter(
  na.omit(df[, c("risk_score", "RNAseq_ARG24")]), 
  x = "risk_score", y = "RNAseq_ARG24",
  alpha = 0.6, color = "black",
  add = "reg.line",
  add.params = list(color = "#377eb8", fill = "#cce5ff"),
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(method = "pearson", label.sep = "\n", label.x.npc = "left", label.y.npc = "top", size = 5)
) +
  theme_classic(base_size = 14) +
  labs(x = "ARG24 Score (RT-qPCR)", y = "ARG24 Score (RNA-seq)") +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

p_corr

# ggsave(p_corr, filename = "../03.out/figure3/08.pcr_vs_rnaseq_correlation.pdf", width = 4, height = 4)
```

### Risk Score Distribution in RT-qPCR Cohort

```{r fig3_pcr_box_roc, eval=TRUE, message=FALSE, warning=FALSE}
library(pROC)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)

# Violin + boxplot
p_box <- ggplot(df[, c("group", "risk_score")], aes(x = group, y = risk_score, color = group)) + 
  geom_violin() +
  geom_boxplot_jitter(width = 0.4, alpha = 0.75) +
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  theme_pubr() + 
  xlab("") + ylab("ARG24 Risk Score") +
  theme(strip.text.x = element_text(size = 13)) +
  stat_compare_means(comparisons = list(c('CR', 'NCR')), method = 'wilcox.test') + 
  scale_color_manual(values = c("#4292C6", "#A50F15")) +
  theme(legend.position = 'none')

p_box

# ggsave(p_box, filename = "../03.out/figure3/09.pcr_riskscore_boxplot.pdf", width = 2.4, height = 3.8)

# ROC Analysis
auc_val <- auc(df$group, df$risk_score)
roc_obj <- roc(df$group, df$risk_score)

# pdf("../03.out/figure3/10.pcr_riskscore_roc.pdf", width = 3.8, height = 3.8)
plot(smooth(roc_obj), col = brewer.pal(8, "Dark2")[7])
legend("bottomright", legend = paste0("ARG24 AUC=", round(auc_val, 3)), col = brewer.pal(8, "Dark2")[7], lty = 1)
title(main = "RT-qPCR Validation Cohort")
# dev.off()
```

## D5-PBCR Evaluation in RT-qPCR Cohort

```{r fig3_pcr_d5pbcr_box_roc, eval=TRUE, message=FALSE, warning=FALSE}
library(pROC)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)

# Define D5-PBCR status group
dat <- df[, c("D5_PBCR", "ARG24")]
dat$D5_PBCR_group <- ifelse(dat$D5_PBCR == "1", "Positive", "Negative")

# Boxplot comparing ARG24 scores by D5-PBCR status
p_d5pbcr <- ggplot(dat, aes(x = D5_PBCR_group, y = ARG24, color = D5_PBCR_group)) + 
  geom_violin() +
  geom_boxplot_jitter(width = 0.4, alpha = 0.75) + 
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  theme_pubr() + 
  xlab("") + ylab("ARG24 Risk Score") +
  theme(strip.text.x = element_text(size = 13)) +
  stat_compare_means(comparisons = list(c("Negative", "Positive")), method = "wilcox.test") + 
  scale_color_manual(values = c("Negative" = "#4292C6", "Positive" = "#A50F15")) +
  theme(legend.position = 'none')

p_d5pbcr

# ggsave(p_d5pbcr, filename = "../03.out/figure3/11.d5pbcr_riskscore_boxplot.pdf", width = 2.8, height = 3.8)

# ROC Analysis for D5-PBCR
auc_d5 <- auc(df$group, df$D5_PBCR)
roc_d5 <- roc(df$group, df$D5_PBCR)

# pdf("../03.out/figure3/12.d5pbcr_riskscore_roc.pdf", width = 3.8, height = 3.8)
plot(roc_d5, col = brewer.pal(8, "Dark2")[8])
legend("bottomright", legend = paste0("D5-PBCR AUC=", round(auc_d5, 3)), col = brewer.pal(8, "Dark2")[8], lty = 1)
title(main = "RT-qPCR Validation Cohort")
# dev.off()

# Export combined clinical and risk score data
# write.csv(df, "../02.data/rjaml_qPCR_39_clinical_ARG24.csv")
```
