# Figure 4: Integration of ARG24 and AML genomics

The varied ARG24 score among AML patients likely reflected the functional consequences of specific genomic abnormalities on cellular response to chemotherapy. To examine how ARG24 score relate to clinical and molecular features of AML, we further integrated the genomic/clinical data of 1,336 patient samples from the RJAML, BeatAML, Leucegene and TCGA-LAML cohorts.  

## Clinical Factors and ARG24 Score Association

```{r fig4_clinical_assoc, eval=TRUE, message=FALSE, warning=FALSE}
# Load libraries
library(tidyverse)
library(ggpubr)
library(ggsci)
library(ggrastr)
library(readxl)
library(patchwork)

# Load data
dat <- read_xlsx("../02.data/Combined_cohorts_riskScore_Clinical_Final.xlsx")

# Define plotting function
plot_clinical <- function(df, xvar, yvar = "RiskScore", comp_list, fill_var = NULL) {
  ggplot(df, aes_string(x = xvar, y = yvar, color = xvar)) +
    geom_violin(alpha = 0.5) +
    geom_boxplot_jitter(width = 0.4, alpha = 0.75) +
    geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
    stat_compare_means(comparisons = comp_list, method = "wilcox.test", label = "p.signif") +
    theme_pubr() +
    ylab("ARG24 Score") + xlab(NULL) +
    scale_color_nejm() +
    theme(legend.position = "none", strip.text.x = element_text(size = 13))
}

# Prepare data and plot
plots <- list(
  plot_clinical(filter(dat, !is.na(Age)) %>% mutate(Age = ifelse(Age < 60, "<60", ">=60")), "Age", comp_list = list(c("<60", ">=60"))),
  plot_clinical(filter(dat, !is.na(Gender)), "Gender", comp_list = list(c("Female", "Male"))),
  plot_clinical(filter(dat, isDeNovo %in% c("Yes", "No")), "isDeNovo", comp_list = list(c("Yes", "No"))),
  plot_clinical(filter(dat, ELN2022 %in% c("Favorable", "Intermediate", "Adverse")) %>% 
                  mutate(ELN2022 = factor(ELN2022, levels = c("Favorable", "Intermediate", "Adverse"))),
                "ELN2022", comp_list = list(c("Favorable", "Intermediate"), c("Intermediate", "Adverse"), c("Favorable", "Adverse"))),
  plot_clinical(filter(dat, Cytogenetic_risk %in% c("Good", "Intermediate", "Poor")), "Cytogenetic_risk",
                comp_list = list(c("Good", "Intermediate"), c("Intermediate", "Poor"), c("Good", "Poor"))),
  plot_clinical(filter(dat, FAB %in% c("M0", "M1", "M2", "M4", "M5", "M6", "M7")), "FAB",
                comp_list = list(c("M0", "M1"), c("M0", "M2"), c("M0", "M4"), c("M0", "M5"), c("M0", "M6"), c("M0", "M7")))
)

# Combine and export
combined <- (plots[[1]] | plots[[2]] | plots[[3]]) / (plots[[4]] | plots[[5]]) / plots[[6]]
combined

# ggsave(combined, filename = "../03.out/figure4/01.clinical.arg24.pdf", width = 12, height = 9)

```

## Cytogenetic Alterations and ARG24 Score

```{r fig4_cytogenetics_ridges, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggpubr)
library(ggsci)
library(ggridges)
library(cowplot) 
library(ggrastr)
library(readxl)

# Load integrated cohort data with risk scores and cytogenetic info
hcat <- read_xlsx("../02.data/Combined_cohorts_riskScore_Clinical_Final.xlsx")

# Select relevant cytogenetic markers
cyto_vars <- c(
  "inv(3): GATA2/MECOM", "t(11q23): MLL_Fusion", "t(9;22): BCR-ABL",
  "t(5;11): NUP98-NSD1", "inv(16): CBFB-MYH11", "t(8;21): RUNX1-RUNX1T1", 
  "Monosomy 5;del(5q)", "Monosomy 7;del(7q)", "Monosomy 17;del(17p)", "Complex Karyotype"
)

cytogenetics <- hcat %>% select(RiskScore, all_of(cyto_vars))

# Convert statuses to binary and reshape
alteration <- cytogenetics %>% 
  pivot_longer(-RiskScore, names_to = "Alteration", values_to = "Status") %>%
  mutate(Status = if_else(Status %in% c("NA", "Negative", "Other"), 0, 1)) %>%
  filter(Status == 1)

# Cap extreme RiskScores for consistent density visual
alteration <- alteration %>% mutate(RiskScore = pmin(RiskScore, 7))

# Density ridge plot of ARG24 score by cytogenetic alteration
p_cytogenetics <- ggplot(alteration, aes(y = fct_reorder(Alteration, RiskScore), x = RiskScore, fill = stat(x))) +
  geom_density_ridges_gradient(
    jittered_points = TRUE, scale = 1.25,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.3
  ) +
  scale_fill_gradient2(
    low = "#D9D9D9", high = "#9970AB", midpoint = median(alteration$RiskScore), name = ""
  ) +
  theme_pubr() +
  labs(x = "ARG24 Risk Score", y = "Cytogenetic Alteration") +
  xlim(c(-2, 8)) +
  theme(
    axis.title.x = element_text(size = 14),
    axis.text.y = element_text(size = 11),
    legend.position = "right"
  )

p_cytogenetics

# Save figure
# ggsave(p_cytogenetics, filename = "../03.out/figure4/02.cytogenetics.arg24.pdf", width = 7, height = 5)
```

## Single Gene Mutations and ARG24 Score

```{r fig4_singlemutation_ridges, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggpubr)
library(ggsci)
library(ggridges)
library(cowplot) 
library(ggrastr)
library(readxl)

# Load and subset mutation data
dat <- read_xlsx("../02.data/Combined_cohorts_riskScore_Clinical_Final.xlsx")
row.names(dat) <- dat$Patient
mutation_mat <- dat[, c(2, 30:36, 38:64)]

# Wilcoxon tests comparing ARG24 scores by mutation status
stat <- lapply(2:ncol(mutation_mat), function(i) {
  tmp <- mutation_mat[, c(1, i)]
  colnames(tmp)[2] <- "Mutation"
  tmp <- tmp %>% filter(Mutation %in% c("Positive", "Negative"))
  wt <- wilcox.test(RiskScore ~ Mutation, data = tmp)
  data.frame(
    Negative_median = median(tmp$RiskScore[tmp$Mutation == "Negative"]),
    Positive_median = median(tmp$RiskScore[tmp$Mutation == "Positive"]),
    P.value = wt$p.value,
    Number = sum(tmp$Mutation == "Positive"),
    Mutations = colnames(mutation_mat)[i]
  )
}) %>% bind_rows()

# Order by median score in mutation-positive samples
stat <- stat %>% mutate(log2FC = log2(Positive_median / Negative_median)) %>% 
  arrange(desc(Positive_median))

# Extract RiskScore and mutation status
hcat <- read_xlsx("../02.data/Combined_cohorts_riskScore_Clinical_Final.xlsx")
cytogenetics <- hcat[, c("RiskScore", stat$Mutations)]

alteration <- cytogenetics %>%
  pivot_longer(-RiskScore, names_to = "Alteration", values_to = "Status") %>%
  mutate(Status = if_else(Status %in% c("NA", "Negative", "Other"), 0, 1)) %>%
  filter(Status == 1) %>%
  mutate(RiskScore = pmin(RiskScore, 7))

# Plot ridge plot of ARG24 score by single mutations
p_single_mut <- ggplot(alteration, aes(y = fct_reorder(Alteration, RiskScore), x = RiskScore, fill = stat(x))) +
  geom_density_ridges_gradient(
    jittered_points = TRUE, scale = 1.25,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.3
  ) +
  scale_fill_gradient2(low = "#4393C3", high = "#9970AB", midpoint = median(alteration$RiskScore), name = "") +
  theme_pubr() +
  labs(x = "ARG24 Risk Score", y = "Gene Mutation") +
  xlim(-2, 8) +
  theme(
    legend.position = "right",
    axis.title.x = element_text(size = 14),
    axis.text.y = element_text(size = 11)
  )

p_single_mut

# Save plot
# ggsave(p_single_mut, filename = "../03.out/figure4/03.single_mutations.arg24.pdf", width = 8, height = 12)
```

## Mutation Pairs and ARG24 Score

```{r fig4_mutationpair_ridges, eval=TRUE, message=FALSE, warning=FALSE, fig.height=14, fig.width=6}
library(tidyverse)
library(ggridges)
library(ggpubr)
library(readr)
library(stringr)

# ----------------------
# Load and preprocess mutation data
# ----------------------
combos <- read_csv("../02.data/Combined_cohorts_riskScore_Clinical_Final.csv") %>% 
  column_to_rownames(var = "Patient")

# Extract mutation columns and binarize mutation status
mut_mat <- combos %>% 
  select(contains("mut")) %>% 
  mutate(across(everything(), ~ replace_na(as.character(.), "NA"))) %>% 
  mutate(across(everything(), ~ recode(., "Positive" = "1", "Negative" = "0", "NA" = "0"))) %>% 
  mutate(across(everything(), as.integer))

# Filter for patients with ≥2 mutations and remove rare mutations (sum > 1)
mut_mat <- mut_mat %>% 
  select(order(colSums(.), decreasing = TRUE)) %>% 
  filter(rowSums(.) >= 2) %>% 
  select(where(~ sum(.) > 1))

# Generate pairwise mutation combinations
combinations <- combn(colnames(mut_mat), 2)
df_comb <- purrr::map_dfc(seq_len(ncol(combinations)), function(i) {
  combo <- combinations[, i]
  setNames(as.data.frame(mut_mat[[combo[1]]] * mut_mat[[combo[2]]]), paste(combo, collapse = " + "))
})

# Retain combinations with at least two co-occurrences
combo2 <- df_comb %>% 
  select(where(~ sum(.) > 1)) %>% 
  rownames_to_column(var = "Patient")

# Save to CSV
# write_csv(combo2, "../02.data/Two_mutations_combination_riskScore_final.csv", col_names = TRUE)

combo2 <- combo2 %>% 
  pivot_longer(-Patient, names_to = "Combo", values_to = "Present") %>% 
  filter(Present > 0) %>% 
  mutate(
    dnmt3a_dup = str_detect(Combo, "DNMT3A \\(") & str_detect(Combo, "DNMT3A_mut"),
    flt3_dup = str_detect(Combo, "FLT3 \\(") & str_detect(Combo, "FLT3_mut")
  ) %>% 
  filter(!dnmt3a_dup & !flt3_dup) %>% 
  select(-dnmt3a_dup, -flt3_dup) %>% 
  mutate(Combo = str_replace_all(Combo, "_mut", ""))

# Summarize frequency for each mutation pair
pair_freq <- combo2 %>% 
  group_by(Combo) %>% 
  summarise(Total = sum(Present), .groups = "drop") %>% 
  arrange(desc(Total))
# write.csv(pair_freq, "../02.data/Two_mutations_combination_Freq_riskScore_final.csv")

```

### ARG24 scores and Mutation Pairs (Two mutations)

```{r eval=TRUE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggridges)
library(ggpubr)
library(readr)

# ----------------------
# Perform statistical tests and plot
# ----------------------
# Load combination risk score matrix
combo_data <- read_csv("../02.data/Two_mutations_combination_riskScore_final.csv")

# Wilcoxon test comparing ARG24 score between pair-present vs. absent
stat <- lapply(3:ncol(combo_data), function(i) {
  tmp <- combo_data[, c(2, i)]
  colnames(tmp)[2] <- "Mutation"
  wt <- wilcox.test(RiskScore ~ Mutation, data = tmp)
  data.frame(
    Negative_median = median(tmp$RiskScore[tmp$Mutation == 0], na.rm = TRUE),
    Positive_median = median(tmp$RiskScore[tmp$Mutation == 1], na.rm = TRUE),
    P.value = wt$p.value,
    Mutations = colnames(combo_data)[i]
  )
}) %>% bind_rows()

# Merge with frequency info and filter for common co-occurrences
freq <- read_csv("../02.data/Two_mutations_combination_Freq_riskScore_final.csv")
stat <- left_join(stat, freq, by = "Mutations")
plot_pairs <- stat %>% filter(Total >= 10)

# Prepare matrix for plotting
mutants <- combo_data %>% select(RiskScore, all_of(plot_pairs$Mutations))
alteration <- mutants %>% 
  pivot_longer(-RiskScore, names_to = "Alteration", values_to = "Status") %>% 
  filter(Status > 0)

# Ridge plot of ARG24 score across co-occurring mutation pairs
p_pair_mut <- ggplot(alteration, aes(y = fct_reorder(Alteration, RiskScore), x = RiskScore, fill = stat(x))) +
  geom_density_ridges_gradient(
    jittered_points = TRUE, scale = 1.25,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.3
  ) +
  scale_fill_gradient2(low = "#01665E", high = "#8C510A", midpoint = median(alteration$RiskScore), name = "") +
  theme_pubr() +
  labs(x = "ARG24 Risk Score", y = "Co-occurring Mutation Pairs") +
  theme(
    legend.position = "right",
    axis.title.x = element_text(size = 14),
    axis.text.y = element_text(size = 11)
  )

p_pair_mut

# Save plot
# ggsave(p_pair_mut, filename = "../03.out/figure4/04.two_mutations.arg24.pdf", width = 8, height = 19)

```

## Mutation Triplets and ARG24 Score

```{r}
library(tidyverse)
library(ggridges)
library(ggpubr)
library(readr)
library(stringr)

# ----------------------
# Load and preprocess mutation data
# ----------------------
combos <- read_csv("../02.data/Combined_cohorts_riskScore_Clinical_Final.csv") %>% 
  column_to_rownames(var = "Patient")


# Extract mutation columns and convert mutation status to binary
mut_mat <- combos %>% 
  select(contains("mut")) %>% 
  mutate(across(everything(), ~ replace_na(as.character(.), "NA"))) %>% 
  mutate(across(everything(), ~ recode(., "Positive" = "1", "Negative" = "0", "NA" = "0"))) %>% 
  mutate(across(everything(), as.integer))

# Filter for patients with ≥3 mutations and remove rare mutations (sum > 1)
mut_mat <- mut_mat %>% 
  select(order(colSums(.), decreasing = TRUE)) %>% 
  filter(rowSums(.) >= 3) %>% 
  select(where(~ sum(.) > 1))

# Generate triplet combinations of mutations
triplet_combos <- combn(colnames(mut_mat), 3)
df_triplet <- purrr::map_dfc(seq_len(ncol(triplet_combos)), function(i) {
  combo <- triplet_combos[, i]
  col1 <- mut_mat[[combo[1]]]
  col2 <- mut_mat[[combo[2]]]
  col3 <- mut_mat[[combo[3]]]
  setNames(as.data.frame(col1 * col2 * col3), paste(combo, collapse = " + "))
})

# Retain triplets with at least two co-occurrences
combo3 <- df_triplet %>% 
  select(where(~ sum(.) > 1)) %>% 
  rownames_to_column(var = "Patient")

# ----------------------
# Tidy the triplet combinations
# ----------------------
combo3 <- combo3 %>% 
  pivot_longer(-Patient, names_to = "Combo", values_to = "Present") %>% 
  filter(Present > 0) %>% 
  mutate(
    dnmt3a_dup = str_detect(Combo, "DNMT3A \\(") & str_detect(Combo, "DNMT3A_mut"),
    flt3_dup = str_detect(Combo, "FLT3 \\(") & str_detect(Combo, "FLT3_mut")
  ) %>% 
  filter(!dnmt3a_dup & !flt3_dup) %>% 
  select(-dnmt3a_dup, -flt3_dup) %>% 
  mutate(Combo = str_replace_all(Combo, "_mut", ""))

# Summarize frequency of each triplet
triplet_freq <- combo3 %>% 
  group_by(Combo) %>% 
  summarise(Total = sum(Present), .groups = "drop") %>% 
  arrange(desc(Total))

# # Optionally save triplet matrix or frequency file
# write_csv(combo3, "../02.data/Three_mutations_combination_riskScore_final.csv")
# write_csv(triplet_freq, "../02.data/Three_mutations_combination_Freq_riskScore_final.csv")

```

```{r eval=TRUE, message = FALSE, warning = FALSE}
# ----------------------
# Statistical analysis and plotting
# ----------------------
# Load triplet ARG24 score matrix
dat <- read_csv("../02.data/Three_mutations_combination_riskScore_final.csv")

# Perform Wilcoxon test for each triplet combination
stat <- lapply(3:ncol(dat), function(i) {
  tmp <- dat[, c(2, i)]
  colnames(tmp)[2] <- "Mutation"
  tmp_0 <- filter(tmp, Mutation == 0)
  tmp_1 <- filter(tmp, Mutation == 1)
  testout <- wilcox.test(tmp_0$RiskScore, tmp_1$RiskScore, alternative = "two.sided")
  data.frame(
    Negative_median = median(tmp_0$RiskScore, na.rm = TRUE),
    Positive_median = median(tmp_1$RiskScore, na.rm = TRUE),
    P.value = testout$p.value,
    Mutations = colnames(dat)[i]
  )
}) %>% bind_rows()

# Merge statistics with triplet frequency
freq <- read_csv("../02.data/Three_mutations_combination_Freq_riskScore_final.csv")
stat <- left_join(stat, freq, by = "Mutations")

# Save statistics to CSV (optional)
# write_csv(stat, "riskScore_genomics/Three_mutations_Mut_vs.WT_stat.csv")

# Load filtered set for plotting (triplets with Total >= 5)
plot_stat <- read_csv("../02.data/Three_mutations_Mut_vs.WT_stat.csv") %>% 
  filter(Total >= 5)

# Prepare data for plotting
mutants <- dat %>% select(RiskScore, all_of(plot_stat$Mutations))
alteration <- mutants %>% 
  pivot_longer(-RiskScore, names_to = "Alteration", values_to = "Status") %>% 
  filter(Status > 0)

# Ridge plot: ARG24 score across triplet mutation combinations
p_triplet_mut <- ggplot(alteration, aes(y = fct_reorder(Alteration, RiskScore), x = RiskScore, fill = stat(x))) +
  geom_density_ridges_gradient(
    jittered_points = TRUE, scale = 1,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.3
  ) +
  scale_fill_gradient2(low = "#01665E", high = "#8C510A", midpoint = median(alteration$RiskScore), name = "") +
  theme_pubr() +
  labs(x = "ARG24 Risk Score", y = "Triplet Mutation Combinations") +
  theme(
    legend.position = "right",
    axis.title.x = element_text(size = 14),
    axis.text.y = element_text(size = 10)
  )

p_triplet_mut

# Save plot
# ggsave(p_triplet_mut, filename = "../03.out/figure4/05.three_mutations.arg24.pdf", width = 9, height = 18)

```

## FLT3 Subtypes and ARG24 Score

```{r fig4_flt3subtypes_density, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggpubr)
library(ggsci)
library(ggridges)
library(cowplot)
library(ggrastr)
library(readxl)

# ----------------------
# Load and combine FLT3-subtype data from multiple sheets
# ----------------------
dat1 <- read_xlsx("../02.data/ITD_TKD_combinations.xlsx", sheet = 1)
dat2 <- read_xlsx("../02.data/ITD_TKD_combinations.xlsx", sheet = 2)
dat3 <- read_xlsx("../02.data/ITD_TKD_combinations.xlsx", sheet = 3)
score <- read_xlsx("../02.data/ITD_TKD_combinations.xlsx", sheet = 4)

dat_combined <- list(dat1, dat2, dat3) %>%
  reduce(full_join, by = "Patient") %>%
  full_join(score, by = "Patient") %>%
  mutate(across(where(is.numeric), ~ replace_na(., 0))) %>%  # replaces NA with 0 in numeric columns
  select(-Patient)

# ----------------------
# Prepare data for density plot
# ----------------------
alteration <- dat_combined %>%
  pivot_longer(-RiskScore, names_to = "Alteration", values_to = "Status") %>%
  filter(Status > 0) %>%
  mutate(RiskScore = pmin(RiskScore, 5))

# Density ridge plot
p_flt3_density <- ggplot(alteration, aes(y = fct_reorder(Alteration, RiskScore), x = RiskScore, fill = stat(x))) +
  geom_density_ridges_gradient(
    jittered_points = TRUE, scale = 1,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.3
  ) +
  scale_fill_gradient2(low = "#4393C3", high = "#9970AB", midpoint = median(alteration$RiskScore), name = "") +
  theme_pubr() +
  labs(x = "ARG24 Risk Score", y = "FLT3 Subtype Combinations") +
  theme(legend.position = "right", axis.title.x = element_text(size = 15))

p_flt3_density

# Save plot
# ggsave(p_flt3_density, filename = "../03.out/figure4/06.flt3_subtypes_density.pdf", width = 7.2, height = 4.5)
```

```{r fig4_flt3subtypes_boxplot, eval=TRUE, message=FALSE, warning=FALSE}
library(RColorBrewer)

# ----------------------
# Annotate alteration groups and types
# ----------------------
col_val <- pal_nejm(alpha = 0.7)(8)

alteration <- alteration %>%
  mutate(
    Type = if_else(Alteration %in% c("FLT3-ITD", "NPM1 + FLT3-ITD", "FLT3-ITD + DNMT3A", "FLT3-ITD + TET2",
                                     "NPM1 + FLT3-ITD + DNMT3A", "NPM1 + FLT3-ITD + TET2"), "ITD", "TKD"),
    Group = case_when(
      Alteration %in% c("FLT3-ITD", "FLT3-TKD") ~ "FLT3",
      Alteration %in% c("NPM1 + FLT3-ITD", "NPM1 + FLT3-TKD") ~ "FLT3+NPM1",
      Alteration %in% c("FLT3-ITD + DNMT3A", "DNMT3A + FLT3-TKD") ~ "FLT3+DNMT3A",
      Alteration %in% c("FLT3-ITD + TET2", "TET2 + FLT3-TKD") ~ "FLT3+TET2",
      Alteration %in% c("NPM1 + FLT3-ITD + DNMT3A", "NPM1 + DNMT3A + FLT3-TKD") ~ "FLT3+NPM1+DNMT3A",
      TRUE ~ "FLT3+NPM1+TET2"
    ),
    Type = factor(Type, levels = c("ITD", "TKD")),
    Group = factor(Group, levels = c("FLT3", "FLT3+NPM1", "FLT3+DNMT3A", "FLT3+TET2", "FLT3+NPM1+DNMT3A", "FLT3+NPM1+TET2"))
  )

# Violin + boxplot of ARG24 scores across FLT3 subtypes
p_flt3_box <- ggplot(alteration, aes(x = Type, y = RiskScore, color = Type)) +
  geom_violin(alpha = 0.6, width = 1, position = position_dodge(width = 0.8), size = 0.75) +
  geom_jitter_rast(alpha = 0.6, position = position_jitterdodge(jitter.width = 0.5, jitter.height = 0, dodge.width = 1)) +
  geom_boxplot(alpha = 0.6, width = 0.2, position = position_dodge(width = 0.8), size = 0.75, outlier.colour = NA) +
  scale_color_manual(values = col_val[4:6]) +
  facet_wrap(~ Group, ncol = 3) +
  theme_classic() +
  stat_compare_means(label = "p.format", comparisons = list(c("ITD", "TKD")), method = "t.test") +
  labs(x = NULL, y = "ARG24 Risk Score") +
  theme(
    legend.position = "none",
    text = element_text(size = 16),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1)
  )

p_flt3_box

# Save plot
# ggsave(p_flt3_box, filename = "../03.out/figure4/07.flt3_subtypes_boxplot.pdf", width = 7, height = 5)
```

## Patient data for validation FLT3 mutations

```{r}
library(ggstatsplot)
library(ggplot2)
library(patchwork)

# -------------------------------
# Load and preprocess clinical cohort data
# -------------------------------
dat <- read.csv("../02.data/compiledPatientCohorts_FLT3.csv", header = TRUE, check.names = FALSE)

# Standardize response labels to CR and NCR
dat$Response <- ifelse(dat$Response == "CR", "CR", "NCR")

# -------------------------------
# Plot 1: FLT3-ITD vs Response
# -------------------------------
p1 <- ggbarstats(
  data = dat,
  x = Response,
  y = `FLT3-ITD`,
  palette = "Paired",
  title = "FLT3-ITD"
) +
  theme_classic(base_size = 13)

# -------------------------------
# Plot 2: FLT3-TKD vs Response
# -------------------------------
p2 <- ggbarstats(
  data = dat,
  x = Response,
  y = `FLT3-TKD`,
  palette = "Paired",
  title = "FLT3-TKD"
) +
  theme_classic(base_size = 13)

# -------------------------------
# Plot 3: Combined FLT3 subtype comparison (ITD vs TKD)
# -------------------------------
# Subset patients with FLT3-ITD or FLT3-TKD mutations
dat1 <- rbind(
  subset(dat, `FLT3-ITD` == "Positive"),
  subset(dat, `FLT3-TKD` == "Positive")
)

# Add mutation type label
dat1$Mutation <- c(
  rep("FLT3-ITD", sum(dat$`FLT3-ITD` == "Positive")),
  rep("FLT3-TKD", sum(dat$`FLT3-TKD` == "Positive"))
)

# Plot
p3 <- ggbarstats(
  data = dat1,
  x = Response,
  y = Mutation,
  palette = "Paired",
  title = "FLT3-ITD vs FLT3-TKD"
) +
  theme_classic(base_size = 13)

# -------------------------------
# Combine and export plots
# -------------------------------
combined_plot <- (p1 | p2 | p3)

# Print
combined_plot
# ggsave(combined_plot, filename = "../03.out/figure4/11.patient_flt3.pdf", width = 9, height = 4)

```

## Mutation Landscape Between ARG24-High and ARG24-Low

```{r fig4_mutation_heatmap, eval=TRUE, message=FALSE, warning=FALSE, fig.height=8, fig.width=16}
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(ggsci)
library(readxl)

# ------------------------------
# Load and preprocess mutation matrix
# ------------------------------
mut_matrix <- read.csv("../02.data/RiskScore_mutmat.csv", header = TRUE, row.names = 1)
mut_matrix <- t(mut_matrix)  # Transpose so genes are rows, patients are columns

# Define color palette
mycol <- c(
  pal_nejm("default")(8), "#8DD3C7", "#CCEBC5", "#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C",
  "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A", "#66C2A5", "#FC8D62",
  "#8DA0CB", "#E78AC3", "#A6D854", "#FFD92F", "#E5C494", "#B3B3B3"
)

# Define how alterations are displayed
alter_fun <- list(
  background = function(x, y, w, h) {
    grid.rect(x, y, w - unit(0.0, "mm"), h - unit(1.5, "mm"), gp = gpar(fill = '#ebebe1ff', col = NA))
  },
  NAN = function(x, y, w, h) {
    grid.rect(x, y, w - unit(0.0, "mm"), h - unit(1.5, "mm"), gp = gpar(fill = "lightgrey", col = NA))
  },
  Positive = function(x, y, w, h) {
    grid.rect(x, y, w - unit(0.0, "mm"), h - unit(1.5, "mm"), gp = gpar(fill = "#1F78B4", col = NA))
  }
)

col <- c("NAN" = "lightgrey", "Positive" = "#1F78B4", "Negative" = "#ebebe1ff")

# ------------------------------
# Load and preprocess clinical annotation
# ------------------------------
pdata <- read_xlsx("../02.data/RiskScore_mutmat_clinical.xlsx")
pdata <- subset(pdata, pdata$Patient %in% colnames(mut_matrix))
pdata$Group <- ifelse(pdata$RiskScore >= median(pdata$RiskScore), "High", "Low")
mut_matrix <- mut_matrix[, pdata$Patient]  # Align columns

# Define color scale for Age
col_age <- colorRamp2(c(20, 90), c("#F7F7F7", "#1B7837"))

# Define clinical annotation for heatmap
ClinicAnno <- HeatmapAnnotation(
  ELN2022 = pdata$ELN2022,
  Denovo = pdata$isDeNovo,
  Age = pdata$Age,
  Gender = pdata$Gender,
  Cohort = pdata$Cohort,
  "inv(3)" = pdata$`inv(3): GATA2/MECOM`,
  MLL_Fusion = pdata$`t(11q23): MLL_Fusion`,
  "t(5;11)" = pdata$`t(5;11): NUP98-NSD1`,
  "inv(16)" = pdata$`inv(16): CBFB-MYH11`,
  "t(8;21)" = pdata$`t(8;21): RUNX1-RUNX1T1`,
  "del(5q)" = pdata$`Monosomy 5;del(5q)`,
  "del(7q)" = pdata$`Monosomy 7;del(7q)`,
  "del(17p)" = pdata$`Monosomy 17;del(17p)`,
  Complex_Karyotype = pdata$`Complex Karyotype`,
  Group = pdata$Group,
  foo = anno_points(
    pdata$RiskScore,
    size = unit(3, "points"),
    height = unit(1.5, "cm")
  ),
  show_annotation_name = TRUE,
  col = list(
    Age = col_age,
    Gender = c("Female" =  "#CAB2D6", "Male" = "#6A3D9A", "NAN" = "#B3B3B3"),
    Cohort = c("BeatAML" = "#E18727FF", "RJAML" = "#20854EFF", "TCGA-LAML" = "#7876B1FF", "Leucegene" = "#6F99ADFF"),
    "inv(3)" = col, MLL_Fusion = col, "t(5;11)" = col, "inv(16)" = col, "t(8;21)" = col,
    "del(5q)" = col, "del(7q)" = col, "del(17p)" = col, Complex_Karyotype = col,
    Denovo = c("Yes" = "#299D8F", "No" = "#E9C46A", "NAN" = "grey"),
    ELN2022 = c("Favorable" = "#66C2A5", "Intermediate" = "#FC8D62", "Adverse" = "#8DA0CB", "NAN" = "grey"),
    Group = c("Low" = "#A6CEE3", "High" = "#FB9A99")
  ),
  annotation_name_gp = gpar(fontsize = 8)
)

# ------------------------------
# Generate oncoPrint heatmap
# ------------------------------
p1 <- oncoPrint(
  mut_matrix,
  col = col,
  alter_fun = alter_fun,
  remove_empty_columns = FALSE,
  alter_fun_is_vectorized = FALSE,
  row_names_side = "left",
  show_pct = FALSE,
  column_order = pdata$Patient,
  right_annotation = NULL,
  top_annotation = ClinicAnno
)

# Display plot
p1

# Save to PDF
# pdf("../03.out/figure4/08.mutation_heatmap.pdf", width = 14, height = 8)
# p1
# dev.off()
```

## Fisher's Exact Test for Mutation Enrichment

```{r fig4_fisher_test, eval=TRUE, message=FALSE, warning=FALSE}
# Load required libraries
library(tidyverse)
library(ggplot2)
library(ggsci)
library(scales)
library(ggbreak)

# ------------------------------
# Load and preprocess input data
# ------------------------------
fisher.dat <- read.csv("../02.data/Mutation_fisher_input.csv", header = TRUE, row.names = 1, check.names = FALSE, stringsAsFactors = FALSE)

# Initialize matrix to store Fisher test results
stat1 <- matrix(NA, nrow = ncol(fisher.dat) - 1, ncol = 4)

# Perform Fisher test for each gene
for (i in 2:ncol(fisher.dat)) {
  m1 <- fisher.dat[, c(1, i)] %>% drop_na()
  m1$RiskScore <- ifelse(m1$RiskScore >= median(m1$RiskScore), "High", "Low")
  
  m2 <- table(m1)
  m3 <- matrix(c(m2[3], m2[4], m2[1], m2[2]), ncol = 2, byrow = TRUE)

  s1 <- fisher.test(m3)
  stat1[i - 1, ] <- c(round(s1$p.value, 5), round(s1$estimate, 3), round(s1$conf.int, 3))
}

# Format output
stat1 <- as.data.frame(stat1)
stat1$Gene <- colnames(fisher.dat)[2:ncol(fisher.dat)]
colnames(stat1) <- c("P.value", "Odds_Ratio", "Lower95%", "Upper95%", "Gene")
stat <- stat1 %>% select(Gene, P.value, Odds_Ratio, `Lower95%`, `Upper95%`) %>% arrange(P.value)
stat$fdr <- p.adjust(stat$P.value, method = "fdr")

# Save results
# write.csv(stat, "../03.out/Mutation_fisher_out.csv")
```

```{r fig4_fisher_dotplot, eval=TRUE, message=FALSE, warning=FALSE}
# Load processed results
tab <- read.csv("../03.out/Mutation_fisher_out.csv", header = TRUE, row.names = 1, check.names = FALSE)

# Transform data
tab$fdr[tab$fdr == 0] <- 1e-6
tab$neglog2fdr <- -log2(tab$fdr)

# Filter significant results
sig_tab <- tab %>% filter(fdr < 0.05) %>% arrange(desc(Odds_Ratio))
sig_tab$Gene <- factor(sig_tab$Gene, levels = rev(sig_tab$Gene))

# Dot plot
p2 <- ggplot(sig_tab, aes(x = Odds_Ratio, y = Gene)) +
  geom_point(aes(color = neglog2fdr), size = 2.5) +
  geom_errorbarh(aes(xmin = `Lower95%`, xmax = `Upper95%`), height = 0.2, color = "black") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  scale_color_gradient2(low = "white", mid = "white", high = "#0072B5FF") +
  scale_x_log10() +
  labs(x = "Odds Ratio (log scale)", y = NULL) +
  theme_classic(base_size = 14) +
  theme(legend.title = element_blank())

p2
# ggsave(p14, filename = "../03.out/figure4/09.Mutations_Fisher.pdf", width = 6.5, height = 4.1)
```

```{r fig4_fisher_forest, eval=TRUE, message=FALSE, warning=FALSE}
# Forest-style plot (alternative view)
p3 <- ggplot(sig_tab, aes(x = Gene, y = Odds_Ratio)) +
  geom_point(aes(fill = neglog2fdr, size = neglog2fdr), shape = 21) +
  geom_errorbar(aes(ymin = `Lower95%`, ymax = `Upper95%`), width = 0.2) +
  scale_y_log10() +
  scale_fill_gradient2(low = "white", mid = "white", high = "#0072B5FF") +
  labs(title = "Forest Plot of Significant Mutations", y = "Odds Ratio (log scale)", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3
# ggsave(p15, filename = "../03.out/figure4/09.Mutations_Fisher_style2.pdf", width = 8, height = 5)
```

## Multivariable Logistic Regression

```{r fig4_multilogit_forestplot, eval=TRUE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(forestmodel)
library(ggplot2)
library(ggsci)
library(scales)
library(ggbreak)

# Load data
logit_data <- read.csv("../02.data/MultivariableReg_riskScore.csv", header = TRUE, check.names = FALSE)
logit_data <- subset(logit_data, Response %in% c("CR", "NCR"))
logit_data$Response <- ifelse(logit_data$Response == "NCR", 1, 0)
logit_data$Gender <- ifelse(logit_data$Gender == "Male", 1, 0)
logit_data$isDeNovo <- ifelse(logit_data$isDeNovo == "Yes", 1, 0)
logit_data$`ELN2022 (Adv. vs. Fav._Int.)` <- ifelse(logit_data$`ELN2022 (Adv. vs. Fav._Int.)` == "Adverse", 1, 0)

# Convert binary features to numeric
binary_vars <- c('t(5;11): NUP98-NSD1', 'inv(16): CBFB-MYH11', 'Complex Karyotype', 
                 'ASXL1', 'CEBPA', 'FLT3-TKD', 'KIT', 'NPM1', 'NRAS', 
                 'PHF6', 'RUNX1', 'SF3B1', 'SRSF2', 'TP53', 'U2AF1')
for (var in binary_vars) {
  logit_data[[var]] <- ifelse(logit_data[[var]] == "Positive", 1, 0)
}

# Define forest panels
forest.panels <- list(
  list(width = 0.01),
  list(width = 0.1, display = ~variable, fontface = "bold", heading = "Variable"),
  list(width = 0.1, display = ~level),
  list(width = 0.05, display = ~n, hjust = 1, heading = "N"),
  list(width = 0.03, item = "vline", hjust = 0.5),
  list(width = 0.55, item = "forest", hjust = 0.5, heading = "Odds ratio", linetype = "dashed", line_x = 0),
  list(width = 0.03, item = "vline", hjust = 0.5),
  list(width = 0.12, display = ~ ifelse(reference, "Reference", sprintf("%0.3f (%0.3f, %0.3f)", trans(estimate), trans(conf.low), trans(conf.high))), display_na = NA),
  list(width = 0.03, display = ~ ifelse(reference, "", format.pval(p.value, digits = 1, eps = 0.001)), display_na = NA, hjust = 1, heading = "p"),
  list(width = 0.03)
)

# Forest plot function
create_forest_plot <- function(data) {
  formula <- as.formula(paste0('Response ~ ', paste(c('RiskScore', 'Age', 'Gender', 'isDeNovo', '`ELN2022 (Adv. vs. Fav._Int.)`', paste0('`', binary_vars, '`')), collapse = ' + ')))
  model <- glm(formula, data = data, family = binomial)
  forest_model(model, panels = forest.panels, factor_separate_line = FALSE)
}

# Plot for combined and individual cohorts
pdf("../03.out/figure4/10.forestplot.combined.pdf", width = 8, height = 5)
create_forest_plot(logit_data)
dev.off()

pdf("../03.out/figure4/10.forestplot.rjaml.pdf", width = 8, height = 5)
create_forest_plot(subset(logit_data, Cohort == "RJAML"))
dev.off()

pdf("../03.out/figure4/10.forestplot.beataml.pdf", width = 8, height = 5)
create_forest_plot(subset(logit_data, Cohort == "BeatAML"))
dev.off()

pdf("../03.out/figure4/10.forestplot.leucegene.pdf", width = 8, height = 5)
create_forest_plot(subset(logit_data, Cohort == "Leucegene"))
dev.off()
```

```{r fig4_multilogit_dotplot, eval=TRUE, message=FALSE, warning=FALSE}
# Summary dotplot of odds ratios by cohort
logit_summary <- data.frame(
  Cohort = factor(c("Combined", "RJAML", "BeatAML", "Leucegene"), levels = rev(c("Combined", "RJAML", "BeatAML", "Leucegene"))),
  Odds_Ratio = c(2.637, 4.257, 3.072, 2.294),
  Lower = c(2.196, 2.616, 2.369, 1.443),
  Upper = c(3.207, 7.584, 4.104, 3.891),
  Test = "Multivariate logistic regression"
)

p4 <- ggplot(logit_summary, aes(x = Odds_Ratio, y = Cohort)) +
  geom_point(aes(fill = Odds_Ratio, size = Odds_Ratio), shape = 21) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.1, color = "black") +
  scale_x_log10() +
  scale_fill_gradient2(low = "white", mid = "white", high = "#9970AB") +
  facet_wrap(. ~ Test) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 1, color = "black"),
    panel.grid.major = element_line(size = 0.1, linetype = "dashed", color = "grey"),
    panel.grid.minor = element_line(size = 0.1, linetype = "dashed", color = "grey")
  )

p4
# ggsave(p4, filename = "../03.out/figure4/13.Multivariable_plot.pdf", width = 3.5, height = 3)
```

## Differential Expression Analysis for Mut vs. WT

```{r}
library(readxl)
library(dplyr)
library(DESeq2)

# ----------------------------
# Configuration
# ----------------------------
INPUT_DIR <- "../02.data"
OUTPUT_DIR <- "../03.out"
SCRIPTS_DIR <- "../01.script"
MIN_GENE_MEAN <- 0.5  # Minimum mean expression threshold
FDR_CUTOFF <- 0.05    # Significance threshold for mutations

# ----------------------------
# Data Loading & Preprocessing
# ----------------------------

# Load DESeq2 wrapper function
source(file.path(SCRIPTS_DIR, "RunDESeq2.R"))

# Load clinical and mutation data
clinical_data <- read_xlsx(file.path(INPUT_DIR, "Combined_cohorts_riskScore_Clinical_Final.xlsx"))
mutation_results <- read.csv(
  file.path(OUTPUT_DIR, "Mutation_fisher_out.csv"), 
  header = TRUE, 
  row.names = 1,
  check.names = FALSE
)

# Filter significant mutations
significant_mutations <- mutation_results %>%
  filter(fdr < FDR_CUTOFF) %>%
  arrange(desc(Odds_Ratio)) %>%
  pull(Gene) %>%
  as.character()

# Select relevant columns from clinical data
clinical_subset <- clinical_data[, c("Patient", significant_mutations)]

# ----------------------------
# Expression Matrix Processing
# ----------------------------

# Load and clean expression matrix
expression_matrix <- read.csv(
  file.path(INPUT_DIR, "AML_multiCohorts_adjustedCounts.csv"),
  header = TRUE,
  row.names = 1
) %>%
  # Remove duplicate genes (keep first occurrence)
  filter(!duplicated(Symbol)) 

# Filter by gene type (if needed)
valid_gene_types <- c("protein_coding", "miRNA", "lincRNA", "antisense",
                     "processed_transcript", "transcribed_processed_pseudogene",
                     "transcribed_unitary_pseudogene", "transcribed_unprocessed_pseudogene",
                     "macro_lncRNA", "TEC", "3prime_overlapping_ncRNA",
                     "sense_intronic", "sense_overlapping", "unprocessed_pseudogene")

expression_matrix <- expression_matrix %>%
  dplyr::filter(Type %in% valid_gene_types) %>%
  # Set row names and remove metadata columns
  {rownames(.) <- .$Symbol; .[, -which(colnames(.) %in% c("Symbol", "Type"))]}

# Filter lowly expressed genes
expression_matrix <- expression_matrix[rowMeans(expression_matrix) > MIN_GENE_MEAN, ]

# ----------------------------
# Differential Analysis Loop
# ----------------------------

# Create output directory if needed
output_dir <- file.path(OUTPUT_DIR, "figure4", "de.out")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

# Perform DE analysis for each significant variant
for (variant in significant_mutations) {
  tryCatch({
    # Prepare and ORDER sample metadata
    sample_metadata <- clinical_subset[, c("Patient", variant)] %>%
      filter(!is.na(.[[2]]) & .[[2]] != "NA") %>%
      arrange(.[[2]])  # Orders by variant status
    
    # Get group counts
    group_counts <- table(sample_metadata[[2]])
    
    # Skip if not exactly 2 groups with sufficient samples
    if (length(group_counts) != 2 || any(group_counts < 2)) {
      warning(paste("Skipping", variant, "- insufficient samples in one or both groups"))
      next
    }
    
    # Subset expression matrix
    exp_mat_subset <- expression_matrix[, sample_metadata$Patient, drop = FALSE]
    
    # Run DESeq2
    RunDESeq2(
      count_mat = exp_mat_subset,
      n.cont = group_counts[1],
      n.treat = group_counts[2],
      prefix = file.path(output_dir, paste0("deseq_out_", variant)),
      sort.p = FALSE,
      merge.normalized = FALSE
    )
    
    message(paste("Completed analysis for", variant))
    
  }, error = function(e) {
    message(paste("Error processing", variant, ":", e$message))
  })
}
```

### Obtain the differentially expressed genes

```{r}
library(dplyr)
library(readr)
library(purrr)

# 1. Define parameters
lfc_threshold <- 1       # Absolute log2FoldChange cutoff
padj_threshold <- 0.05   # Adjusted p-value cutoff
input_dir <- "../03.out/figure4/de.out" 
output_dir <- "../03.out/figure4/filtered_DEGs"

# 2. Create output directory
if(!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
}

# 3. Get file list with full paths
file_names <- list.files(
  path = input_dir,
  pattern = "deseq_out_.*_statistic_only\\.txt$",
  full.names = TRUE
)

# Check if files were found
if(length(file_names) == 0) {
  stop("No matching files found in directory: ", input_dir)
}

# 4. Process each file
deg_results <- map(file_names, ~{
  tryCatch({
    # Read data
    df <- read_tsv(.x, show_col_types = FALSE)
    
    # Check required columns exist
    required_cols <- c("log2FoldChange", "padj")
    if(!all(required_cols %in% colnames(df))) {
      stop("Missing required columns in file: ", basename(.x))
    }
    
    # Filter and annotate
    df_filtered <- df %>%
      filter(abs(log2FoldChange) > lfc_threshold, 
             padj < padj_threshold) %>%
      arrange(desc(abs(log2FoldChange))) %>%
      mutate(variant = str_extract(basename(.x), 
                      "deseq_out_(.*?)_statistic_only", 
                      group = 1))
    
    # Save filtered results
    output_file <- file.path(output_dir, 
                           paste0("filtered_", basename(.x)))
    write_tsv(df_filtered, output_file)
    
    # Return summary
    tibble(
      file = basename(.x),
      total_genes = nrow(df),
      deg_count = nrow(df_filtered),
      up_regulated = sum(df_filtered$log2FoldChange > 0, na.rm = TRUE),
      down_regulated = sum(df_filtered$log2FoldChange < 0, na.rm = TRUE)
    )
  }, error = function(e) {
    warning("Error processing ", basename(.x), ": ", e$message)
    return(NULL)
  })
}) %>% compact() %>% bind_rows()

# 5. Save outputs
if(nrow(deg_results) > 0) {
  write_csv(deg_results, file.path(output_dir, "DEG_summary.csv"))
  message("Successfully processed ", nrow(deg_results), " files.")
} else {
  warning("No files were successfully processed.")
}
```

```{r}
library(ggplot2)
library(dplyr)
library(forcats)  # For factor manipulation

df <- read.csv("../03.out/figure4/filtered_DEGs/DEG_summary.csv", header = T)
df$file <- gsub("deseq_out_", "", df$file)
df$file <- gsub("_statistic_only.txt", "", df$file)

# Clean mutation names by removing special characters
df <- df %>%
  mutate(file = gsub("inv\\(16\\): |t\\(5;11\\): ", "", file),
         file = gsub("-", "_", file))  # Replace hyphens with underscores


# Create plot
deg_plot <- ggplot(df, aes(x = fct_reorder(file, deg_count), y = deg_count)) +
  geom_bar(stat = "identity", 
           fill = "#4E79A7",  # Professional blue shade
           width = 0.7,
           alpha = 0.8) +  # Slight transparency
  geom_text(aes(label = deg_count), 
            hjust = -0.1, 
            size = 3.5, 
            color = "black") +
  coord_flip() +  # Horizontal bars for readability
  labs(x = NULL,
       y = "Number of Differentially Expressed Genes",
       title = "Transcriptional Impact of AML Mutations",
       subtitle = "Total DEGs identified across mutations (n = 31,337 genes analyzed)") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(face = "bold", size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    axis.line.x = element_line(color = "black"),
    plot.margin = margin(1, 1, 1, 1, "cm")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Space for labels

# Print plot
print(deg_plot)

ggsave(deg_plot, filename = "../03.out/figure4/mutation_DEG_counts_simple.pdf", 
       width = 8, height = 6)

```


## Define the ARG24-associated genes

```{r}
library(dplyr)
library(readxl)
library(DESeq2)

# ----------------------------
# Configuration
# ----------------------------
INPUT_DIR <- "../02.data"
OUTPUT_DIR <- "../03.out/figure4"
MIN_TPM_THRESHOLD <- 0.5  # Minimum mean expression cutoff
TPM_FILE <- file.path(INPUT_DIR, "AML_multiCohorts_adjustedTPM.csv")
CLINICAL_FILE <- file.path(INPUT_DIR, "Combined_cohorts_riskScore_Clinical_Final.xlsx")

# ----------------------------
# Data Loading & Preprocessing
# ----------------------------

# Load and clean TPM data with error handling
tpm_data <- tryCatch({
  read.csv(TPM_FILE, header = TRUE, stringsAsFactors = FALSE) %>%
    # Remove duplicate genes (keeping highest expressed)
    arrange(desc(rowMeans(select(., -c(ID, Symbol, Type))))) %>%
    distinct(Symbol, .keep_all = TRUE) %>%
    # Filter lowly expressed genes
    filter(rowMeans(select(., -c(ID, Symbol, Type)), na.rm = TRUE) >= MIN_TPM_THRESHOLD) %>%
    # Convert to matrix format
    tibble::column_to_rownames("Symbol") %>%
    select(-c(ID, Type)) %>%
    as.matrix()
}, error = function(e) {
  stop("Failed to load TPM data: ", e$message)
})

# Load clinical data with validation
clinical_data <- tryCatch({
  read_xlsx(CLINICAL_FILE) %>%
    select(Patient, RiskScore) %>%
    filter(complete.cases(.)) %>%
    mutate(Patient = as.character(Patient))  # Ensure consistent typing
}, error = function(e) {
  stop("Failed to load clinical data: ", e$message)
})

# ----------------------------
# Data Alignment & Transformation
# ----------------------------

# Verify sample overlap
common_samples <- intersect(colnames(tpm_data), clinical_data$Patient)
if(length(common_samples) < 10) {
  warning("Low sample overlap (n = ", length(common_samples), ")")
}

# Create processed dataset
processed_data <- tpm_data[, common_samples] %>%
  t() %>%  # Transpose to samples x genes
  as.data.frame() %>%
  mutate(ARG24 = clinical_data$RiskScore[match(rownames(.), clinical_data$Patient)]) %>%
  select(ARG24, everything()) %>%
  # Log2 transformation with pseudocount
  mutate(across(-ARG24, ~ log2(.x + 1)))

# ----------------------------
# Correlation Analysis
# ----------------------------

# Parallelized correlation computation
library(future.apply)
plan(multisession)  # Enable parallel processing

cor_results <- future_lapply(
  colnames(processed_data)[-1], 
  function(gene) {
    test <- cor.test(
      x = processed_data$ARG24,
      y = processed_data[[gene]],
      method = "pearson",
      exact = FALSE
    )
    tibble(
      Gene = gene,
      rho = test$estimate,
      P.value = test$p.value
    )
  }
) %>% bind_rows() %>%
  mutate(
    FDR = p.adjust(P.value, method = "BH"),
    Direction = ifelse(rho > 0, "Positive", "Negative")
  ) %>%
  arrange(FDR, P.value)

# ----------------------------
# Output & Visualization
# ----------------------------

# Save results
write_csv(cor_results, file.path(OUTPUT_DIR, "ARG24_gene_correlations.csv"))
```

### Plot

```{r}
## Plot
library(ggplot2)
library(ggrepel)
library(ggrastr)
library(viridis)   

stat1 <- read.csv(file.path(OUTPUT_DIR, "ARG24_gene_correlations.csv"), header = T)
stat1$rank_score <- -log2(stat1$FDR) * abs(stat1$rho)
dat <- stat1[order(stat1$rank_score, decreasing = TRUE), ]
dat$Rank <- 1:nrow(stat1)
dat1 <- head(dat, 20)

arg24_genes <- head(dat, 2000)
write.csv(arg24_genes, "../03.out/figure4/arg24_genes_top2000.csv")
# Sort data by rank_score and find the top 2000 threshold
top_2000_threshold <- dat$rank_score[order(dat$rank_score, decreasing = TRUE)][2000]

# Plotting
p <- ggplot(dat, aes(x = Rank, y = rank_score, color = rank_score)) +
  geom_point_rast(aes(size = abs(rank_score)), alpha = 0.8, raster.dpi = 350, stroke = 0) +
  scale_color_viridis(option = "plasma", direction = -1, name = "Rank Score") +
  labs(x = "Rank", y = "Rank score") +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    panel.grid.major = element_line(size = 0.2, linetype = "dashed", color = "grey"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  ) +
  geom_hline(yintercept = top_2000_threshold, color = "red", linetype = "dotted", lwd = 0.8) +
  annotate("text", x = max(dat$Rank) * 0.95, y = top_2000_threshold, label = paste0("Top 2000 Genes (", round(top_2000_threshold, 2), ")"),
           vjust = -0.5, hjust = 1, color = "red", fontface = "bold", size = 4) +
  scale_size_continuous(range = c(2, 6)) # Adjust point size range for better visibility

# Adding highlighted points and labels
p1 <- p +
  geom_point(data = dat1, aes(size = abs(rank_score)), shape = 21, stroke = 1.5, color = "black", fill = "white") +
  geom_label_repel(data = dat1, aes(label = Gene), color = "black", 
                   box.padding = unit(0.4, "lines"), point.padding = unit(0.5, "lines"),
                   max.overlaps = 100, segment.color = "grey50", size = 4, fontface = "bold", fill = "lightyellow")

p1
# Save the plot
ggsave(p1, filename = "../03.out/figure4/ARG24_Correlated_genes.pdf", 
       width = 7, height = 5.3)
```

## Pathway overlap analysis of genes 

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)
library(ggpubr)
library(patchwork)
library(tidyverse)

# ----------------------------
# Configuration
# ----------------------------
input_dir <- "../03.out/figure4/filtered_DEGs"
arg24_file <- "../03.out/figure4/arg24_genes_top2000.csv"

# ----------------------------
# Data Loading Functions
# ----------------------------

# Function to safely load gene files
load_genes <- function(file_path, gene_col = "Gene") {
  if (str_detect(file_path, "\\.csv$")) {
    read_csv(file_path, show_col_types = FALSE) %>% 
      pull({{gene_col}}) %>% 
      na.omit() %>% 
      unique()
  } else if (str_detect(file_path, "\\.txt$")) {
    read_tsv(file_path, show_col_types = FALSE) %>% 
      pull({{gene_col}}) %>% 
      na.omit() %>% 
      unique()
  } else {
    stop("Unsupported file format for: ", file_path)
  }
}

# ----------------------------
# Load All Gene Sets
# ----------------------------

# Initialize list with ARG24 genes
gene_lists <- list(
  ARG24 = load_genes(arg24_file)
)

# Dynamically load all mutation DEG files
deg_files <- list.files(
  path = input_dir,
  pattern = "filtered_deseq_out_.*_statistic_only\\.(txt|csv)$",
  full.names = TRUE
)

for (file in deg_files) {
  # Extract clean mutation name
  mutation_name <- str_remove(basename(file), 
                          "filtered_deseq_out_|_statistic_only.*$") %>%
    str_replace_all("[^[:alnum:]]", "_")  # Replace special chars
  
  # Load genes (try both common column names)
  genes <- tryCatch({
    load_genes(file, "Gene")
  }, error = function(e) {
    tryCatch({
      load_genes(file, "ID")
    }, error = function(e) {
      warning("Failed to load genes from: ", file, "\nError: ", e$message)
      NULL
    })
  })
  
  if (!is.null(genes)) {
    gene_lists[[mutation_name]] <- genes
  }
}

# ----------------------------
# Quality Control & Output
# ----------------------------

# Generate summary
gene_summary <- tibble(
  GeneSet = names(gene_lists),
  N_Genes = map_int(gene_lists, length),
  Example_Genes = map_chr(gene_lists, ~paste(head(.x, 3), collapse = ", "))
)

print(gene_summary)

# Save combined list
saveRDS(gene_lists, file.path(input_dir, "combined_gene_lists.rds"))

# Map gene symbols to ENTREZ IDs
genes <- lapply(gene_lists, function(gene_set) {
  mapIds(
    org.Hs.eg.db,
    keys = gene_set,
    column = "ENTREZID",
    keytype = "SYMBOL",
    multiVals = "first"
  ) %>% na.omit() %>% unique()
})

# Obtain all gene-pathway associations from the database
all_pathway_genes <- AnnotationDbi::select(
  org.Hs.eg.db, 
  keys = keys(org.Hs.eg.db, "GO"), 
  columns = c("ENTREZID", "GO", "ONTOLOGY"), 
  keytype = "GO"
) %>% filter(!is.na(GO))

# Calculate total gene counts per pathway and filter for at least 5 genes
total_genes_per_pathway <- all_pathway_genes %>%
  group_by(GO) %>%
  summarize(Total_Genes = n_distinct(ENTREZID)) %>%
  filter(Total_Genes >= 5)

# Function to calculate overlap ratio for each gene list
calculate_overlap_ratio <- function(gene_list, pathway_genes, total_counts) {
  pathway_overlap <- pathway_genes %>%
    filter(ENTREZID %in% gene_list) %>%
    group_by(GO, ONTOLOGY) %>%
    summarize(Overlap_Count = n_distinct(ENTREZID), .groups = "drop") %>%
    left_join(total_counts, by = "GO") %>%
    mutate(Overlap_Ratio = Overlap_Count / Total_Genes) %>%
    arrange(desc(Overlap_Ratio))
  
  # Convert to data frame to avoid issues with tibble
  pathway_overlap <- as.data.frame(pathway_overlap)
  
  return(pathway_overlap)
}

# Calculate overlap ratios for each gene list
overlap_results <- lapply(names(genes), function(list_name) {
  overlap_data <- calculate_overlap_ratio(genes[[list_name]], all_pathway_genes, total_genes_per_pathway)
  
  # Safely select and rename columns
  overlap_data <- overlap_data[, c("GO", "Overlap_Ratio")]
  colnames(overlap_data)[2] <- list_name  # Dynamically assign the new column name
  
  return(overlap_data)
})


# Combine results from all gene lists into a single data frame
combined_overlap_results <- Reduce(function(x, y) merge(x, y, by = "GO", all = TRUE), overlap_results)

colnames(combined_overlap_results) <- gsub("_statistic_only_txt", "", colnames(combined_overlap_results))
# Save the results as a CSV file
write.csv(combined_overlap_results, "../03.out/figure4/pathway_overlap_analysis_result.csv", row.names = FALSE)

```

```{r}
library(corrplot)
library(Hmisc) # For rcorr() to get p-values
library(RColorBrewer)

combined_overlap_results <- read.csv("../03.out/figure4/pathway_overlap_analysis_result.csv", header = T)
cor_data <- combined_overlap_results %>%
  as_tibble() %>%
  column_to_rownames("GO") %>%
  dplyr::select(ARG24, everything()) %>%  # Explicitly use dplyr::select
  na.omit()

cor_matrix <- cor(cor_data, method = "pearson", use = "complete.obs")

arg24_cor <- cor_matrix[-1, 1, drop = FALSE] |> 
  as.data.frame() |>
  setNames("ARG24") |>
  arrange(desc(abs(ARG24)))  # Sort by absolute correlation strength

# Calculate p-values with FDR adjustment
cor_test <- rcorr(as.matrix(cor_data))
p_values <- cor_test$P[-1, 1, drop = FALSE][rownames(arg24_cor), , drop = FALSE] 

col_palette <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(200)

# pdf("../03.out/figure4/pathway_cor_colplot.pdf", width = 8, height = 5)
corrplot(
  data.matrix(arg24_cor), 
  method = "circle",
  col = colorRampPalette(brewer.pal(9, 'Blues'))(100),
  p.mat = data.matrix(p_values),
  sig.level = 0.05,
  insig = "label_sig",
  tl.col = 'black',
  col.lim = c(0.1, 0.4)
)
# dev.off()

```

```{r}
library(ggpubr)
library(ggplot2)
library(patchwork)

# Scatter plot function
plot_scatter <- function(data, x_col, y_col, color, x_label, y_label) {
  ggscatter(
    data = data,
    x = x_col,
    y = y_col,
    add = "reg.line",                      # Add regression line
    conf.int = TRUE,                       # Add confidence interval
    add.params = list(color = "black", fill = "lightgray"), # Regression line params
    color = color,                         # Scatter point color
    size = 2                               # Point size
  ) +
    stat_cor(method = "pearson") +
    labs(x = x_label, y = y_label) +       # Axis labels
    theme_classic() +                      # Use a classic theme
    theme(
      strip.text = element_text(size = 12, face = "bold"),  # Customize facet labels
      plot.title = element_text(hjust = 0.5, face = "bold"),# Center-align title
      plot.margin = margin(5, 5, 5, 5),                    # Reduce plot margins
      panel.spacing = unit(0.5, "lines")                   # Reduce spacing between facets
    )
}


# Scatter plots between different gene overlap ratios
p1 <- plot_scatter(cor_data, "PHF6", "ARG24", "#1B9E77",
                   "% Pathway Overlap \n(PHF6 mutation-affected genes)", 
                   "% Pathway Overlap (ARG24-correlated genes)")

p2 <- plot_scatter(cor_data, "CEBPA", "ARG24", "#7570B3",
                   "% Pathway Overlap \n(CEBPA mutation-affected genes)", 
                   "% Pathway Overlap (ARG24-correlated genes)")

p3 <- plot_scatter(cor_data, "TP53", "ARG24", "#D95F02",
                   "% Pathway Overlap \n(TP53 mutation-affected genes)", 
                   "% Pathway Overlap (ARG24-correlated genes)")

plots <- p1 | p2 | p3
plots

ggsave(plots, filename = "../03.out/figure4/top_pathway_overlap.pdf", width = 10.8, height = 4)

```

