# Figure 5: ARG24 score is associated with cellular hierarchy and predicts AML relapse 

Zeng et. al. (Nat Med, 2022) recently described a cellular hierarchy of AML leukemic cells representing distinct maturation states including LSPC-Quiescent, LSPC-Primed, LSPC-Cycle, GMP-like, ProMono-like, Mono-like and cDC-like. We investigated ARG24 score across these AML hierarchies and found that it was highest among cells in the HSPC-like state. Zeng et al. also revealed 4 distinct subtypes in bulk cohort based on the composition of their leukemia cell hierarchy. We projected the ARG24 scores on this map and validated the ARG24 scores in these 4 distinct subtypes, and it was highest in the Primitive state while lowest in GMP states.  

## ARG24 Scores Across AML Cellular Hierarchies (Single-cell)

```{r fig5_scaml_arg24_umap, eval=TRUE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(Seurat)
library(tidyr)
library(corrplot)
library(RColorBrewer)

# Load Seurat object
scaml <- readRDS("../02.data/scaml_Diag.rds")

# Replace UMAP coordinates with Zeng UMAPs
umap1 <- scaml@meta.data$Zeng_UMAP1
umap2 <- scaml@meta.data$Zeng_UMAP2
names(umap1) <- names(umap2) <- rownames(scaml@meta.data)
scaml@reductions[["umap"]]@cell.embeddings[, 1] <- umap1
scaml@reductions[["umap"]]@cell.embeddings[, 2] <- umap2

# Set identities and colors
Idents(scaml) <- 'Zeng_CellType'
my36colors <- c(
  '#E5D2DD', '#53A85F', '#F1BB72', '#F3B1A0', '#D6E7A3', '#57C3F3', '#476D87',
  '#E95C59', '#E59CC4', '#AB3282', '#23452F', '#BD956A', '#8C549C', '#585658', 
  '#9FA3A8', '#E0D4CA', '#5F3D69', '#58A4C3', "#b20000",'#E4C755', '#F7F398',
  '#AA9A59', '#E63863', '#E39A35', '#C1E6F3', '#6778AE', '#91D0BE', '#B53E2B',
  '#712820', '#DCC1DD', '#CCE0F5', '#CCC9E6', '#625D9E', '#68A180', '#3A6963',
  '#968175')

# UMAP plot by Zeng cell type
# pdf("../03.out/figure5/01.scaml_umap.pdf", width = 5.8, height = 4)
DimPlot(scaml, group.by = "Zeng_CellType", label = TRUE, repel = TRUE, raster = FALSE, cols = my36colors)
# dev.off()
```

```{r fig5_scaml_arg24_score, eval=TRUE, message=FALSE, warning=FALSE}
# Define ARG24 risk score calculation function

risk_score <- function(dat, candidate_genes, lasso_report) {
  genes <- intersect(colnames(dat), candidate_genes)
  risk_score_table <- dat[, genes]
  for (g in genes) {
    risk_score_table[, g] <- as.matrix(risk_score_table[, g]) * lasso_report[g, 1]
  }
  risk_scores <- rowSums(risk_score_table)
  data.frame(ARG24 = risk_scores)
}

# Calculate ARG24 score from expression
gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE, row.names = 1)
intersected_genes <- intersect(rownames(scaml), rownames(gene.min))
sc_expr <- as.data.frame(scaml@assays$RNA@data[intersected_genes, ])
sc_expr_t <- data.frame(t(sc_expr))
arg24_scores <- risk_score(sc_expr_t, intersected_genes, gene.min)
scaml <- AddMetaData(scaml, metadata = arg24_scores)

```

```{r fig5_scaml_arg24_vln_corrplot, eval=TRUE, message=FALSE, warning=FALSE}
# Violin plot of ARG24 scores across cell types
VlnPlot(scaml, features = "ARG24")

# Mean ARG24 scores by cell type
mean_arg24 <- aggregate(scaml@meta.data$ARG24, by = list(scaml@meta.data$Zeng_CellType), FUN = mean)
scaled_arg24 <- data.frame(ARG24 = scale(mean_arg24$x))
rownames(scaled_arg24) <- mean_arg24$Group.1

# Correlation circle plot
rdbu_palette <- rev(colorRampPalette(brewer.pal(11, 'RdBu'))(100))
# pdf("../03.out/figure5/02.scaml_corplot.pdf", width = 5.8, height = 6)
corrplot(data.matrix(scaled_arg24), method = "circle", tl.cex = 1, is.corr = FALSE, tl.col = 'black', cl.ratio = 0.75, col = rdbu_palette)
# dev.off()
```

## ARG24 Scores and AML Hierarchy in Bulk Cohorts

```{r fig5_bulk_pca_violin, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)
library(ggplot2)
library(ggbeeswarm)
library(RColorBrewer)

# Load bulk data with ARG24 and hierarchy info
dat <- read.csv("../02.data/2022NM_cellAbundances.csv", header = TRUE)
dat$Cluster <- factor(dat$Cluster, levels = c("Primitive", "GMP", "Intermediate", "Mature"))
dat$RiskScore[which(dat$RiskScore >= 5)] <- 5

# PCA plot by Cluster
p0 <- ggplot(dat, aes(x = PC1, y = PC2, fill = Cluster)) +
  geom_point(size = 2, shape = 21, alpha = 0.7) +
  theme_classic() +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "top", axis.line = element_line(size = 0.75, colour = "black"))

# PCA plot by RiskScore
p1 <- ggplot(dat, aes(x = PC1, y = PC2, fill = RiskScore)) +
  geom_point(size = 2, shape = 21, alpha = 0.7) +
  theme_classic() +
  scale_fill_gradient2(low = "#74ADD1", mid = "white", high = "#F46D43") +
  theme(legend.position = "top", axis.line = element_line(size = 0.75, colour = "black"))

# Violin plot by subtype
p2 <- ggplot(dat, aes(x = Cluster, y = RiskScore, color = Cluster)) +
  geom_violin() +
  geom_boxplot(width = 0.4, alpha = 0.85) +
  geom_quasirandom(width = 0.25, alpha = 0.25) +
  theme_pubr() +
  xlab("") + ylab("Risk score") +
  theme(strip.text.x = element_text(size = 13), legend.position = 'None') +
  stat_compare_means(comparisons = list(
    c('Primitive', 'GMP'),
    c('Primitive', 'Intermediate'),
    c('Primitive', 'Mature'),
    c('GMP', 'Intermediate'),
    c('GMP', 'Mature'),
    c('Intermediate', 'Mature')
  ), method = 'wilcox.test') +
  scale_color_brewer(palette = "Set2")

# Combine plots
p_combined <- ggarrange(p0, p1, p2, ncol = 3, nrow = 1)

# pdf("../03.out/figure5/03.bulk_pca.pdf", width = 10, height = 4)
p_combined
# dev.off()
```

## ARG24 Score Dynamics from Diagnosis to Relapse

```{r , eval=TRUE, message=FALSE, warning=FALSE}
# Diagnosis-relapse paired analysis
library(readxl)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(ggrastr)
library(RColorBrewer)

meta_info <- read_xlsx("../02.data/bulk_diag_r-r_clinical.xlsx")
gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE)
colnames(gene.min)[1] <- "Symbol"

# Define risk score function
risk_score <- function(dat, candidate_genes, lasso_report) {
  genes <- intersect(colnames(dat), candidate_genes)
  risk_score_table <- dat[, genes]
  for (g in genes) {
    risk_score_table[, g] <- as.matrix(risk_score_table[, g]) * lasso_report[g, 1]
  }
  data.frame(risk_score = rowSums(risk_score_table))
}

# Preprocess gene expression matrix
mat_132 <- read.csv("../02.data/bulk_diag_relapse_batchRemoved_tpm_132.csv", header = TRUE, row.names = 1)
mat_132 <- merge(gene.min, mat_132, by = "Symbol")
rownames(mat_132) <- mat_132$Symbol
mat_132 <- mat_132[, -c(1:3)]
mat_132 <- data.frame(t(mat_132))
colnames(mat_132)[9] <- "INHBA-AS1"
colnames(mat_132)[16:17] <- c("RP11-158I9.8", "RP5-1119A7.14")
colnames(mat_132)[24] <- "UNC5B-AS1"

# Compute risk scores
genes <- colnames(mat_132)
gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE, row.names = 1)
mat_132_RiskScores <- risk_score(mat_132, genes, gene.min)
mat_132_RiskScores$Sample <- rownames(mat_132_RiskScores)
mat_132_RiskScores <- merge(mat_132_RiskScores, meta_info, by = "Sample")
# Save the risk scores (optional)
# write.csv(mat_132_RiskScores, "../02.data/mat_132_RiskScores.csv")
```

```{r fig5_dx_relapse_validation, eval=TRUE, message=FALSE, warning=FALSE}
mat_132_RiskScores <- read.csv("../02.data/mat_132_RiskScores.csv")

# Violin plot comparing timepoints
p1 <- ggplot(mat_132_RiskScores, aes(x = timepoint, y = risk_score)) +
  geom_violin(aes(fill = timepoint), alpha = 0.35) +
  geom_boxplot(aes(fill = timepoint), width = 0.3, alpha = 0.5) +
  geom_point(aes(color = timepoint), size = 2, shape = 19, alpha = 0.95) +
  geom_line(aes(group = Patient), color = 'gray', lwd = 0.5) +
  scale_fill_manual(values = c("#5AAE61", "#9970AB")) +
  scale_color_brewer(palette = "Accent") +
  theme_pubr() +
  labs(x = '', y = 'ARG24 score') +
  stat_compare_means(comparisons = list(c('Diagnosis', 'Relapse')), paired = TRUE)

# Compare change direction
data <- mat_132_RiskScores[, c("Patient", "timepoint", "risk_score")]
dat <- dcast(Patient ~ timepoint, data = data)
dat$Relap_Diag <- dat$Relapse - dat$Diagnosis
dat$type <- ifelse(dat$Relap_Diag > 0, "Increased", "Decreased")

# Diagnosis score by relapse change type
p2 <- ggplot(dat, aes(x = type, y = Diagnosis, color = type)) +
  geom_violin() +
  geom_boxplot_jitter(width = 0.4, alpha = 0.75) +
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  theme_pubr() +
  labs(x = '', y = 'Risk score') +
  stat_compare_means(comparisons = list(c('Increased', 'Decreased')), method = 'wilcox.test') +
  scale_color_manual(values = c("#5AAE61", "#9970AB"))

# Combine and save
plots <- p1 + p2
plots
# ggsave(plots, filename = "../03.out/figure5/04.Dx_Rel_violin.pdf", width = 6, height = 3.5)
```

## ARG24 Score Predicts AML Relapse and CIR

```{r fig5_km_efs_cir, eval=TRUE, message=FALSE, warning=FALSE}
# EFS analysis
library(readxl)
library(survival)
library(survminer)

mat <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
mat$group <- ifelse(mat$risk_score >= median(mat$risk_score), "High ARG24", "Low ARG24")
surv_obj <- Surv(mat$EFS_Time, mat$EFS_Status)
surv_fit <- survfit(surv_obj ~ mat$group)
surv_diff <- survdiff(surv_obj ~ mat$group)
p_val <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n) - 1)
hr <- (surv_diff$obs[1] / surv_diff$exp[1]) / (surv_diff$obs[2] / surv_diff$exp[2])
hr_ci_low <- exp(log(hr) - qnorm(0.975) * sqrt(1 / surv_diff$exp[2] + 1 / surv_diff$exp[1]))
hr_ci_high <- exp(log(hr) + qnorm(0.975) * sqrt(1 / surv_diff$exp[2] + 1 / surv_diff$exp[1]))

p_label <- ifelse(p_val < 0.001, "p < 0.001", paste("p =", round(p_val, 3)))
plot1 <- ggsurvplot(
  surv_fit, data = mat, palette = "nejm",
  pval = paste(p_label, paste("HR =", round(hr, 2)), paste("95% CI:", round(hr_ci_low, 2), "-", round(hr_ci_high, 2)), sep = "\n"),
  pval.size = 5, xlab = "Survival time in days",
  conf.int = FALSE, risk.table = FALSE,
  legend.title = "RJAML cohort",
  legend.labs = c(
    paste0("High ARG24 (N = ", sum(mat$group == "High ARG24"), ")"),
    paste0("Low ARG24 (N = ", sum(mat$group == "Low ARG24"), ")")
  ), legend = c(0.65, 0.85)
)

# pdf("../03.out/figure5/04.efs.km.pdf", width = 4.2, height = 4)
plot1
# dev.off()
```

```{r fig5_km_cir, eval=TRUE, message=FALSE, warning=FALSE}
library(cmprsk)
df <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
df$group <- ifelse(df$risk_score >= median(df$risk_score), "High ARG24", "Low ARG24")
df <- df[!is.na(df$CIR_time) & !is.na(df$CIR_status), ]
ci <- cuminc(ftime = df$CIR_time, fstatus = df$CIR_status, group = as.factor(df$group))
# pdf("../03.out/figure5/05.cir.km.pdf", width = 4, height = 4.4)
plot(ci, xlab = "Time (days)", ylab = "Cumulative incidence of relapse", col = 1:length(ci), lty = 1:length(ci))
# dev.off()
```

## Validation in Independent Cohort (phs001027)

```{r fig5_rfs_phs001027, eval=TRUE, message=FALSE, warning=FALSE}
# Relapse-free survival in phs001027
library(readxl)
dat <- read_xlsx("../02.data/phs001027_clinical.xlsx")
dat$group <- ifelse(dat$risk_score >= median(dat$risk_score), "High", "Low")
plot1 <- ggsurvplot(
  survfit(Surv(Time_to_Relapse, Relapse) ~ group, data = dat),
  palette = "nejm", pval.size = 5, pval = TRUE,
  xlab = "Survival time in days", conf.int = FALSE,
  risk.table = FALSE, surv.median.line = 'hv',
  legend.title = "",
  legend.labs = c(
    paste0("High ARG24 (N = ", sum(dat$group == "High"), ")"),
    paste0("Low ARG24 (N = ", sum(dat$group == "Low"), ")")
  ), legend = c(0.65, 0.85)
)

# pdf("../03.out/figure5/07.phs001027_rfs.pdf", width = 4.2, height = 4)
plot1
# dev.off()
```

## ARG24 Score Distinguishes Relapse in Treatment-naive AML

```{r fig5_violin_relapse_vs_nonrelapse, eval=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggpubr)
library(ggbeeswarm)
library(ggrastr)
library(patchwork)
library(RColorBrewer)
library(readxl)

# RJAML cohort
mat <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
mat$Relapse <- ifelse(mat$Relapse == "1", "Relapse", "Non-relapse")

# Plot for RJAML cohort
p1 <- ggplot(mat, aes(x = Relapse, y = risk_score, color = Relapse)) + 
  geom_violin() +
  geom_boxplot(width = 0.4, alpha = 0.75) + 
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  theme_pubr() + 
  xlab("") + 
  ylab("ARG24 score") +
  theme(strip.text.x = element_text(size = 13)) +
  stat_compare_means(comparisons = list(c('Relapse', 'Non-relapse')), method = 'wilcox.test') + 
  scale_color_manual(values = c("#5AAE61", "#9970AB")) + 
  theme(legend.position = 'None')

# GSE165430 cohort (CN-AML)
dat <- read.csv("../02.data/GSE165430_268_riskScore.csv", header = TRUE)
dat$type <- ifelse(dat$type == "relapse", "Relapse", "Non-relapse")

# Plot for GSE165430 cohort
p2 <- ggplot(dat, aes(x = type, y = risk_score, color = type)) + 
  geom_violin() +
  geom_boxplot(width = 0.4, alpha = 0.75) + 
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  theme_pubr() + 
  xlab("") + 
  ylab("ARG24 score") +
  theme(strip.text.x = element_text(size = 13)) +
  stat_compare_means(comparisons = list(c('Relapse', 'Non-relapse')), method = 'wilcox.test') +   
  scale_color_manual(values = c("#5AAE61", "#9970AB")) + 
  theme(legend.position = 'None')

# Combine and save
combined_plot <- p1 + p2
combined_plot
ggsave(combined_plot, filename = "../03.out/figure5/06.Dx_R_NR_violin.pdf", width = 5, height = 3.5)
```

## Multivariable Cox Model for Predicting EFS

```{r fig5_multivariable_cox_efs, eval=TRUE, message=FALSE, warning=FALSE}
library(forestmodel)
library(readxl)
library(survival)

# Load and preprocess data
dat <- read_xlsx("../02.data/0_RJAML_retained_clinical.xlsx")
dat$`ARG24 score` <- dat$risk_score
dat$`D5-PBCR` <- ifelse(dat$`D5-PBCR` == 'Pos', 1, 0)
dat$`ELN2022 (Adv. vs. Fav._Int.)` <- ifelse(dat$ELN2022 == 'Adverse', 1, 0)

# Define forest plot layout
forest.panels <- list(
  list(width = 0.01),
  list(width = 0.1, display = ~variable, fontface = "bold", heading = "Variable"),
  list(width = 0.1, display = ~level),
  list(width = 0.03, item = "vline", hjust = 0.5),
  list(width = 0.55, item = "forest", hjust = 0.5, heading = "Hazard ratio", linetype = "dashed", line_x = 0),
  list(width = 0.03, item = "vline", hjust = 0.5),
  list(width = 0.12, display = ~ ifelse(reference, "Reference", sprintf("%0.3f (%0.3f, %0.3f)", trans(estimate), trans(conf.low), trans(conf.high))), display_na = NA),
  list(width = 0.03, display = ~ ifelse(reference, "", format.pval(p.value, digits = 1, eps = 0.001)), display_na = NA, hjust = 1, heading = "p"),
  list(width = 0.03)
)

# Fit Cox model for EFS
efs_formula <- as.formula('Surv(EFS_Time, EFS_Status) ~ `ARG24 score` + Age + Gender + HSCT + `D5-PBCR` + `ELN2022 (Adv. vs. Fav._Int.)` + `BM Blast` + WBC + HB + PLT + NPM1 + `FLT3-ITD`')
efs_model <- coxph(efs_formula, data = dat)

# pdf("../03.out/figure5/10.Multivariable_EFS.pdf", width = 7, height = 4)
print(forest_model(efs_model, panels = forest.panels, factor_separate_line = FALSE))
# dev.off()
```

## ARG24 Score and Evolutionary Mutational Combinations in Relapse

ARG24 score is a common hallmark across diverse evolutionary paths to AML relapse after chemotherapy.
The change of ARG24 score from diagnosis to relapse can also help to contextualize patterns of clonal evolution in AML. For example, in the context of NPM1-mutant AML, FLT3-ITD alterations are recurrently gained at relapse, whereas NRAS and FLT3-TKD alterations are recurrently lost at relapse. Data from Christopher MJ et.al N Engl J Med. 2018.

```{r fig5_mutational_combo_density_boxplot, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
library(ggpubr)
library(readr)
library(ggrastr)
library(ggsci)

# Load data
dat <- read_csv("../02.data/Two_mutations_combination_riskScore_final.csv") %>%
  select(c("RiskScore", "NPM1 + FLT3-ITD", "NPM1 + FLT3-TKD", "NPM1 + NRAS"))

alteration <- dat %>%
  gather(key = 'Alteration', value = 'Status', -RiskScore) %>%
  filter(Status > 0)

# Ridge density plot
p1 <- alteration %>%
  ggplot(aes(y = fct_reorder(Alteration, RiskScore), x = RiskScore, fill = stat(x))) +
  geom_density_ridges_gradient(
    jittered_points = TRUE, scale = 1.25,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.3
  ) +
  scale_fill_gradient2(low = "#01665E", high = "#8C510A", midpoint = median(alteration$RiskScore), name = "") +
  theme_pubr() +
  ylab('') +
  xlab('Risk score') +
  theme(
    legend.position = 'right',
    axis.title.x = element_text(size = 15)
  )
p1
# ggsave(p1, filename = "../03.out/figure5/11.NPM1_FLT3_NRAS_arg24.pdf", width = 6, height = 3)

# Boxplot
alteration$Alteration <- factor(alteration$Alteration, levels = c("NPM1 + FLT3-ITD", "NPM1 + NRAS", "NPM1 + FLT3-TKD"))
col_val <- pal_nejm(alpha = 0.7)(8)

p2 <- ggplot(data = alteration, aes(x = Alteration, y = RiskScore, color = Alteration)) +
  geom_violin(alpha = 0.6, width = 1, position = position_dodge(width = 0.8), size = 0.75) +
  geom_jitter_rast(alpha = 0.6, position = position_jitterdodge(jitter.width = 0.5, jitter.height = 0, dodge.width = 1)) +
  geom_boxplot(alpha = 0.6, width = 0.2, position = position_dodge(width = 0.8), size = 0.75, outlier.colour = NA) +
  scale_color_manual(values = col_val[4:6]) +
  theme_classic() +
  theme(
    legend.position = "none",
    text = element_text(size = 16),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1)
  ) +
  stat_compare_means(label = "p.format", comparisons = list(c("NPM1 + FLT3-ITD", "NPM1 + NRAS"),
                                                            c("NPM1 + FLT3-ITD", "NPM1 + FLT3-TKD")), method = 't.test') +
  ylab("Risk score") +
  xlab("")
p2
# ggsave(p2, filename = "../03.out/figure5/11.NPM1_FLT3_NRAS_arg24_box.pdf", width = 4, height = 4)
```

## Clonal Dynamics from Diagnosis to Relapse (Fishplots)

```{r fig5_fishplot_196371, eval=TRUE, message=FALSE, warning=FALSE}
library(rjson)
library(devtools)
library(fishplot)

# Patient 196371
frac.table.CCF = matrix(
  c(100, 96.6, 44.1, 3.6, 24.6,
    100, 96.7, 68.5, 59.9, 1.7),
  ncol = 2)
parents = c(0, 1, 2, 3, 3)
timepoints = c(1, 2)
fish = createFishObject(frac.table.CCF, parents, timepoints = timepoints)
fish = layoutClones(fish)
fish = setCol(fish, c("#888888", "#EF0000", "#FF6000", "#7B1732", "#F2B33F"))

# pdf("../03.out/figure5/12.fishplot_196371.pdf", width = 5, height = 3.4)
fishPlot(fish, shape = "spline", title.btm = "Christopher et al\nPt 196371",
         cex.title = 0.5, vlines = c(1, 2), bg.type = 'solid', bg.col = 'white',
         vlab = c("Diagnosis", "Relapse"))
# dev.off()
```

```{r fig5_fishplot_327733, eval=TRUE, message=FALSE, warning=FALSE}
# Patient 327733
frac.table.CCF = matrix(
  c(100, 80, 73.8, 70.2, 4.5,
    100, 92.8, 46.6, 45.7, 30.9),
  ncol = 2)
parents = c(0, 1, 2, 3, 4)
fish = createFishObject(frac.table.CCF, parents, timepoints = timepoints)
fish = layoutClones(fish)

# pdf("../03.out/figure5/12.fishplot_327733.pdf", width = 5, height = 3.4)
fishPlot(fish, shape = "spline", title.btm = "Christopher et al\nPt 327733",
         cex.title = 0.3, vlines = c(1, 2), bg.type = 'solid', bg.col = 'white',
         vlab = c("Diagnosis", "Relapse"))
# dev.off()
```

## ARG24 Score Trajectories in Selected Patients (Dx vs Relapse)

```{r fig5_dx_rel_patient_trajectories, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)
dat <- read.csv("../02.data/Christopher_2018_NEJM_Dx_Rel.csv", header = TRUE, check.names = FALSE)
dat <- dat[!duplicated(dat$Gene), ]
row.names(dat) <- dat$Gene
dat <- dat[, -1]
dat <- log2(dat + 1)
gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE, row.names = 1)
inter_gene <- intersect(row.names(dat), row.names(gene.min))
dat <- dat[inter_gene, ]
dat <- t(dat)
gene.min <- data.frame(gene.min[inter_gene, ])
row.names(gene.min) <- colnames(dat)
colnames(gene.min) <- "Coef"

calculate_risk_scores <- function(features, coefficients) {
  scores <- sweep(features, 2, coefficients, `*`)
  risk_scores <- rowSums(scores)
  data.frame(risk_score = risk_scores)
}

dx_rel_arg24 <- calculate_risk_scores(dat, gene.min[, 1])
dx_rel_arg24$sample <- row.names(dx_rel_arg24)
dx_rel_arg24 <- separate(dx_rel_arg24, col = "sample", into = c("Patient", "Timepoint"))
dx_rel_arg24$Timepoint <- gsub("primary", "Diagnosis", dx_rel_arg24$Timepoint)
dx_rel_arg24$Timepoint <- gsub("post", "Relapse", dx_rel_arg24$Timepoint)

dx_rel_arg24$Patient <- factor(dx_rel_arg24$Patient, levels = c(196371, 327733, 150288, 303642))
dx_rel_arg24$Timepoint <- factor(dx_rel_arg24$Timepoint, levels = c("Diagnosis", "Relapse"))

p1 <- ggplot(dx_rel_arg24, aes(x = Timepoint, y = risk_score, group = Patient, color = Patient)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Patient, scales = "free") +
  labs(x = "Timepoint", y = "Risk Score") +
  theme_minimal()
p1

# ggsave(p1, filename = "../03.out/figure5/13.4patients_Dx_Rel.pdf", width = 6, height = 4)
```

## ARG24 Score Predicts AML Overall Survival (Multi-cohort)

```{r fig5_multicohort_km_dotplot, eval=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(survival)
library(survminer)
library(ggplot2)
library(ggsci)
library(scales)
library(ggbreak)

# Load data
dat <- read_xlsx("../02.data/Survival_multicohorts.xlsx")
dat <- dat[, c("sample", "RiskScore", "Cohort", "OS_Time", "OS_Status")]

cohorts <- c("RJAML", "GSE146173", "BeatAML", "GSE165656", "Leucegene",  "HOVON", "TCGA-LAML", "GSE37642")
plots <- list()

for (i in seq_along(cohorts)) {
  cohort_data <- subset(dat, Cohort == cohorts[i])
  cohort_data$group <- ifelse(cohort_data$RiskScore >= median(cohort_data$RiskScore), 'High', 'Low')
  surv_obj <- Surv(cohort_data$OS_Time, cohort_data$OS_Status)
  surv_fit <- survfit(surv_obj ~ cohort_data$group)
  surv_diff <- survdiff(surv_obj ~ cohort_data$group)
  p_val <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n) - 1)
  hr <- (surv_diff$obs[1] / surv_diff$exp[1]) / (surv_diff$obs[2] / surv_diff$exp[2])
  hr_ci_low <- exp(log(hr) - qnorm(0.975) * sqrt(1 / surv_diff$exp[2] + 1 / surv_diff$exp[1]))
  hr_ci_high <- exp(log(hr) + qnorm(0.975) * sqrt(1 / surv_diff$exp[2] + 1 / surv_diff$exp[1]))
  hr_label <- paste("Hazard Ratio = ", round(hr, 2))
  ci_label <- paste("95% CI: ", round(hr_ci_low, 2), " - ", round(hr_ci_high, 2))
  p_val_label <- paste("p =", signif(p_val, 2))
  plots[[i]] <- ggsurvplot(
    surv_fit,
    data = cohort_data,
    palette = "nejm",
    pval = paste(p_val_label, hr_label, ci_label, sep = "\n"),
    pval.size = 5,
    xlab = "Survival time in days",
    conf.int = FALSE,
    risk.table = FALSE,
    legend.title = paste(cohorts[i], "cohort"),
    legend.labs = c(
      paste0("High (N = ", sum(cohort_data$group == "High"), ")"),
      paste0("Low (N = ", sum(cohort_data$group == "Low"), ")")
    ),
    legend = c(0.65, 0.85)
  )
}

p <- arrange_ggsurvplots(plots, print = FALSE, ncol = 4, nrow = 2)

# pdf("../03.out/figure5/11.km_plots.pdf", width = 16, height = 8)
p
# dev.off()

# Dot plot summary
library(dplyr)
tab <- data.frame(
  Cohort = c("RJAML", "BeatAML", "Leucegene", "TCGA-LAML", "GSE146173", "GSE165656", "HOVON", "GSE37642"),
  Hazard_Ratio = c(5.59, 2.17, 2.47, 2.13, 1.98, 2.15, 1.33, 1.8),
  Pvalue = c(1.45e-11, 7.02e-10, 1.30e-05, 4.26e-05, 2.45e-05, 0.00083, 0.0033, 0.0034),
  Lower = c(3.51, 1.69, 1.63, 1.48, 1.43, 1.33, 1.1, 1.21),
  Upper = c(8.92, 2.79, 3.76, 3.07, 2.75, 3.48, 1.62, 2.68)
)
tab$Cohort <- factor(tab$Cohort, levels = rev(tab$Cohort))
tab$Test <- "Overall survival"

p1 <- ggplot(tab, aes(x = Hazard_Ratio, y = Cohort)) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.1, color = "black") + 
  geom_point(aes(fill = -log2(Pvalue), size = Hazard_Ratio), shape = 21) +
  scale_x_log10() +
  scale_fill_gradient(low = "white", high = "#8C510A") +
  facet_wrap(. ~ Test) + 
  theme_classic() +
  labs(x = "Hazards Ratio", y = "") +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 1, color = "black"),
    panel.grid.major = element_line(size = 0.1, linetype = "dashed", color = "grey"),
    panel.grid.minor = element_line(size = 0.1, linetype = "dashed", color = "grey")
  )

p1

# ggsave(p1, filename = "../03.out/figure5/12.Multicohorts_km.pdf", width = 3.8, height = 3)
```

## Multivariable OS analysis across AML cohorts

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(survival)
library(survminer)
library(ggplot2)
library(ggsci)
library(scales)
library(ggbreak)
library(forestmodel)

# Load and prepare data
dat <- read_xlsx("../02.data/Survival_multicohorts.xlsx")
colnames(dat)[33] <- "FLT3.ITD"  # Ensure consistent naming
dat$Gender <- ifelse(dat$Gender == "Male", 1, 0)  # Encode Gender

# Define cohorts to be analyzed
cohorts <- c("RJAML", "GSE146173", "BeatAML", "GSE165656", "Leucegene", "HOVON", "TCGA-LAML", "GSE37642")

# Define forest plot layout
forest.panels <- list(
  list(width = 0.01),
  list(width = 0.1, display = ~variable, fontface = "bold", heading = "Variable"),
  list(width = 0.1, display = ~level),
  list(width = 0.03, item = "vline", hjust = 0.5),
  list(width = 0.55, item = "forest", hjust = 0.5, heading = "Hazard Ratio", linetype = "dashed", line_x = 0),
  list(width = 0.03, item = "vline", hjust = 0.5),
  list(width = 0.12, display = ~ifelse(reference, "Reference",
                                       sprintf("%0.3f (%0.3f, %0.3f)", trans(estimate), trans(conf.low), trans(conf.high))),
       display_na = NA),
  list(width = 0.03, display = ~ifelse(reference, "", format.pval(p.value, digits = 1)),
       display_na = NA, hjust = 1, heading = "p"),
  list(width = 0.03)
)

# Wrapper function for generating forest plots
create_forest_plot <- function(data, cohort, binary_vars) {
  formula_str <- paste0("Surv(OS_Time, OS_Status) ~ ", paste(c("RiskScore", "Age", "Gender", binary_vars), collapse = " + "))
  model <- coxph(as.formula(formula_str), data = data)
  pdf_filename <- paste0("../03.out/figure5/13.forestplot.os_", tolower(gsub("-", "_", cohort)), ".pdf")
  pdf(pdf_filename, width = 7, height = 4)
  print(forest_model(model, panels = forest.panels, factor_separate_line = FALSE))
  dev.off()
}

# Cohort-specific modeling
## RJAML
dat_rjaml <- subset(dat, Cohort == "RJAML")
create_forest_plot(dat_rjaml, "RJAML", binary_vars = c("Blast", "WBC", "NPM1", "FLT3.ITD"))

## BeatAML
dat_beataml <- subset(dat, Cohort == "BeatAML")
dat_beataml$isDeNovo <- ifelse(dat_beataml$isDeNovo == "Yes", 1, 0)
create_forest_plot(dat_beataml, "BeatAML", binary_vars = c("Blast", "WBC", "isDeNovo", "NPM1", "FLT3.ITD"))

## Leucegene
dat_leucegene <- subset(dat, Cohort == "Leucegene")
create_forest_plot(dat_leucegene, "Leucegene", binary_vars = c("Blast", "NPM1", "FLT3.ITD"))

## TCGA-LAML
dat_tcga <- subset(dat, Cohort == "TCGA-LAML")
create_forest_plot(dat_tcga, "TCGA-LAML", binary_vars = c("Blast", "WBC", "NPM1", "FLT3.ITD"))

## GSE146173
dat_gse146173 <- subset(dat, Cohort == "GSE146173")
dat_gse146173$isDeNovo <- ifelse(dat_gse146173$isDeNovo == "Yes", 1, 0)
dat_gse146173$`ELN2017 (Adv. vs. Fav._Int.)` <- ifelse(dat_gse146173$ELN2017 == "Adverse", 1, 0)
create_forest_plot(dat_gse146173, "GSE146173", binary_vars = c("isDeNovo", "`ELN2017 (Adv. vs. Fav._Int.)`"))

## GSE165656
dat_gse165656 <- subset(dat, Cohort == "GSE165656")
dat_gse165656$isDeNovo <- ifelse(dat_gse165656$isDeNovo == "Yes", 1, 0)
dat_gse165656$`ELN2017 (Adv. vs. Fav._Int.)` <- ifelse(dat_gse165656$ELN2017 == "Adverse", 1, 0)
create_forest_plot(dat_gse165656, "GSE165656", binary_vars = c("WBC"))

## HOVON
dat_HOVON <- subset(dat, Cohort == "HOVON")
dat_HOVON$`Cytogenetic (Poor vs. Good_Int.)` <- ifelse(dat_HOVON$Cytogenetic_risk == "Poor", 1, 0)
create_forest_plot(dat_HOVON, "HOVON", binary_vars = c("`Cytogenetic (Poor vs. Good_Int.)`", "FAB"))

#-----------------------------------------
# Summary forest dot plot across cohorts
#-----------------------------------------
summary_tab <- data.frame(
  Cohort = c("RJAML", "BeatAML", "Leucegene", "TCGA-LAML", "GSE146173", "GSE165656"),
  Hazard_Ratio = c(1.614, 1.337, 1.539, 1.156, 1.199, 1.553),
  Pvalue = c(2e-14, 3e-08, 2e-06, 0.005, 0.001, 0.002),
  Lower = c(1.428, 1.207, 1.290, 1.044, 1.076, 1.175),
  Upper = c(1.825, 1.481, 1.835, 1.280, 1.336, 2.053)
)
summary_tab$Cohort <- factor(summary_tab$Cohort, levels = rev(summary_tab$Cohort))
summary_tab$Test <- "Multivariable overall survival"

# Plot summary dot plot
p1 <- ggplot(summary_tab, aes(x = Hazard_Ratio, y = Cohort)) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.1, color = "black") +
  geom_point(aes(fill = -log2(Pvalue), size = Hazard_Ratio), shape = 21) +
  scale_x_log10() +
  scale_fill_gradient(low = "white", high = "#01665E") +
  facet_wrap(. ~ Test) +
  theme_classic() +
  labs(x = "Hazard Ratio", y = "") +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    axis.text.x = element_text(color = "black"),
    panel.grid.major = element_line(size = 0.1, linetype = "dashed", color = "grey"),
    panel.grid.minor = element_line(size = 0.1, linetype = "dashed", color = "grey")
  )

p1
# Save summary plot
# ggsave(p1, filename = "../03.out/figure5/14.Multicohorts_Multivariable_km.pdf", width = 3.8, height = 3)

```

