# Figure 6: ARG24 score provide theraputic insights

We firstly integrated the drug response data including selective drug sensitivity scores (sDSS) for 515 compounds across 132 samples (also profiled by RNA-seq). The DSS is modified area under the curve values and are calculated as shown in Yadav et al publication (Yadav B et.al, Scientific Reports 2014;4:5193). The selective drug sensitivity scores (sDSS) is healthy control normalized DSS that gives estimated cancer-selective drug responses. The higher the sDSS values indicate drug sensitivities and negative sDSS values represent drug resistance.

## TPM Normalization and Risk Score Calculation (Functional Precision Medicine Tumor Board AML)

```{r fig6_fpm_normalize_tpm, eval=TRUE, message=FALSE, warning=FALSE}
library(readxl)

count <- read.csv("../02.data/File_8_RNA_seq_Raw_Reads_163S_4Healthy.csv")
colnames(count)[1] <- "id2"
length <- read.csv("../02.data/gencode.v22_geneLength.csv")

count <- merge(length, count, by = "id2")
row.names(count) <- count$id2
count <- count[, -c(1:5)]
row.names(length) <- length$id2
length <- length[row.names(count), ]
Length <- as.vector(length$length)

TPM <- t(t(count / t(Length)) * 1e6 / colSums(count / t(Length)))
tpm <- cbind(length, TPM)[, -c(1:3)]

# Save TPM data (optional)
# write.csv(tpm, "../02.data/File_8_RNA_seq_TPM_163S_4Healthy.csv")

# Load the drug profiling data
drug <- read.csv("../02.data/File_3_Drug_response_sDSS_164S_17Healthy.csv", header = TRUE, check.names = FALSE)
samples <- intersect(colnames(tpm), colnames(drug))
tpm <- tpm[, c("Symbol", "Type", samples)]
drug <- drug[, c("Drug_name", samples)]

# Identify overlapped samples profiled by RNA-seq and Drug profiling
samples <- intersect(colnames(tpm), colnames(drug))

# Subset TPM and drug data to include only overlapped samples
tpm <- tpm[, c("Symbol", "Type", samples)]
count <- count[, samples]
drug <- drug[, c("Drug_name", samples)]

gene.min <- read.csv("../02.data/24-gene_signature.csv", row.names = 1)
mat <- subset(tpm, Symbol %in% rownames(gene.min))
row.names(mat) <- mat$Symbol
mat <- data.frame(t(log2(mat[, -c(1,2)] + 1)))
colnames(mat)[19:20] <- c("INHBA-AS1", "RP5-1119A7.14")
colnames(mat)[23:24] <- c("UNC5B-AS1", "RP11-158I9.8")

risk_score <- function(dat, candidate_genes, lasso_report) {
  genes <- intersect(colnames(dat), candidate_genes)
  weighted <- sweep(dat[, genes], 2, lasso_report[genes, 1], `*`)
  data.frame(risk_score = rowSums(weighted))
}

mat_RiskScores <- risk_score(mat, colnames(mat), gene.min)
drug <- data.frame(t(drug))
colnames(drug) <- drug[1, ]; drug <- drug[-1, ]
drug_exp <- cbind(mat_RiskScores, drug)
# write.csv(drug_exp, "../02.data/Functional_Precision_Medicine_drug_exp.csv")

### Make the matrix for DE analysis

riskscore_quantiles <- quantile(drug_exp$risk_score, probs = c(0.25, 0.75))

# Get sample names
bottom_25_samples <- rownames(drug_exp)[drug_exp$risk_score <= riskscore_quantiles[1]]
top_25_samples <- rownames(drug_exp)[drug_exp$risk_score >= riskscore_quantiles[2]]

select_samples <- c(bottom_25_samples, top_25_samples)

# Subset TPM and count data to include only selected samples

tpm <- tpm[, c("Symbol", "Type", select_samples)]
count <- count[, select_samples]
# 
# write.csv(count, "../03.out/figure6/fpmtb_top_bottom_25_count.csv")
# write.csv(tpm, "../03.out/figure6/fpmtb_top_bottom_25_tpm.csv")

## 2. Create the dot plot with cutoff lines
# First, create a rank-ordered version of the data
drug_exp_ranked <- drug_exp[order(drug_exp$risk_score), ]
drug_exp_ranked$rank <- 1:nrow(drug_exp_ranked)

# Create the plot
library(ggplot2)
library(scales)
library(RColorBrewer)

# Enhanced squish transformation with better handling of extremes
squish_trans <- function(lower, upper) {
  trans_new(
    name = "squish",
    transform = function(x) {
      ifelse(x > upper, upper + log10(x - upper + 1),
             ifelse(x < lower, lower - log10(abs(x - lower) + 1), x))
    },
    inverse = function(x) {
      ifelse(x > upper, 10^(x - upper) + upper - 1,
             ifelse(x < lower, -10^(abs(x - lower)) + lower + 1, x))
    }
  )
}

# Calculate bounds and risk groups
lower <- quantile(drug_exp$risk_score, 0.05)
upper <- quantile(drug_exp$risk_score, 0.95)

# 3. Create the transformation object
my_trans <- squish_trans(lower, upper)

drug_exp_ranked$risk_group <- cut(drug_exp_ranked$risk_score,
                                 breaks = c(-Inf, riskscore_quantiles[1], 
                                           riskscore_quantiles[2], Inf),
                                 labels = c("Low 25%", "Middle 50%", "High 25%"))

# Create the plot with enhanced aesthetics
p <- ggplot(drug_exp_ranked, aes(x = rank, y = risk_score, 
                                color = risk_group, size = risk_score)) +
  geom_point(alpha = 0.7, shape = 19) +  # Use filled circles with transparency
  
  # Improved quantile lines
  geom_hline(yintercept = riskscore_quantiles, 
             linetype = "dashed", linewidth = 0.8,
             color = c("#E41A1C", "#377EB8")) +
  
  # Color scale (ColorBrewer palette)
  scale_color_manual(values = c("#E41A1C", "#4DAF4A", "#377EB8"),
                    name = "Risk Group") +
  
  # Size scale for dots
  scale_size_continuous(range = c(1.5, 5), 
                       guide = guide_legend(title = "Risk Score")) +
  
  # Squish transformation for y-axis
  scale_y_continuous(trans = my_trans,  # Use the pre-defined transformation
                    breaks = pretty_breaks(n = 8),
                    expand = expansion(mult = 0.05)) +
  
  # Labels and theme
  labs(x = "Sample Rank", 
       y = "Risk Score (compressed extremes)",
       title = "Risk Score Distribution",
       subtitle = "Dashed lines indicate top/bottom 25% percentiles") +
  
  # Professional theme
  theme_classic(base_size = 12) +
  theme(
    legend.position = "right",
    legend.box = "vertical",
    panel.grid.major = element_line(color = "grey90", linewidth = 0.2),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey40"),
    axis.title = element_text(face = "bold")
  )

# Add annotation for quantiles
p <- p + annotate("text", x = Inf, y = riskscore_quantiles[1], 
                 label = "25th percentile", hjust = 1.1, vjust = -0.5,
                 color = "#E41A1C", size = 3.5) +
         annotate("text", x = Inf, y = riskscore_quantiles[2], 
                 label = "75th percentile", hjust = 1.1, vjust = -0.5,
                 color = "#377EB8", size = 3.5)

p
```

## Correlation Between ARG24 Score and Drug Sensitivity (FPMTB Dataset)

```{r fig6_fpm_drug_correlation_volcano, eval=TRUE, message=FALSE, warning=FALSE}
library(corrplot)
library(RColorBrewer)
library(EnhancedVolcano)

drug_exp <- read.csv("../02.data/Functional_Precision_Medicine_drug_exp.csv", row.names = 1, check.names = FALSE)

stat1 <- data.frame(ID = colnames(drug_exp)[-1])
stat1$P.value <- NA
stat1$rho <- NA
for (i in 2:ncol(drug_exp)) {
  test <- cor.test(drug_exp[, 1], drug_exp[, i])
  stat1$P.value[i - 1] <- test$p.value
  stat1$rho[i - 1] <- test$estimate
}
stat1$fdr <- p.adjust(stat1$P.value, method = "BH")
row.names(stat1) <- stat1$ID

# pdf("../03.out/figure6/01.fpm_drugs_valcano.pdf", width = 6.2, height = 8)
EnhancedVolcano(
  stat1,
  lab = rownames(stat1),
  selectLab = subset(stat1, P.value < 0.05 & abs(rho) > 0.1)$ID,
  x = 'rho',
  y = 'P.value',
  xlab = 'Correlation with drug sensitivity (sDSS)',
  pCutoff = 0.05,
  FCcutoff = 0.2,
  xlim = c(-0.4, 0.4),
  ylim = c(0, 3)
)
# dev.off()

significant <- subset(stat1, P.value < 0.05 & abs(rho) > 0.2)
significant <- significant[order(significant$rho, decreasing = TRUE), ]
corr_dat <- data.frame(rho = significant$rho)
p_dat <- data.frame(p.value = significant$P.value)
row.names(corr_dat) <- row.names(p_dat) <- row.names(significant)

# pdf("../03.out/figure6/02.fpm_drugs_colplot.pdf", width = 3, height = 5)
corrplot(
  data.matrix(corr_dat), 
  method = "circle",
  p.mat = data.matrix(p_dat),
  sig.level = 0.05,
  insig = "label_sig",
  tl.col = 'black',
  col = colorRampPalette(brewer.pal(11, 'PRGn'))(100),
  col.lim = c(-0.75, 0.75)
)
# dev.off()
```

## Representative Scatter Plots for Verdinexor and Selinexor (FPM Dataset)

```{r fig6_fpm_scatter_selinexor_verdinexor, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)
library(patchwork)

mat <- read.csv("../02.data/Functional_Precision_Medicine_drug_exp.csv", row.names = 1, check.names = FALSE)

make_scatter <- function(df, drug_name) {
  ggscatter(
    na.omit(df[, c("risk_score", drug_name)]),
    x = "risk_score", y = drug_name,
    size = "risk_score", color = drug_name,
    alpha = 0.5,
    add = "reg.line", 
    add.params = list(color = "grey", fill = "lightgray"),
    conf.int = TRUE,
    cor.coef = TRUE,
    cor.coeff.args = list(method = "pearson", label.x = 2, label.sep = "\n")
  ) +
    theme_bw() +
    scale_color_gradient(low = "#CCECE6", high = "#006D2C") +
    labs(
      title = "Functional Precision Medicine Tumor Board AML",
      x = "ARG24 Score (Chemotherapy resistance)", 
      y = paste0("Ex vivo ", drug_name, " sensitivity (sDSS)")
    )
}

p1 <- make_scatter(mat, "Verdinexor")
p2 <- make_scatter(mat, "Selinexor")
combined <- p1 / p2
combined
# ggsave(combined, filename = "../03.out/figure6/03.fpm_selino_colplot.pdf", width = 4.5, height = 7.4)
```

## Drug Correlation Analysis in BeatAML Dataset

```{r fig6_beataml_correlation_analysis, eval=TRUE, message=FALSE, warning=FALSE}
library(corrplot)
library(readxl)
library(RColorBrewer)
library(EnhancedVolcano)

t1 <- read.csv("../02.data/BeatAML_riskScore_corr_input.csv", header = TRUE, row.names = 1, check.names = FALSE)
t1 <- t1[, -c(1:3)]

stat2 <- data.frame(ID = colnames(t1)[-1])
stat2$P.value <- NA
stat2$rho <- NA
for (i in 2:ncol(t1)) {
  res <- cor.test(t1[, 1], -1 * t1[, i], method = "pearson")
  stat2$P.value[i - 1] <- res$p.value
  stat2$rho[i - 1] <- res$estimate
}

stat2$fdr <- p.adjust(stat2$P.value, method = "BH")
row.names(stat2) <- stat2$ID

# pdf("../03.out/figure6/04.beataml_drugs_valcano.pdf", width = 6.2, height = 8)
EnhancedVolcano(
  stat2,
  lab = rownames(stat2),
  x = 'rho',
  y = 'P.value',
  xlab = 'Correlation with drug sensitivity (-AUC value)',
  pCutoff = 0.05,
  FCcutoff = 0.1,
  xlim = c(-0.3, 0.3),
  ylim = c(0, 4)
)
# dev.off()

stat2 <- subset(stat2, P.value < 0.05 & abs(rho) > 0.1)
stat2 <- stat2[order(stat2$rho, decreasing = TRUE), ]
corr_dat <- data.frame(rho = stat2$rho)
p_dat <- data.frame(p.value = stat2$P.value)
row.names(corr_dat) <- row.names(p_dat) <- row.names(stat2)

# pdf("../03.out/figure6/05.beataml_drugs_colplot.pdf", width = 2, height = 3)
corrplot(
  data.matrix(corr_dat),
  method = "circle",
  tl.cex = 1,
  p.mat = data.matrix(p_dat),
  sig.level = 0.05,
  insig = "label_sig",
  pch.col = "black",
  pch.cex = 1,
  is.corr = FALSE,
  tl.col = 'black',
  cl.ratio = 0.5,
  col = colorRampPalette(brewer.pal(11, 'BrBG'))(100),
  col.lim = c(-0.3, 0.3)
)
# dev.off()
```

```{r}
library(tidyverse)
library(GSVA)
library(GSEABase)
library(dplyr)
library(tibble)

# Read data and preprocess
mat <- read.csv("../03.out/figure6/beataml_400_tpm.csv", header = T, row.names = 1)

# Step-by-step safe transformation
mat_processed <- mat %>%
  as_tibble() %>%
  mutate(
    avg_expr = rowMeans(across(-c(Symbol, Type)), na.rm = TRUE)
  ) %>%
  arrange(desc(avg_expr)) %>%
  distinct(Symbol, .keep_all = TRUE) %>%
  column_to_rownames("Symbol") %>%
  dplyr::select(-Type, -avg_expr)

# Log2 transform (add pseudocount to avoid -Inf)
mat_log <- log2(mat_processed + 1)

# Read gene sets
gene.set <- getGmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Set matrixStats options to avoid warning
options(matrixStats.useNames = TRUE) 

# Run GSVA with proper parameters
mat_log <- as.matrix(mat_log)
gsvapar <- gsvaParam(mat_log, gene.set, maxDiff=TRUE)
gsva_es <- gsva(gsvapar)
gsva_es <- data.frame(t(gsva_es))

# Save results
write.csv(gsva_es, "../03.out/figure6/gsva_nuclear_export_pathways_results_beataml.csv")
```


### Select the top 25% and bottom 25%

```{r}
# Create the plot
library(ggplot2)
library(scales)
library(RColorBrewer)

df <- read.csv("../02.data/BeatAML_riskScore_corr_input.csv", header = TRUE, row.names = 1, check.names = FALSE)

riskscore_quantiles <- quantile(df$risk_score, probs = c(0.25, 0.75))

# Get sample names
bottom_25_samples <- rownames(df)[df$risk_score <= riskscore_quantiles[1]]
top_25_samples <- rownames(df)[df$risk_score >= riskscore_quantiles[2]]

select_samples <- c(bottom_25_samples, top_25_samples)

# Subset TPM and count data to include only selected samples
count <- read.csv("../03.out/figure6/beataml_400_count.csv", header = T, row.names = 1)
tpm <- read.csv("../03.out/figure6/beataml_400_tpm.csv", header = T, row.names = 1)

tpm <- tpm[, c("ID", "Symbol", "Type", select_samples)]
count <- count[, select_samples]
# 
# write.csv(count, "../03.out/figure6/beataml_top_bottom_25_count.csv")
# write.csv(tpm, "../03.out/figure6/beataml_top_bottom_25_tpm.csv")

## 2. Create the dot plot with cutoff lines
# First, create a rank-ordered version of the data
df_ranked <- df[order(df$risk_score), ]
df_ranked$rank <- 1:nrow(df_ranked)

# Enhanced squish transformation with better handling of extremes
squish_trans <- function(lower, upper) {
  trans_new(
    name = "squish",
    transform = function(x) {
      ifelse(x > upper, upper + log10(x - upper + 1),
             ifelse(x < lower, lower - log10(abs(x - lower) + 1), x))
    },
    inverse = function(x) {
      ifelse(x > upper, 10^(x - upper) + upper - 1,
             ifelse(x < lower, -10^(abs(x - lower)) + lower + 1, x))
    }
  )
}

# Calculate bounds and risk groups
lower <- quantile(df$risk_score, 0.05)
upper <- quantile(df$risk_score, 0.95)

# 3. Create the transformation object
my_trans <- squish_trans(lower, upper)

df_ranked$risk_group <- cut(df_ranked$risk_score,
                     breaks = c(-Inf, riskscore_quantiles[1], 
                                riskscore_quantiles[2], Inf),
                     labels = c("Low 25%", "Middle 50%", "High 25%"))

# Create the plot with enhanced aesthetics
p <- ggplot(df_ranked, aes(x = rank, y = risk_score, 
                                color = risk_group, size = risk_score)) +
  geom_point(alpha = 0.7, shape = 19) +  # Use filled circles with transparency
  
  # Improved quantile lines
  geom_hline(yintercept = riskscore_quantiles, 
             linetype = "dashed", linewidth = 0.8,
             color = c("#E41A1C", "#377EB8")) +
  
  # Color scale (ColorBrewer palette)
  scale_color_manual(values = c("#E41A1C", "#4DAF4A", "#377EB8"),
                    name = "Risk Group") +
  
  # Size scale for dots
  scale_size_continuous(range = c(1.5, 5), 
                       guide = guide_legend(title = "Risk Score")) +
  
  # Squish transformation for y-axis
  scale_y_continuous(trans = my_trans,  # Use the pre-defined transformation
                    breaks = pretty_breaks(n = 8),
                    expand = expansion(mult = 0.05)) +
  
  # Labels and theme
  labs(x = "Sample Rank", 
       y = "Risk Score (compressed extremes)",
       title = "Risk Score Distribution",
       subtitle = "Dashed lines indicate top/bottom 25% percentiles") +
  
  # Professional theme
  theme_classic(base_size = 12) +
  theme(
    legend.position = "right",
    legend.box = "vertical",
    panel.grid.major = element_line(color = "grey90", linewidth = 0.2),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey40"),
    axis.title = element_text(face = "bold")
  )

# Add annotation for quantiles
p <- p + annotate("text", x = Inf, y = riskscore_quantiles[1], 
                 label = "25th percentile", hjust = 1.1, vjust = -0.5,
                 color = "#E41A1C", size = 3.5) +
         annotate("text", x = Inf, y = riskscore_quantiles[2], 
                 label = "75th percentile", hjust = 1.1, vjust = -0.5,
                 color = "#377EB8", size = 3.5)

p
```


## Selinexor Sensitivity vs. ARG24 Score in BeatAML Dataset

```{r fig6_beataml_selinexor_scatter, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)

mat <- read.csv("../02.data/BeatAML_riskScore_corr_input.csv", header = TRUE, row.names = 1, check.names = FALSE)
mat$neg_Selinexor <- -1 * mat$Selinexor

p3 <- ggscatter(
  na.omit(mat[, c("risk_score", "neg_Selinexor", "Selinexor")]),
  x = "risk_score",
  y = "neg_Selinexor",
  size = "risk_score",
  color = "neg_Selinexor",
  alpha = 0.5,
  title = "BeatAML",
  add = "reg.line",
  add.params = list(color = "grey", fill = "lightgray"),
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(method = "pearson", label.x = 2, label.sep = "\n")
) +
  theme_bw() +
  scale_color_gradient(low = "#C6DBEF", high = "#01665E") +
  labs(
    x = "ARG24 Score (Chemotherapy resistance)",
    y = "Ex vivo Selinexor sensitivity (-AUC)"
  )

p3
# ggsave(p3, filename = "../03.out/figure6/06.beataml_selino_colplot.pdf", width = 4.5, height = 3.3)
```

## Correlation Between ARG24 Score and XPO1 Expression in each cohort

```{r fig6_xpo1_expression_correlation, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)
library(patchwork)

dat <- read.csv("../02.data/XPO1_ARG24_cor_mat.csv", header = TRUE, row.names = 1, check.names = FALSE)
dat$XPO1 <- log2(dat$XPO1 + 1)

p4 <- ggscatter(
  dat,
  x = "RiskScore",
  y = "XPO1",
  color = "Cohort",
  alpha = 0.5,
  add = "reg.line",
  add.params = list(color = "grey", fill = "lightgray"),
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(method = "pearson", label.x = 2, label.sep = "\n")
) +
  theme_bw() +
  facet_wrap(. ~ Cohort) +
  labs(
    x = "ARG24 Score (Chemotherapy resistance)",
    y = "XPO1 expression"
  )

p4

mat <- read.csv("../02.data/XPO1_ARG24_cor_CCLE_AML_cell-lines.csv", header = TRUE)
mat$XPO1 <- log2(mat$XPO1)

p5 <- ggscatter(
  mat,
  x = "risk_score",
  y = "XPO1",
  color = "XPO1",
  alpha = 0.5,
  add = "reg.line",
  add.params = list(color = "grey", fill = "lightgray"),
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(method = "pearson", label.x = 2, label.sep = "\n")
) +
  theme_bw() +
  scale_color_gradient(low = "#FEE8C8", high = "#D7301F") +
  labs(
    x = "ARG24 Score (Chemotherapy resistance)",
    y = "XPO1 expression"
  )

p5

# plots <- p4 | p5
# plots
# ggsave(plots, filename = "../03.out/figure6/07.arg24_xpo1.pdf", width = 8, height = 3.3)
```

## Detailed Visualization: ARG24 Score vs. XPO1 Expression (Scatter, Density, and Group Comparison)

```{r eval=TRUE, message = FALSE, warning = FALSE}
library(ggpubr)
library(ggplot2)
library(ggExtra)
library(patchwork)
library(ggrastr)
library(tidyverse)
library(ggstatsplot)
library(RColorBrewer)

# Load and preprocess the dataset
dat <- read.csv("../02.data/XPO1_ARG24_cor_mat.csv", header = TRUE, row.names = 1, check.names = FALSE)
dat$XPO1 <- log2(dat$XPO1 + 1)
dat$ARG24_group <- ifelse(dat$RiskScore >= median(dat$RiskScore), "High", "Low")
dat$ARG24_group <- factor(dat$ARG24_group, levels = c("Low", "High"))
col <- brewer.pal(2, "Set2")

## Panel A: Scatter plot with density margin
p1 <- ggplot(dat, aes(x = RiskScore, y = XPO1, color = ARG24_group)) +
  geom_point(aes(size = RiskScore, shape = Cohort), alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = col[3]) +
  scale_color_manual(values = col) +
  labs(x = "ARG24 Score", y = "XPO1 expression") +
  theme_bw()

# Pearson correlation test
stat <- cor.test(dat$RiskScore, dat$XPO1, method = "pearson")
lab <- paste0("R = ", round(stat$estimate, 3), "\nP = ", signif(stat$p.value, 3))

# Annotate correlation results on the plot
p2 <- p1 + annotate("text", x = 1.2, y = 0.55, label = lab)

# Add marginal density plots
p3 <- ggMarginal(p2, type = "density", groupColour = TRUE, groupFill = TRUE)

# Display Panel A
p3
# ggsave(p3, filename = "../03.out/figure6/08.arg24_xpo1_allcohorts.pdf", width = 5.4, height = 4.2)


## Panel B: Group-wise comparison using violin + boxplot
p_violin <- ggplot(dat, aes(x = ARG24_group, y = XPO1, color = ARG24_group)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.4, alpha = 0.75, outlier.shape = NA) +
  geom_quasirandom_rast(width = 0.25, alpha = 0.5) +
  scale_color_manual(values = c("Low" = col[1], "High" = col[2])) +
  labs(x = NULL, y = "XPO1 expression (log2-transformed)") +
  theme_pubr(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1),
    text = element_text(size = 12)
  ) +
  stat_compare_means(comparisons = list(c("Low", "High")), method = 'wilcox.test')

# Display Panel B
p_violin
# ggsave(p_violin, filename = "../03.out/figure6/09.xpo1_arg24_group.pdf", width = 2, height = 3.6)
```

## Selinexor Drug Response Curves

```{python}
# import pandas as pd
# import numpy as np
# import matplotlib.pyplot as plt
# import seaborn as sns
# from scipy.optimize import curve_fit
# from pathlib import Path
# 
# # Load the Excel file
# file_path = Path("../data/selinexor_IC50_2024.10.15_longFormat.xlsx")
# df = pd.read_excel(file_path)
# 
# # Reshape the data to long format
# df_long = df.melt(id_vars=["Selinexor", "Cell_line"], 
#                   value_vars=["Rep1", "Rep2", "Rep3"],
#                   var_name="Replicate", value_name="Viability")
# 
# # Define 4-parameter logistic function
# def four_param_logistic(x, a, b, c, d):
#     return d + (a - d) / (1 + (x / c)**b)
# 
# # Fit and plot curves
# cell_lines = df_long["Cell_line"].unique()
# ic50_results = []
# 
# # Initialize figure
# fig, ax = plt.subplots(figsize=(10, 6))
# sns.set(style="whitegrid")
# 
# for cell_line in cell_lines:
#     data = df_long[df_long["Cell_line"] == cell_line]
#     xdata = data["Selinexor"].values
#     ydata = data["Viability"].values
# 
#     try:
#         # Fit curve
#         popt, _ = curve_fit(four_param_logistic, xdata, ydata, 
#                             bounds=([0, 0, 0, 0], [200, 10, 100, 200]), 
#                             p0=[100, 1, 1, 0])
#         a, b, c, d = popt
#         ic50_results.append({"Cell_line": cell_line, "IC50": c})
#         
#         # Generate fitted curve
#         xfit = np.logspace(np.log10(min(xdata[xdata > 0])), np.log10(max(xdata)), 100)
#         yfit = four_param_logistic(xfit, *popt)
#         ax.plot(xfit, yfit, label=f"{cell_line} (IC50={c:.3f})")
#         
#         # Raw data
#         sns.scatterplot(x="Selinexor", y="Viability", data=data, ax=ax, label=f"{cell_line} raw", alpha=0.6)
#     except RuntimeError:
#         ic50_results.append({"Cell_line": cell_line, "IC50": np.nan})
#         print(f"Could not fit curve for {cell_line}")
# 
# # Finalize plot
# ax.set_xscale("log")
# ax.set_xlabel("Selinexor (µM)")
# ax.set_ylabel("Cell Viability (%)")
# ax.set_title("Drug Response Curves for Selinexor")
# ax.legend(bbox_to_anchor=(1.05, 1), loc='upper left')
# plt.tight_layout()
# 
# # Save to PDF
# pdf_path = "../out/figure6/Selinexor_Drug_Response_Curves.pdf"
# fig.savefig(pdf_path)

```

## ARG24 Score vs. Selinexor IC50 in AML Cell Lines

```{r fig6_arg24_selinexor_ic50, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)
library(reshape2)
library(patchwork)
library(ggplot2) 

# Load risk score and IC50 data
mat <- read.csv("../02.data/risk_score_XPO1_IC50.csv", header = TRUE)

# Scatter plot: ARG24 score vs. Selinexor IC50
p1 <- ggscatter(
  data = mat,
  x = "risk_score",
  y = "Selinexor_IC50",
  color = "Selinexor_IC50",
  size = 3,
  alpha = 0.5,
  add = "reg.line",
  add.params = list(color = "#666666", fill = "lightgray"),
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(method = "pearson", label.sep = "\n")
) +
  theme_bw() +
  labs(
    x = "ARG24 Score",
    y = "IC50 (Selinexor)"
  ) +
  scale_color_gradient(low = "#E7D4E8", high = "#762A83") +
  geom_text(aes(label = Cell_Line), hjust = 0.5, vjust = -1, size = 3, color = "black")

p1

# Save plot to file (uncomment to activate)
# ggsave(p1, filename = "../03.out/figure6/10.risk_score_AML_cells_IC50.pdf", width = 5.1, height = 3.8)
```

## ARG24 Score vs. Selinexor IC50 in Patients

```{r fig6_patient_qpcr_heatmap, eval=TRUE, message=FALSE, warning=FALSE}
library(pheatmap)

# Step 1: Load gene expression (qPCR) and gene coefficient data
dat <- read.csv("../02.data/patient_selinexor_pcr_gapdh.csv", header = TRUE, row.names = 1, check.names = FALSE)
row.names(dat) <- paste0("P", 1:8)
gene.min <- read.csv("../02.data/24-gene_signature.csv", header = TRUE, row.names = 1, check.names = FALSE)

# Step 2: Ensure gene order and subset
genes <- rownames(gene.min)
dat <- dat[, genes]

# Step 3: Calculate ARG24 score
calculate_risk_scores <- function(features, coefficients) {
  scores <- sweep(features, 2, coefficients, `*`)
  data.frame(ARG24 = rowSums(scores))
}

rjaml_pcr <- calculate_risk_scores(dat, gene.min[, "Coef"])

# Step 4: Reorder data by ARG24 score
dat_ordered <- dat[order(rjaml_pcr$ARG24, decreasing = TRUE), ]

# Step 5: IC50 values and alignment
ic50 <- data.frame(IC50 = c(0.7288, 1.440, 4.112, 3.365, 1.443, 1.643, 1.120, 1.938))
row.names(ic50) <- row.names(dat)
ic50_ordered <- ic50[rownames(dat_ordered), ]

# Step 6: Annotation for heatmap
annotation_row <- data.frame(
  ARG24 = rjaml_pcr$ARG24,
  LogIC50 = ic50$IC50
)
row.names(annotation_row) <- row.names(dat)

# Step 7: Plot heatmap
# pdf("../03.out/figure6/11.qpcr_heatmap.pdf", width = 6, height = 3)
pheatmap(
  dat_ordered,
  scale = "row",
  cluster_rows = FALSE,
  cluster_cols = TRUE,
  annotation_row = annotation_row,
  fontsize_row = 10,
  fontsize_col = 8,
  show_rownames = TRUE,
  show_colnames = TRUE,
  breaks = seq(-2, 2, length.out = 100),
  color = colorRampPalette(c("#7c9d97", "#F0F0F0", "#e9b383"))(100)
)
# dev.off()
```

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(ggpubr)

# Step 8: Scatter plot of ARG24 score vs. Selinexor IC50
mat <- data.frame(
  Cell_Line = row.names(rjaml_pcr),
  risk_score = rjaml_pcr$ARG24,
  Selinexor_IC50 = ic50$IC50
)

p1 <- ggscatter(
  data = mat,
  x = "risk_score",
  y = "Selinexor_IC50",
  color = "Selinexor_IC50",
  alpha = 0.5,
  add = "reg.line",
  add.params = list(color = "#666666", fill = "lightgray"),
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(method = "pearson", label.sep = "\n")
) +
  geom_text(aes(label = Cell_Line), hjust = 0.5, vjust = -1, size = 3, color = "black") +
  theme_bw() +
  labs(x = "ARG24 Score", y = "IC50 (Selinexor)") +
  scale_color_gradient(low = "#FEE0D2", high = "#67000D")

p1

# Save to file
# ggsave(p1, filename = "../03.out/figure6/12.risk_score_AML_selinexor_IC50.pdf", width = 5.1, height = 3.8)
```

## ssGSEA

```{r}
library(tidyverse)
library(GSVA)
library(GSEABase)
library(dplyr)
library(tibble)

# Read data and preprocess
mat <- read.csv("../03.out/figure6/fpmtb_132_tpm.csv", header = TRUE, row.names = 1, check.names = FALSE)

# Step-by-step safe transformation
mat_processed <- mat %>%
  rownames_to_column("GeneID") %>%
  as_tibble() %>%
  mutate(
    avg_expr = rowMeans(across(-c(GeneID, Symbol, Type)), na.rm = TRUE)
  ) %>%
  arrange(desc(avg_expr)) %>%
  distinct(Symbol, .keep_all = TRUE) %>%
  column_to_rownames("Symbol") %>%
  dplyr::select(-GeneID, -Type, -avg_expr)

# Log2 transform (add pseudocount to avoid -Inf)
mat_log <- log2(mat_processed + 1)

# Read gene sets
gene.set <- getGmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Set matrixStats options to avoid warning
options(matrixStats.useNames = TRUE) 

# Run GSVA with proper parameters
mat_log <- as.matrix(mat_log)
gsvapar <- gsvaParam(mat_log, gene.set, maxDiff=TRUE)
gsva_es <- gsva(gsvapar)
gsva_es <- data.frame(t(gsva_es))

gsva_es <- gsva_es[order(gsva_es$Integrated), ]
top_33 <- rownames(head(gsva_es, 33))
bot_33 <- rownames(tail(gsva_es, 33))
samples <- c(top_33, bot_33)

tpm <- read.csv("../03.out/figure6/fpmtb_132_tpm.csv", header = TRUE, row.names = 1, check.names = FALSE)
count <- read.csv("../03.out/figure6/fpmtb_132_count.csv", header = TRUE, row.names = 1, check.names = FALSE)

tpm <- tpm %>% rownames_to_column("ID")
tpm <- tpm[, c("ID", "Symbol", "Type", samples)]
count <- count[, samples]

write.csv(tpm, "../03.out/figure6/fpmtb_top_bottom_25_tpm.csv")
write.csv(count, "../03.out/figure6/fpmtb_top_bottom_25_count.csv")
# Save results
write.csv(gsva_es, "../03.out/figure6/gsva_nuclear_export_pathways_results.csv")
```

## DE analysis in FPMTB cohot

```{r}
library(DESeq2)
library(tidyverse)
## run DESEQ2
source("../01.script/RunDESeq2.R")

tpm <- read.csv("../03.out/figure6/fpmtb_top_bottom_25_tpm.csv", header=T, row.names = 1)

tpm <- subset(tpm, Type %in% c("protein_coding", "miRNA", "lincRNA", "antisense",
                               "processed_transcript",
                               "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene",
                               "transcribed_unprocessed_pseudogene","macro_lncRNA", "TEC",
                               "3prime_overlapping_ncRNA","sense_intronic", "sense_overlapping",
                               "unprocessed_pseudogene"))

# filter low-express genes
tpm <- tpm[rowMeans(tpm[, 4:ncol(tpm)]) >= 0.25, ]

count <- read.csv("../03.out/figure6/fpmtb_top_bottom_25_count.csv", header=T, row.names = 1)
count <- count[tpm$ID, ]

RunDESeq2(count_mat = count, n.cont = 33, n.treat = 33,
          prefix = "../03.out/figure6/DESeq2_arg24_high_low_fpmtb", sort.p = FALSE,
          merge.normalized = TRUE,
          normalized_mat = tpm)

```

## ARG24-high vs. ARG24-low (FPMTB)

```{r}
library(clusterProfiler)
library(fgsea)
library(enrichplot)
library(dplyr)
library(tidyverse)

degene_mat <- read.table("../03.out/figure6/DESeq2_arg24_high_low_fpmtb_with_normalized_mat.txt", header = TRUE)

# Prepare ranked gene list for GSEA
DE_list <- degene_mat %>%
  dplyr::select(ID = 9, Corr = 3) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na() %>%
  mutate(rank = rank(Corr, ties.method = "random"))

geneList <- DE_list$Corr
names(geneList) <- DE_list$ID
geneList <- sort(geneList, decreasing = TRUE)

# Load gene sets
genesets <- read.gmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Convert TERM2GENE to a list
gene_sets_list <- split(genesets$gene, genesets$term)

# Run standard fgsea 
fgsea.res <- fgsea(
  pathways = gene_sets_list,
  stats = geneList,
  minSize = 1,
  maxSize = 1200,
  nperm = 100000
)

fgsea.res

pdf("../03.out/figure6/GOBP_NUCLEAR_EXPORT_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOBP_NUCLEAR_EXPORT"]], geneList) +
  ggtitle("Enrichment of GOBP_NUCLEAR_EXPORT")
dev.off()

pdf("../03.out/figure6/GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT"]], geneList) +
  ggtitle("Enrichment of GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT")
dev.off()

pdf("../03.out/figure6/GOBP_MRNA_TRANSPORT_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOBP_MRNA_TRANSPORT"]], geneList) +
  ggtitle("Enrichment of GOBP_MRNA_TRANSPORT")
dev.off()

```

## Valcanoplot (FPMTB)

```{r}
# Load necessary libraries
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(ggrastr)

# Load data
diff <- read.table("../03.out/figure6/DESeq2_arg24_high_low_fpmtb_with_normalized_mat.txt", header = TRUE)

# Cap log2FoldChange to [-4, 4] and padj to [1e-6, 1]
diff <- diff %>%
  mutate(
    log2FoldChange = case_when(
      log2FoldChange > 4 ~ 4,
      log2FoldChange < -4 ~ -4,
      TRUE ~ log2FoldChange
    ),
    padj = case_when(
      padj < 1e-6 ~ 1e-6,
      TRUE ~ padj
    )
  )

# Classify genes based on log2FoldChange and adjusted p-value
diff <- diff %>%
  mutate(g = case_when(
    log2FoldChange > 0.5 & padj < 0.05 ~ "up",
    log2FoldChange < -0.5 & padj < 0.05 ~ "down",
    TRUE ~ "stable"
  ))

# Extract significantly upregulated and downregulated genes
diff_up <- diff %>% filter(g == "up")
diff_down <- diff %>% filter(g == "down")

# Load gene sets
genesets <- read.gmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Extract GOBP_NUCLEAR_EXPORT and GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT gene symbols
GOBP_NUCLEAR_EXPORT <- genesets %>% filter(term == "GOBP_NUCLEAR_EXPORT") %>% pull(gene)
GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT <- genesets %>% filter(term == "GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT") %>% pull(gene)

# Classify genes in the differential expression dataset
diff <- diff %>%
  mutate(gene_col = case_when(
    Symbol %in% GOBP_NUCLEAR_EXPORT ~ "GOBP_NUCLEAR_EXPORT",
    Symbol %in% GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT ~ "GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT",
    TRUE ~ "all"
  ))

# Prepare data subsets
data_bg <- diff %>% filter(gene_col == "all")
data_geneset <- diff %>% filter(grepl("GOBP_", gene_col))
data_GOBP_NUCLEAR_EXPORT <- diff %>% filter(gene_col == "GOBP_NUCLEAR_EXPORT")
data_GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT <- diff %>% filter(gene_col == "GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT")

# Define the base plot
p <- ggplot(diff, aes(x = log2FoldChange, y = -log10(padj))) +
  # Add shaded regions for significant up- and down-regulated genes
  annotate(
    geom = "rect", fill = "#ea838e", alpha = 0.3,
    xmin = min(diff_up$log2FoldChange), xmax = max(diff_up$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  annotate(
    geom = "rect", fill = "#827fd6", alpha = 0.3,
    xmin = min(diff_down$log2FoldChange), xmax = max(diff_down$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  # Plot background points
  geom_point_rast(
    data = data_bg, shape = 21, color = "black", alpha = 0.2,
    size = 1.3, stroke = 0.6
  ) +
  # Plot gene set and key points
  geom_point_rast(
    data = data_geneset,
    aes(fill = gene_col), shape = 21, color = "black",
    size = 2, stroke = 0.8
  ) +
  # Custom fill colors for gene sets and key genes
  scale_fill_manual(
    values = c(
      "GOBP_NUCLEAR_EXPORT" = "#C98B88",
      "GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT" = "#6EB9C3"
    )
  ) +
  # Add reference lines
  geom_vline(xintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(
    yintercept = -log10(0.05), color = "#ff7d82",
    linewidth = 0.6, linetype = "dotted"
  ) +
  # Customize x-axis
  scale_x_continuous(limits = c(-4, 4), breaks = seq(-4, 4, 1)) +
  # Customize y-axis
  scale_y_continuous(limits = c(-2, 7), breaks = seq(0, 6.5, 2), expand = c(0, 0)) +
  # Apply classic theme and customize
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  ) +
  # Customize axis labels
  xlab("ARG24-High vs ARG24-Low") +
  ylab("Significance (-log10 adj.p-val.)")

# Add lineranges and annotations to the plot (p1)
p1 <- p +
  # Add lineranges for GOBP_NUCLEAR_EXPORT genes
  geom_linerange(
    data = data_GOBP_NUCLEAR_EXPORT,
    aes(x = log2FoldChange, ymin = -1, ymax = -0.35),
    color = "#C98B88", size = 0.5, linewidth = 0.1
  ) +
  # Add lineranges for GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT genes
  geom_linerange(
    data = data_GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT,
    aes(x = log2FoldChange, ymin = -1.75, ymax = -1.1),
    color = "#6EB9C3", size = 0.5, linewidth = 0.1
  ) +
  # Add text annotations for GOBP_NUCLEAR_EXPORT and GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT
  annotate(
    "text", x = -3, y = -0.7, label = "GOBP_NUCLEAR_EXPORT",
    color = "#C98B88", size = 4.8, hjust = 0
  ) +
  annotate(
    "text", x = -3, y = -1.4, label = "GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT",
    color = "#6EB9C3", size = 4.8, hjust = 0
  ) +
  # Add enrichment statistics for GOBP_NUCLEAR_EXPORT
  annotate(
    "text", x = 3.0, y = -0.7, label = "ES: 0.330\np.value = 0.008",
    color = "#C98B88", size = 4.5, lineheight = 0.8, hjust = 1
  ) +
  # Add enrichment statistics for GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT
  annotate(
    "text", x = 3.0, y = -1.4, label = "ES: 0.296\np.value = 0.022",
    color = "#6EB9C3", size = 4.5, lineheight = 0.8, hjust = 1
  )

p1
ggsave(p1, filename = "../03.out/figure6/valcano_arg24_high_low_fpmtb.pdf", width = 6.5, height = 6.6)
```

```{r}
# Subset TPM and count data to include only selected samples
sample <- read.csv("../03.out/figure6/beataml_selected.csv", header = T)
count <- read.csv("../03.out/figure6/beataml_400_count.csv", header = T, row.names = 1)
tpm <- read.csv("../03.out/figure6/beataml_400_tpm.csv", header = T, row.names = 1)

select_samples <- sample$X

tpm <- tpm[, c("ID", "Symbol", "Type", select_samples)]
count <- count[, select_samples]
# 
write.csv(count, "../03.out/figure6/beataml_top_bottom_25_count.csv")
write.csv(tpm, "../03.out/figure6/beataml_top_bottom_25_tpm.csv")

```

## DE analysis in BeatAML cohot

```{r}
library(DESeq2)
library(tidyverse)
## run DESEQ2
source("../01.script/RunDESeq2.R")

tpm <- read.csv("../03.out/figure6/beataml_top_bottom_25_tpm.csv", header=T, row.names = 1)

tpm <- subset(tpm, Type %in% c("protein_coding", "miRNA", "lincRNA", "antisense",
                               "processed_transcript",
                               "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene",
                               "transcribed_unprocessed_pseudogene","macro_lncRNA", "TEC",
                               "3prime_overlapping_ncRNA","sense_intronic", "sense_overlapping",
                               "unprocessed_pseudogene"))

# filter low-express genes
tpm <- tpm[rowMeans(tpm[, 4:ncol(tpm)]) >= 0.25, ]

count <- read.csv("../03.out/figure6/beataml_top_bottom_25_count.csv", header=T, row.names = 1)
count <- count[tpm$ID, ]

RunDESeq2(count_mat = count, n.cont = 100, n.treat = 100,
          prefix = "../03.out/figure6/DESeq2_arg24_high_low_beataml", sort.p = FALSE,
          merge.normalized = TRUE,
          normalized_mat = tpm)

```

## ARG24-high vs. ARG24-low (BeatAML)

```{r}
library(clusterProfiler)
library(fgsea)
library(enrichplot)
library(dplyr)

degene_mat <- read.table("../03.out/figure6/DESeq2_arg24_high_low_beataml_with_normalized_mat.txt", header = TRUE)

# Prepare ranked gene list for GSEA
DE_list <- degene_mat %>%
  dplyr::select(ID = 9, Corr = 3) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na() %>%
  mutate(rank = rank(Corr, ties.method = "random"))

geneList <- DE_list$Corr
names(geneList) <- DE_list$ID
geneList <- sort(geneList, decreasing = TRUE)

# Load gene sets
genesets <- read.gmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Convert TERM2GENE to a list
gene_sets_list <- split(genesets$gene, genesets$term)

# Run standard fgsea 
fgsea.res <- fgsea(
  pathways = gene_sets_list,
  stats = geneList,
  minSize = 1,
  maxSize = 1200,
  nperm = 100000
)

fgsea.res

pdf("../03.out/figure6/beataml_GOBP_NUCLEAR_EXPORT_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOBP_NUCLEAR_EXPORT"]], geneList) +
  ggtitle("Enrichment of GOBP_NUCLEAR_EXPORT")
dev.off()

pdf("../03.out/figure6/beataml_GOMF_NUCLEAR_EXPORT_SIGNAL_RECEPTOR_ACTIVITY_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOMF_NUCLEAR_EXPORT_SIGNAL_RECEPTOR_ACTIVITY"]], geneList) +
  ggtitle("Enrichment of GOMF_NUCLEAR_EXPORT_SIGNAL_RECEPTOR_ACTIVITY")
dev.off()

pdf("../03.out/figure6/beataml_GOBP_MRNA_TRANSPORT_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOBP_MRNA_TRANSPORT"]], geneList) +
  ggtitle("Enrichment of GOBP_MRNA_TRANSPORT")
dev.off()

pdf("../03.out/figure6/beataml_GOBP_NUCLEAR_RNA_SURVEILLANCE_gsea.pdf", width = 5, height = 2.5)
plotEnrichment(gene_sets_list[["GOBP_NUCLEAR_RNA_SURVEILLANCE"]], geneList) +
  ggtitle("Enrichment of GOBP_NUCLEAR_RNA_SURVEILLANCE")
dev.off()

```

## Valcanoplot (BeatAML)

```{r}
# Load necessary libraries
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(ggrastr)

# Load data
diff <- read.table("../03.out/figure6/DESeq2_arg24_high_low_beataml_with_normalized_mat.txt", header = TRUE)

# Cap log2FoldChange to [-3, 3] and padj to [1e-12, 1]
diff <- diff %>%
  mutate(
    log2FoldChange = case_when(
      log2FoldChange > 3 ~ 3,
      log2FoldChange < -3 ~ -3,
      TRUE ~ log2FoldChange
    ),
    padj = case_when(
      padj < 1e-12 ~ 1e-12,
      TRUE ~ padj
    )
  )

# Classify genes based on log2FoldChange and adjusted p-value
diff <- diff %>%
  mutate(g = case_when(
    log2FoldChange > 0.5 & padj < 0.05 ~ "up",
    log2FoldChange < -0.5 & padj < 0.05 ~ "down",
    TRUE ~ "stable"
  ))

# Extract significantly upregulated and downregulated genes
diff_up <- diff %>% filter(g == "up")
diff_down <- diff %>% filter(g == "down")

# Load gene sets
genesets <- read.gmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Extract GOBP_NUCLEAR_EXPORT and GOBP_MRNA_TRANSPORT gene symbols
GOBP_NUCLEAR_EXPORT <- genesets %>% filter(term == "GOBP_NUCLEAR_EXPORT") %>% pull(gene)
GOBP_MRNA_TRANSPORT <- genesets %>% filter(term == "GOBP_MRNA_TRANSPORT") %>% pull(gene)

# Classify genes in the differential expression dataset
diff <- diff %>%
  mutate(gene_col = case_when(
    Symbol %in% GOBP_NUCLEAR_EXPORT ~ "GOBP_NUCLEAR_EXPORT",
    Symbol %in% GOBP_MRNA_TRANSPORT ~ "GOBP_MRNA_TRANSPORT",
    TRUE ~ "all"
  ))

# Prepare data subsets
data_bg <- diff %>% filter(gene_col == "all")
data_geneset <- diff %>% filter(grepl("GOBP_", gene_col))
data_GOBP_NUCLEAR_EXPORT <- diff %>% filter(gene_col == "GOBP_NUCLEAR_EXPORT")
data_GOBP_MRNA_TRANSPORT <- diff %>% filter(gene_col == "GOBP_MRNA_TRANSPORT")

# Define the base plot
p <- ggplot(diff, aes(x = log2FoldChange, y = -log10(padj))) +
  # Add shaded regions for significant up- and down-regulated genes
  annotate(
    geom = "rect", fill = "#ea838e", alpha = 0.3,
    xmin = min(diff_up$log2FoldChange), xmax = max(diff_up$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  annotate(
    geom = "rect", fill = "#827fd6", alpha = 0.3,
    xmin = min(diff_down$log2FoldChange), xmax = max(diff_down$log2FoldChange),
    ymin = -log10(0.05), ymax = max(-log10(diff$padj))
  ) +
  # Plot background points
  geom_point_rast(
    data = data_bg, shape = 21, color = "black", alpha = 0.2,
    size = 1.3, stroke = 0.6
  ) +
  # Plot gene set and key points
  geom_point_rast(
    data = data_geneset,
    aes(fill = gene_col), shape = 21, color = "black",
    size = 2, stroke = 0.8
  ) +
  # Custom fill colors for gene sets and key genes
  scale_fill_manual(
    values = c(
      "GOBP_NUCLEAR_EXPORT" = "#C98B88",
      "GOBP_MRNA_TRANSPORT" = "#6EB9C3"
    )
  ) +
  # Add reference lines
  geom_vline(xintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "#b2b2b2", linewidth = 0.6) +
  geom_hline(
    yintercept = -log10(0.05), color = "#ff7d82",
    linewidth = 0.6, linetype = "dotted"
  ) +
  # Customize x-axis
  scale_x_continuous(limits = c(-3, 3), breaks = seq(-3, 3, 1)) +
  # Customize y-axis
  scale_y_continuous(limits = c(-2, 14), breaks = seq(0, 14, 2), expand = c(0, 0)) +
  # Apply classic theme and customize
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  ) +
  # Customize axis labels
  xlab("ARG24-High vs ARG24-Low") +
  ylab("Significance (-log10 adj.p-val.)")

# Add lineranges and annotations to the plot (p1)
p1 <- p +
  # Add lineranges for GOBP_NUCLEAR_EXPORT genes
  geom_linerange(
    data = data_GOBP_NUCLEAR_EXPORT,
    aes(x = log2FoldChange, ymin = -1, ymax = -0.35),
    color = "#C98B88", size = 0.5, linewidth = 0.1
  ) +
  # Add lineranges for GOBP_MRNA_TRANSPORT genes
  geom_linerange(
    data = data_GOBP_MRNA_TRANSPORT,
    aes(x = log2FoldChange, ymin = -1.75, ymax = -1.1),
    color = "#6EB9C3", size = 0.5, linewidth = 0.1
  ) +
  # Add text annotations for GOBP_NUCLEAR_EXPORT and GOBP_MRNA_TRANSPORT
  annotate(
    "text", x = -3, y = -0.7, label = "GOBP_NUCLEAR_EXPORT",
    color = "#C98B88", size = 4.8, hjust = 0
  ) +
  annotate(
    "text", x = -3, y = -1.4, label = "GOBP_MRNA_TRANSPORT",
    color = "#6EB9C3", size = 4.8, hjust = 0
  ) +
  # Add enrichment statistics for GOBP_NUCLEAR_EXPORT
  annotate(
    "text", x = 3.0, y = -0.7, label = "ES: 0.261\np.value = 0.036",
    color = "#C98B88", size = 4.5, lineheight = 0.8, hjust = 1
  ) +
  # Add enrichment statistics for GOBP_MRNA_TRANSPORT
  annotate(
    "text", x = 3.0, y = -1.4, label = "ES: 0.289\np.value = 0.021",
    color = "#6EB9C3", size = 4.5, lineheight = 0.8, hjust = 1
  )

p1
ggsave(p1, filename = "../03.out/figure6/valcano_arg24_high_low_beataml.pdf", width = 6.5, height = 6.6)
```

## Integration of DE out
Differential expression analysis data were integrated by performing Fisher’s method through the metaseqR Bioconductor package

```{r}
library(tidyverse)
library(metaseqR2)

# Load data
diff_fpmtb <- read.table("../03.out/figure6/DESeq2_arg24_high_low_fpmtb_with_normalized_mat.txt", 
                         header = TRUE, stringsAsFactors = FALSE)
diff_beataml <- read.table("../03.out/figure6/DESeq2_arg24_high_low_beataml_with_normalized_mat.txt", 
                           header = TRUE, stringsAsFactors = FALSE)

# Get common gene symbols
common_genes <- intersect(diff_fpmtb$Symbol, diff_beataml$Symbol)

# Clean and subset datasets
prepare_diff_data <- function(df, dataset_label) {
  df %>%
    filter(Symbol %in% common_genes) %>%
    distinct(Symbol, .keep_all = TRUE) %>%
    select(Symbol, log2FoldChange, padj) %>%
    rename_with(~ paste0(.x, "_", dataset_label), -Symbol)
}

diff_fpmtb_clean <- prepare_diff_data(diff_fpmtb, "fpmtb")
diff_beataml_clean <- prepare_diff_data(diff_beataml, "beataml")

# Merge datasets by Symbol and set row names
merged_df <- inner_join(diff_fpmtb_clean, diff_beataml_clean, by = "Symbol") %>%
  column_to_rownames("Symbol")

# Now combine the adj.p-values using the Simes method
pSimes <- apply(merged_df[, c("padj_fpmtb", "padj_beataml")], 1, combineSimes)
merged_df$comb_padj <- pSimes
merged_df$Symbol <- row.names(merged_df)

```

```{r}
library(clusterProfiler)
library(tidyverse)
library(ggplot2)
library(ggrepel)

# Load gene sets
genesets <- read.gmt("../03.out/figure6/NUCLEAR_EXPORT_Pathways.v2024.1.Hs.gmt")

# Select specific pathways and extract unique gene symbols
selected_terms <- c("GOBP_NUCLEAR_EXPORT", 
                    "GOBP_REGULATION_OF_NUCLEOCYTOPLASMIC_TRANSPORT",
                    "GOBP_MRNA_TRANSPORT")

genesets_genes <- genesets %>% 
  filter(term %in% selected_terms) %>%
  distinct(gene) %>%
  pull(gene)

# Filter significantly differentially expressed genes in selected gene sets
df_filtered <- merged_df %>%
  filter(comb_padj < 0.05, Symbol %in% genesets_genes) %>%
  mutate(across(c(log2FoldChange_fpmtb, log2FoldChange_beataml),
                ~pmin(pmax(., -0.8), 0.8)))  # Cap values at ±0.8

# Inspect resulting dataframe
head(df_filtered)

# Define quadrant group
df_filtered <- df_filtered %>%
  mutate(
    quadrant = case_when(
      log2FoldChange_fpmtb >= 0 & log2FoldChange_beataml >= 0 ~ "Q1: Up in both",
      log2FoldChange_fpmtb < 0 & log2FoldChange_beataml >= 0  ~ "Q2: Down in fpmtb, Up in beatAML",
      log2FoldChange_fpmtb < 0 & log2FoldChange_beataml < 0   ~ "Q3: Down in both",
      log2FoldChange_fpmtb >= 0 & log2FoldChange_beataml < 0  ~ "Q4: Up in fpmtb, Down in beatAML"
    )
  )

# Plot
p <- ggplot(df_filtered, aes(x = log2FoldChange_fpmtb, y = log2FoldChange_beataml)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(size = -log10(comb_padj), fill = quadrant), shape = 21, stroke = 0.2, color = "black") +
  geom_text_repel(
    data = filter(df_filtered, quadrant == "Q1: Up in both"),
    aes(label = Symbol),
    size = 3,
    box.padding = 0.3,
    max.overlaps = Inf
  ) +
  scale_fill_manual(values = c(
    "Q1: Up in both" = "#D078A7",
    "Q2: Down in fpmtb, Up in beatAML" = "#00976A",
    "Q3: Down in both" = "#006CAF",
    "Q4: Up in fpmtb, Down in beatAML" = "#A09D9B"
  )) +
  scale_size_continuous(name = expression(-log[10](padj))) +
  theme_classic() +
  labs(
    x = "log2FC in FPMTB cohort",
    y = "log2FC in BeatAML cohort",
    fill = "Quadrant"
  )
p

ggsave(p, filename = "../03.out/figure6/dot_plots_two_cohorts.pdf", width = 7.2, height = 4.8)

```

